---
title: "To do"
output: html_notebook
---

* Make report from Flexgsea results (half done)
* Clustering with other comparisons (not just inv vs rest)
* GSEA with "interesting gene list"
* Variable influence factor (VIF) - Maybe
* Diffex and KM short vs long-term inv
* KMs instead of dot plots in 12B
* Check on whether the IGs found in involution are the same class as those typically found in breast milk
* Recreate code for mountain plots with flexgsea (Steal some ideas from ClusterProlifer? Or use UpSet plots instead?)

Open questions/theories from survival interaction analysis:

Do low B cell patients have a high fibrinogen signature (FGB/FGG/FGFBP1)? Could the fibrinogen be suppressing B cell activity, resulting in increased mortality?
What about Akt1 signaling (DUSP13 and FLT3) and cancer-testis antigens (MAGE family)?

## Odds & ends

ALTERNATE THRESHOLD possibility:

Performs a filtering pass on the dataset to remove any features (rows) without at least 1 read per million in n of the samples, where n is the size of the smallest group of replicates. Recommended in:
Anders S, McCarthy DJ, Chen Y, Okoniewski M, Smyth GK, Huber W, Robinson MD (2013). " [Count-based differential expression analysis of RNA sequencing data using R and Bioconductor." Nat. Protocols, 8, 1765-1786.](https://www.nature.com/articles/nprot.2013.099)

```{r}
library(DESeq2)
library(tidyverse)
colnames(colData(dds))
as.data.frame(colData(dds)[,c("study_group", "TIL_percent")]) %>% 
  filter(!is.na(TIL_percent)) %>%
  ggplot(aes(x = study_group, y = TIL_percent)) +
  geom_boxplot() +
  ggtitle("TIL percent")
```

```{r}
print("Fix")
```


