---
title: "Organize Vectra samples"
author: "Kat Moore and Tycho Bismeijer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
params:
    metadata: ""
    batch_info: ""
    example_halo_archive: ""
    vectra_dir: ""
    out: ""
    filedict: ""
---

Here we discover the Vectra data exported from Halo in the `data/vectra` directory.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, message=FALSE}
library(here)
library(lubridate)
library(tidyverse)
```

## Overview

In Nov 2021, we went from having a single exported batch to work with, to having all batches available.

```{r}
batches <- list.files(here(params$vectra_dir, 'halo'), pattern = "Batch")
print(batches)
```

There are two panels in each batch. The markers are as follows:

MPIF26 = CD3_CD20_CD27_FoxP3_PanCK
MPIF27 = CD3_CD8_CD138_CD20_PanCK

We cross-reference these samples with the list of known failed stainings later on.

### Sample Annotation

All samples now have a T-number.

```{r}
sample_annotation <- readxl::read_excel(here(params$metadata), 1)
head(sample_annotation)
```

### Directory structure

For every sample, the data is in a dated sub-directory.

```{r}
list.files(here(params$example_halo_archive))
```

Basic stats are in `summary_results.csv`.

```{r}
read_csv(here(params$example_halo_archive, "summary_results.csv"))
```

Full results are in an `object_results.csv` file.

```{r}
object_results_files <- list.files(here(params$example_halo_archive), "*.object_results.csv")
stopifnot(length(object_results_files) == 1)
read_csv(here(params$example_halo_archive, object_results_files[1])) %>% head()
```

The image segmentations are in the Images directory with a .annotations extension: these are xml
files. Unclear what the difference is between the two annotations files.

```{r}
list.files(here(params$example_halo_archive, "Images"))
```

## Reported failures in metadata

A number of analyses failed.

```{r}
sample_annotation %>%
  filter(sample_type == 'slide' & Included == 0) %>%
  select(sample_ID, experimental_platform, patient_ID, batch_Leuven, batch_HALO, comments)
```

### Duplicates

Some patients evidently provided more than one tissue block.

```{r}
patients_with_duplicate_slides <- sample_annotation %>%
    filter(sample_type == 'slide', Included == 1) %>%
    group_by(patient_ID, experimental_platform) %>%
    count() %>%
    filter(n > 1) %>%
    `$`('patient_ID') %>%
    unique()

sample_annotation %>%
  filter(
    sample_type == 'slide',
    patient_ID %in% patients_with_duplicate_slides) %>%
  arrange(patient_ID) %>%
  select(patient_ID, sample_ID, experimental_platform)
```

## Locate files in HALO archives

### Object results

The data is located in `r params$vectra_dir`/halo in directories, with the first directory
identifying the batch starting with batch, then a directory with the panel name, then a directory
with the t-number.


```{r}
object_files <- tibble(
  object_path = list.files(
    here(params$vectra_dir, 'halo', batches),
    recursive = T, full.names = T, pattern = ".object_results.csv"
  )) %>%
  #Extract the t_number, batch and panel from the object results
  mutate(
    split_object_path=fs::path_split(object_path),
    HALO_archive = map_chr(split_object_path, ~ .[[length(.)-5]]),
    HALO_batch = map_chr(split_object_path, ~ .[[length(.)-4]]) %>%
        str_to_lower() %>%
        str_remove_all(" "),
    panel = map_chr(split_object_path, ~ .[[length(.)-3]]),
    t_number = map_chr(split_object_path, ~ .[[length(.)-2]])) %>%
  select(-split_object_path) %>%
  relocate(object_path, .after = everything()) %>%
  relocate(t_number, .before = everything()) %>%
  arrange(t_number)

head(object_files)
```

The number of sample-panel combinations with an object file is `r nrow(object_files)` for
`r length(unique(object_files$t_number))` samples.

Sample-panel combinations for which we did not find an object file:

```{r}
anti_join(
  filter(sample_annotation, sample_type == 'slide' & Included == 1),
  object_files,
  by = c(sample_ID = 't_number', experimental_platform = 'panel'))
```

and those for which we have a file, but don't expect it:

```{r}
anti_join(
  object_files,
  filter(sample_annotation, sample_type == 'slide' & Included == 1),
  by=c(t_number = 'sample_ID', panel = 'experimental_platform'))
```

and check for doubles

```{r}
object_files %>%
  group_by(t_number, panel) %>%
  count() %>%
  filter(n > 1) %>%
  left_join(object_files, by=c('t_number', 'panel')) %>%
  arrange(t_number, panel)
```


### Summary results

This file is always called `summary_results.csv`, regardless of the sample it belongs to.

```{r}
summary_files <- tibble(
  summary_path = list.files(
    here(params$vectra_dir, 'halo', batches),
    recursive = T, full.names = T, pattern = "summary_results.csv"
  )) %>%
  #Extract the t_number, batch and panel from the summary results
  mutate(
    split_summary_path=fs::path_split(summary_path),
    HALO_archive = map_chr(split_summary_path, ~ .[[length(.)-5]]),
    HALO_batch = map_chr(split_summary_path, ~ .[[length(.)-4]]) %>%
        str_to_lower() %>%
        str_remove_all(" "),
    panel = map_chr(split_summary_path, ~ .[[length(.)-3]]),
    t_number = map_chr(split_summary_path, ~ .[[length(.)-2]])) %>%
  select(-split_summary_path) %>%
  relocate(summary_path, .after = everything()) %>%
  relocate(t_number, .before = everything()) %>%
  arrange(t_number)

head(summary_files)
```

The number of sample-panel combinations with an summary file is `r nrow(summary_files)` for
`r length(unique(summary_files$t_number))` samples.

Sample-panel combinations for which we did not find an summary file:

```{r}
anti_join(
  filter(sample_annotation, sample_type == 'slide' & Included == 1),
  summary_files,
  by = c(sample_ID = 't_number', experimental_platform = 'panel'))
```

and those for which we have a file, but don't expect it:

```{r}
anti_join(
  summary_files,
  filter(sample_annotation, sample_type == 'slide' & Included == 1),
  by=c(t_number = 'sample_ID', panel = 'experimental_platform'))
```

and check for doubles

```{r}
summary_files %>%
  group_by(t_number, panel) %>%
  count() %>%
  filter(n > 1) %>%
  left_join(summary_files, by=c('t_number', 'panel')) %>%
  arrange(t_number, panel)
```

### Annotation files

This file is has an extension of `.annotations` and is an XML file in the `Images` directory.

```{r}
annotations_files <- tibble(
  annotations_path = list.files(
    here(params$vectra_dir, 'halo', batches),
    recursive = T, full.names = T, pattern = ".annotations"
  )) %>%
  #Extract the t_number, batch and panel from the annotations results
  mutate(
    split_annotations_path=fs::path_split(annotations_path),
    HALO_archive = map_chr(split_annotations_path, ~ .[[length(.)-6]]),
    HALO_batch = map_chr(split_annotations_path, ~ .[[length(.)-5]]) %>%
        str_to_lower() %>%
        str_remove_all(" "),
    panel = map_chr(split_annotations_path, ~ .[[length(.)-4]]),
    t_number = map_chr(split_annotations_path, ~ .[[length(.)-3]]),
    job = grepl('job[0-9]+\\.annotations', annotations_path)) %>%
  select(-split_annotations_path) %>%
  relocate(annotations_path, .after = everything()) %>%
  relocate(t_number, .before = everything()) %>%
  arrange(t_number)

head(annotations_files)
```

The number of sample-panel combinations with an annotations file is `r nrow(annotations_files)` for
`r length(unique(annotations_files$t_number))` samples.

Sample-panel combinations for which we did not find an annotations file:

```{r}
anti_join(
  filter(sample_annotation, sample_type == 'slide' & Included == 1),
  annotations_files,
  by = c(sample_ID = 't_number', experimental_platform = 'panel'))
```

and those for which we have a file, but don't expect it:

```{r}
anti_join(
  annotations_files,
  filter(sample_annotation, sample_type == 'slide' & Included == 1),
  by=c(t_number = 'sample_ID', panel = 'experimental_platform'))
```

and check for doubles. Some samples have an extra file with job in the name.

```{r}
annotations_files %>%
  group_by(t_number, panel, job) %>%
  count() %>%
  filter(n > 1) %>%
  left_join(annotations_files, by=c('t_number', 'panel')) %>%
  arrange(t_number, panel)
```

### File dictionary

```{r}
filedict <- full_join(object_files, summary_files,
                      by=c("t_number", "panel", "HALO_batch", "HALO_archive"))
filedict <- full_join(filedict, filter(annotations_files, job == FALSE),
                      by=c("t_number", "panel", "HALO_batch", "HALO_archive"))
```

Combined we have `r nrow(filedict)` sample-panel combinations. 


## Symlink files into a more desired directory structure

Create directory structure.

```{r}
symdir <- here(params$vectra_dir, "raw")
dir.create(symdir, showWarnings = F)

dir.create(file.path(symdir, "objects"), showWarnings = F)
dir.create(file.path(symdir, "summary"), showWarnings = F)

#Deleting previous symlinks is best practice, so that if the input changes,
#the old files are not accidentally included
file.remove(list.files(file.path(symdir, "annotations"),
                       include.dirs = F, full.names = T, recursive = T))
file.remove(list.files(file.path(symdir, "annotations", "job_annotations"),
                       include.dirs = F, full.names = T, recursive = T))
file.remove(list.files(file.path(symdir, "objects"),
                       include.dirs = F, full.names = T, recursive = T))
file.remove(list.files(file.path(symdir, "summary"),
                       include.dirs = F, full.names = T, recursive = T))
```

### Object results

Add desired file locations to file dictionary.

```{r}
filedict$object_link <- file.path(
  symdir, "objects",
  paste(
    filedict$t_number,
    filedict$panel,
    filedict$HALO_batch,
    "object_results.csv",
    sep = "_"
  )
)
```

Create the symbolic links from desired location to actual location.

```{r}
#Will harmlessly give a "file exists" warning and return FALSE if already done
suppressWarnings(file.symlink(
  from = filedict$object_path,
  to = filedict$object_link))
```

### Summary results

Add symlinks to file dict.

```{r, eval=T}
filedict$summary_link <- file.path(
  symdir, "summary",
  paste(
    filedict$t_number,
    filedict$panel,
    filedict$HALO_batch,
    "summary_results.csv",
    sep = "_"))
```

Create the symbolic links.

```{r}
suppressWarnings(file.symlink(
  from = filedict$summary_path,
  to = filedict$summary_link
))
```

### Annotations

Add symlinks to file dict.

```{r, eval=T}
filedict$annotations_link <- file.path(
  symdir, "annotations",
  paste(
    filedict$t_number,
    filedict$panel,
    filedict$HALO_batch,
    "annotations.xml",
    sep = "_"))
```

Create the symbolic links.

```{r}
suppressWarnings(file.symlink(
  from = filedict$annotations_path,
  to = filedict$annotations_link
))
```

## Write output

Contains original file location and symlink path for all of the result types.

```{r}
dir.create(here(file.path(params$vectra_dir, "metadata")), showWarnings = F)
write_csv(filedict, here(params$filedict))
```

```{r}
sessionInfo()
```
