---
title: "IHC receptor status and PAM50"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
---

```{r, include=F}
library(DESeq2)
library(genefu)
library(here)
library(ggbeeswarm)
library(ggrepel)
library(ggpubr)
library(tidyverse)
```

# Load data

```{r}
dds <- readRDS(file = here("data/Rds/02_QC_dds.Rds"))
```

Study group and mol subtype overview

```{r}
table(dds$study_group)
table(dds$molecular_subtype)
table(dds$study_group, dds$molecular_subtype)
```

Extract coldata for convenience

```{r}
sd = as.data.frame(colData(dds))
```


## Variance stablizing transformation

From the documentation:

This function calculates a variance stabilizing transformation (VST) from the fitted dispersion-mean relation(s) and then transforms the count data (normalized by division by the size factors or normalization factors), yielding a matrix of values which are now approximately homoskedastic (having constant variance along the range of mean values).

The VST wrapper provides much faster estimation of the dispersion trend used to determine the formula for the VST. The speed-up is accomplished by subsetting to a smaller number of genes in order to estimate this dispersion trend. The subset of genes is chosen deterministically, to span the range of genes' mean normalized count. This wrapper for the VST is not blind to the experimental design: the sample covariate information is used to estimate the global trend of genes' dispersion values over the genes' mean normalized count. It can be made strictly blind to experimental design by first assigning a design of ~1 before running this function, or by avoiding subsetting and using varianceStabilizingTransformation. 

We want the transformation to be entirely blind to experimental design, so we first apply design of ~1 to dds before calling vst. Calling varianceStabilizingTransformation is extremely slow.

```{r}
design(dds) = formula(~ 1)

vsd = vst(dds, blind=T)
```


# Receptor expression

* HER2: "ENSG00000141736"
* ER:   "ENSG00000091831"
* PGR:  "ENSG00000082175"

Subset transformed data to these receptors of interest:

```{r}
roi = tibble(gene_name = c("HER2", "ER", "PR"),
             gene_id = c("ENSG00000141736","ENSG00000091831", "ENSG00000082175"))
roi

vroi = assay(vsd[rownames(vsd) %in% roi$gene_id, ])
```

```{r}
vroi[1:3,1:3]
```


Create plot-able data frame

```{r}
vsub = vroi %>% as.data.frame() %>% rownames_to_column("gene_id") %>%
  gather(key = "sample_name",-gene_id, value = "normCount") %>%
  left_join(roi, ., by="gene_id") %>%
  left_join(.,select(sd, sample_name, HER2, ER, PR, molecular_subtype), by= "sample_name")

head(vsub)
```

## Overview receptor expression by study group

```{r}
vsub %>% ggplot(aes(x=molecular_subtype, y = normCount)) + geom_boxplot() +
  facet_grid(~gene_name) + theme(axis.text.x = element_text(angle = 90))
```

Function for plotting separate receptors

```{r}
plot_receptor = function(data, receptor, x.column="molecular_subtype", y.column="normCount"){
  
  require(tidyverse)
  require(ggbeeswarm)
  
  data %>% filter(gene_name == !!receptor) %>% 
    ggplot(aes(x=get(x.column), y = get(y.column), color = get(receptor))) +
    ggbeeswarm::geom_beeswarm() +
    xlab(x.column) + ylab(y.column) + labs(color=paste(receptor, "IHC status")) + 
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle(paste(receptor, "variance stabilized transformed mRNA expression"))
}
```

Factors will help us out while plotting.

```{r}
vsub$HER2 = factor(vsub$HER2)
vsub$ER = factor(vsub$ER)
vsub$PR = factor(vsub$PR)
```


## Her2 IHC

ppbc_inv_124 is a clear outlier within the LumB_Her2neg group.

```{r}
her2_outlier = vsub[vsub$HER2=="0" &
                        vsub$gene_name=="HER2" &
                        vsub$normCount > 15, ]

plot_receptor(vsub, "HER2") +
  geom_text_repel(aes(label=sample_name), show.legend = F,
            data=her2_outlier)
```


## ER IHC

non_prbc_143, 45, and 105 are notable outliers here.

```{r}
er_outlier = vsub[vsub$ER == 0 &
                        vsub$gene_name=="ER" &
                        vsub$normCount > 11, ]


plot_receptor(vsub, "ER") +
  geom_text_repel(aes(label=sample_name), show.legend = F,
                  data=er_outlier)
```


## PR IHC

```{r}
pr_outlier = vsub[vsub$PR == 0 &
                        vsub$gene_name=="PR" &
                        vsub$normCount > 10, ]


plot_receptor(vsub, "PR") +
  geom_text_repel(aes(label=sample_name),
                  data=pr_outlier, show.legend = F)

```

## Outlier overview

```{r}
ihc_outlier =
right_join(select(sd, sample_name, HER2, ER, PR),
  bind_rows(
  tibble(sample_name = her2_outlier$sample_name,
       ihc_outlier = "HER2_high"),
  tibble(sample_name=er_outlier$sample_name,
       ihc_outlier = "ER_high"),
    tibble(sample_name=pr_outlier$sample_name,
       ihc_outlier = "PR_high")),
  by="sample_name")

ihc_outlier
```

```{r}
write_csv(ihc_outlier, path = here("results/qc/03_ihc_outliers.csv"))
```


# PAM50 determination

Our annotation file must include columns with this nomenclature:

* EntrezGene.ID
* Gene name
* Gene.Symbol
* Probe

The probe is the identifier used in the gene matrix, in this case ensembl id.

```{r}
entrez_annot = read_tsv(here("data/external/entrez_biomart_ensemblgenes_v94.txt"))

entrez_annot = entrez_annot %>%
  rename(probe = `Gene stable ID`,
         gene_type = `Gene type`,
         EntrezGene.ID = `NCBI gene ID`) %>%
  mutate(Gene.symbol = `Gene name`) %>%
  filter(!is.na(EntrezGene.ID))

head(entrez_annot)
```

## Deduplicate annotation file

Deal with the biomart artefact of multiple EntrezIDs being retrieved for certain ensembl IDs and vice versa. Make sure none of these duplicates are PAM50 genes.

```{r}
deduplicate_annotation <- function(annot){
  
  require(genefu)
  data('pam50', package='genefu', envir=environment())
  pam50.names <- rownames(pam50$centroids)
  dups <- annot[which(duplicated(annot$probe)), "Gene.symbol"] 
  print("PAM50 genes with multiple ensembl ids:")
  print(table(pam50.names %in% dups))
  
  annot <- annot[order(annot$EntrezGene.ID),]
  annot <- annot[!(duplicated(annot$probe)),]

  #Complains about being a tibble
  annot <- as.data.frame(annot)
  rownames(annot) <- annot$probe
  
  stopifnot(sum(duplicated(annot$probe))==0)
  
  return(annot)
}
```

```{r}
entrez_annot = deduplicate_annotation(entrez_annot)
```

## Voom transformation

We will need to voom-transform our count data before running the classification. (Source: Tycho. To ask: Can we use the vst instead? Suspected answer: No, because the voom transformation brings the count matrix to the same order as a typical microarray, whereas a vst does not.)

```{r}
voom_from_deseq <- function(dds){
  
  require(DESeq2)
  require(edgeR)
  require(limma)
  
  rawcount = counts(dds, normalized=F)
  dge <- edgeR::DGEList(rawcount)
  dge <- edgeR::calcNormFactors(dge, method='TMM')
  voom <- limma::voom(dge, plot=F)
  vt <- t(voom$E) #Must be transposed for genefu compatibility
  return(vt)
}

```

```{r}
vt = voom_from_deseq(dds)
```

Ensure that the annotation file has only the genes present in the data set

```{r Filter annotation file}
entrez_annot <- entrez_annot %>% filter(probe %in% colnames(vt))
```

## Genefu

```{r Molecular subtyping}
pred <- genefu::molecular.subtyping(sbt.model="pam50", data = vt, annot=entrez_annot, do.mapping = T)
table(pred$subtype)
```

Add PAM50 info to dds coldata.

```{r}
stopifnot(identical(colData(dds)[,"sample_name"], names(pred$subtype)))
dds$PAM50 = pred$subtype
```

## Confusion matrix PAM50 vs molecular subtype

Confusion is greatest for LumB Her2+ samples: They end up spread between Her2, LumB and LumA.


```{r}
table(dds$PAM50, dds$molecular_subtype)
```

## Barplot confusion matrix

```{r}
barplot_from_table = function(table, xlab, ylab){
  df = as.data.frame(table)
  df %>% ggplot(aes(x=Var1, y = Freq, fill=Var2)) + 
    geom_bar(stat="identity") + xlab(xlab) + labs(fill=ylab)
}
```


```{r}
barplot_from_table(table(dds$PAM50, dds$molecular_subtype),
                   xlab="PAM50", ylab="molecular_subtype") +
  ggtitle("Visualizing the confusion matrix between PAM50 and molecular subtype")
```

# PCA PAM50 vs molecular subtype

```{r}
get_PCA <- function(dds, labelled_samples = NULL, color.group, shape.group){
  
  require(DESeq2)
  require(tidyverse)
  require(ggrepel)
  
  vsd = vst(dds, blind=FALSE)
  
  sampleDists <- dist(t(assay(vsd)))
  
  pcaData <- plotPCA(vsd, intgroup=c(color.group, shape.group), returnData=TRUE) #To do: Make intgroup variables
  percentVar <- round(100 * attr(pcaData, "percentVar"))
  
  if (is.null(labelled_samples)==T){
    ggplot(pcaData, aes(PC1, PC2, color=get(color.group), shape=get(shape.group))) +
      geom_point(size=3) +
      xlab(paste0("PC1: ",percentVar[1],"% variance")) +
      ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
      coord_fixed() + labs(color = color.group, shape = shape.group)
  } else {
    pcaData = pcaData %>% mutate(labels = if_else(name %in% labelled_samples, name, NULL))
    ggplot(pcaData, aes(PC1, PC2, color=get(color.group), shape=get(shape.group), label=labels)) +
      geom_point(size=3) +
      xlab(paste0("PC1: ",percentVar[1],"% variance")) +
      ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
      coord_fixed() + ggrepel::geom_label_repel(size=4, show.legend = F) +
      labs(color = color.group, shape = shape.group)
  }

}
```

```{r}
pca_pam50 = get_PCA(dds, color.group = "PAM50", shape.group = "study_group") +
  ggtitle("PCA using PAM50")
```

```{r}
pca_molsub = get_PCA(dds, color.group = "molecular_subtype", shape.group = "study_group") +
  ggtitle("PCA using molecular subtype")
```

PAM50 clearly results in the better clustering.

```{r, fig.width=10}
ggarrange(pca_pam50, pca_molsub)
```

Label the outliers from the IHC receptor expression analysis.

```{r}
pca_molsub_labels = get_PCA(dds, color.group = "molecular_subtype", shape.group = "study_group",
                            labelled_samples = ihc_outlier$sample_name) +
  ggtitle("PCA using molecular subtype, IHC outliers")
```

It doesn't look like the IHC outliers generally cluster away from their own group.

```{r}
pca_molsub_labels
```

# Receptor status and PAM50

We'll need a new data frame that includes the PAM50 classes

```{r}
vpam = vroi %>% as.data.frame() %>% rownames_to_column("gene_id") %>%
  gather(key = "sample_name",-gene_id, value = "normCount") %>%
  left_join(roi, ., by="gene_id") %>%
  left_join(.,select(as.data.frame(colData(dds)), sample_name, HER2, ER, PR, PAM50, molecular_subtype), by= "sample_name")

vpam$HER2 = factor(vpam$HER2)
vpam$ER = factor(vpam$ER)
vpam$PR = factor(vpam$PR)

head(vpam)
```

## Overview receptor expression by study group

It seems that HER2 expression is by far the most ambiguous among the PAM50 classifications.

```{r}
vpam %>% ggplot(aes(x=PAM50, y = normCount)) + geom_boxplot() +
  facet_grid(~gene_name) + theme(axis.text.x = element_text(angle = 90))
```


Seems like most of the subgroups have a trail which is more HER2 positive than the rest. How sure is genefu about these samples?

Reduce the 5 probability columns (1 per subtype) to just the probabilty of the assigned class (which is always the max of the columns).

```{r}
get_max_prob = function(prob_matrix, samples=rownames(prob_matrix)){
  mat = prob_matrix[rownames(prob_matrix) %in% samples, ]
  max_prob = apply(mat, 1, function(x) max(x, na.rm=T))
  
  colmax = colnames(mat)[max.col(mat)]
  
  df = tibble(sample_name = names(max_prob),
              prediction = colmax,
              probability = max_prob)
  
  return(df)
}
```


Retrieve a list of probabilties by prediction. Add them to coldata.

```{r}
prob_pred = get_max_prob(pred$subtype.proba)
head(prob_pred)

stopifnot(identical(prob_pred$sample_name,dds$sample_name))
dds$PAM50_prob = prob_pred$probability
```

Add certainity to the facet plot by color

## Overview by probability

```{r}
left_join(prob_pred, vpam, by="sample_name") %>%
  ggplot(aes(x=PAM50, y = normCount, color=probability)) +
  ggbeeswarm::geom_beeswarm() +
  facet_grid(~gene_name) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_gradient(low="gray", high="black")
```

## Her2 PAM50 probability

Her2 is the trickiest sample to classify.

```{r}
plot_receptor(left_join(prob_pred, vpam, by="sample_name"), "HER2", x.column="PAM50") +
  aes(color=probability, shape=HER2) + labs(shape="HER2 IHC") +
  scale_color_gradient(low="gray", high="darkblue")
```

At the very least we should include the high triple negative samples in the list of those to check.

```{r}
pam50_outliers = filter(vpam, PAM50 == "Basal" & gene_name=="HER2" & normCount>14) %>%
  select(sample_name, HER2, ER, PR, PAM50) %>% mutate(pam50_outlier = "highHER2_tripleneg")
pam50_outliers
```


## ER PAM50 probability

No major outliers stand out here.

```{r}
plot_receptor(left_join(prob_pred, vpam, by="sample_name"), "ER", x.column="PAM50") +
  aes(color=probability, shape=ER) + labs(shape="ER IHC") +
  scale_color_gradient(low="gray", high="darkred")
```


## PR PAM50 probability

Once again we have a clear outlier in the basal group.

```{r}
plot_receptor(left_join(prob_pred, vpam, by="sample_name"), "PR", x.column="PAM50") +
  aes(color=probability, shape=PR) + labs(shape="PR IHC") +
  scale_color_gradient(low="gray", high="darkgreen")
```

To check: which is the obvious outlier in the basal samples?


```{r}
pam50_outliers = bind_rows(pam50_outliers, 
          filter(vpam, PAM50 == "Basal" & gene_name=="PR" & normCount>10) %>%
  select(sample_name, HER2, ER, PR, PAM50) %>% mutate(pam50_outlier = "highPR_tripleneg"))
pam50_outliers
```

# Remove PAM50 outliers

Were these samples already in the to-check list?

```{r}
pam50_outliers$sample_name %in% ihc_outlier$sample_name
```

```{r}
write_csv(pam50_outliers, path = here("results/qc/pam50_outliers.csv"))
```

Subset the data frame based on this.

```{r}
print(paste("Prior to discarding outliers, there are", ncol(dds), "samples"))

dds = dds[,!colnames(dds) %in% pam50_outliers$sample_name]

print(paste("After discarding PAM50 outliers, there are", ncol(dds), "samples"))
```


# Write dds data

```{r}
saveRDS(dds, file = here("data/Rds/03_dds_PAM50.Rds"))
write_csv(as.data.frame(colData(dds)), path=here("data/metadata/03_sample_annot_filtered_PAM50.csv"))
```

```{r}
save.image(here("reports/03_PAM50_IHC.Rdata"))
```

