---
title: "Tumor purity"
output: html_notebook
---

```{r, include=F}
library(DESeq2)
library(here)
library(estimate)
library(tidyverse)
```

# Load data


```{r}
dds <- readRDS(here("data/Rds/03_dds_PAM50.Rds"))

tx_annot <- read_tsv(here("data/metadata/01_tx_annot.tsv"))
```

We only need the gene id and hgnc symbol for this script.

```{r}
tx_annot = tx_annot %>% select(gene_id, hgnc_symbol) %>% distinct()
```

# Conversion to FPKM with HUGO symbols

Esimate requires transcript length normalization and gene names instead of ensembl ids.

Debatable: whether to use robust = T or F when calculating fpkm for the purposes of determining tumor purity with Estimate. When True, estimateSizeFactors is called to more accurately normalize by library size. Use the robust = F option to disable the median ratio normalization and just get classical fpkm via colSums.

Since we prefer a more robust correction for library size, choose robust = T unless we hear otherwise.

```{r}
fpkm <- fpkm(dds, robust = T)
```

Add the HUGO symbols

```{r}
fpkm <- fpkm %>% as.data.frame() %>% rownames_to_column("gene_id") %>% right_join(select(tx_annot, gene_id, hgnc_symbol),., "gene_id")

head(fpkm)
```

Make sure we don't have any NAs in the HGNC field.

```{r}
print(paste("Removing", sum(is.na(fpkm$hgnc_symbol)), "genes without HUGO ids from the fpkm matrix."))
fpkm <- fpkm %>% filter(!is.na(hgnc_symbol))
```

```{r}
print(paste(nrow(fpkm), "genes remaining in the fpkm matrix."))
```

Drop the ensembl id column.

```{r}
fpkm <- fpkm %>% select(-gene_id)
```

# De-duplicate gene symbols

Non-unique identifiers are the bane of bioinformaticians everywhere. 

When dealing with duplicated gene names, take the mean fpkm.

```{r}
summarize_expression_duplicate_ids <- function(mat, id_column, f=colMeans, final_gene_symbol_colname="GeneSymbol"){
  require(dplyr)
  
  print(paste("Starting with gene expression matrix containing", nrow(mat), "rows."))
  
  #Easiest way to write functions with dplyr is to standardize the column name
  
  if(id_column != "symbol"){
    colnames(mat)[colnames(mat)==id_column] <- "symbol"
  }
  
  #Make frequency table
  id_table <- as.data.frame(table(mat$symbol))
  
  #Identify duplicate genes
  dups <- id_table$Var1[id_table$Freq > 1]
  stopifnot(length(dups) == length(unique(dups)))
  print(paste("Number of genes with duplicate names:", length(dups)))
  
  #Set aside rows with unique gene names
  nodup_df <- mat[!mat$symbol %in% dups,]
  
  #Set aside rows with duplicate ids
  dup_df <- mat[mat$symbol %in% dups,]
  stopifnot(nrow(nodup_df) + nrow(dup_df) == nrow(mat))
  
  #Sort by recurring id
  dup_df <- dup_df[order(dup_df$symbol),]
  
  print(paste("Number of rows with duplicate gene ids:", nrow(dup_df)))
  
  #Mean expression fpkm of genes with the same symbol
  mean_exps <- matrix(ncol = ncol(dup_df)-1, nrow=0) #Empty matrix, -1 gene symbol column

  for (i in 1:length(unique(dup_df$symbol))){
    #Subset rows with same symbol, discard symbol column, then apply aggregate function
    exp <- f(as.matrix(dup_df[dup_df$symbol==unique(dup_df$symbol)[i], -1])) 
    mean_exps <- rbind(mean_exps, exp)
  }
  stopifnot(nrow(mean_exps) == length(unique(dup_df$symbol)))
  
  rownames(mean_exps) <- unique(dup_df$symbol)
  mean_exps <- as.data.frame(mean_exps) %>% rownames_to_column("symbol")
  
  dedupped_df <- rbind(mean_exps, nodup_df)
  dedupped_df <- dedupped_df[order(dedupped_df$symbol),]
  
  stopifnot(length(unique(dedupped_df$symbol))==length(dedupped_df$symbol)) #All symbols should not be unique
  stopifnot(nrow(mat) - #starting number
            nrow(dup_df) + #rows with duplicate genes...
            length(dups) == #...which condense down into this many unique genes...
            nrow(dedupped_df)) #...should equal the number of rows in the final matrix

  print(paste("Number of genes after applying", substitute(f),  "to duplicate ids:", nrow(dedupped_df)))
  
  #For estimate, the column with identifiers HAS to be called GeneSymbol or EntrezGeneID
  colnames(dedupped_df)[colnames(dedupped_df)=="symbol"] <- final_gene_symbol_colname
  
  return(dedupped_df)
}
```

```{r}
clean_fpkm <- summarize_expression_duplicate_ids(fpkm, "hgnc_symbol")
```

```{r}
head(clean_fpkm)
```


The input for ESTIMATE can supposedly be a dataframe, but the function throws an error unless provided with a file..

```{r}
write.table(clean_fpkm, sep = "\t", file = here("data","RNA-seq","hugo_fpkm.txt"), col.names = T, row.names =F, quote=F)
```

# ESTIMATE

Estimate really only has two functions of interest for us. The first is filterCommonGenes, which unifies genes symbols provided versus the 10,412 used for determining tumor purity.

Annoyingly, it insists on writing a file instead of returning an object within the R environment. 


```{r}
dir.create(here("results","tumor_purity"))

filterCommonGenes(input.f=here("data","RNA-seq","hugo_fpkm.txt"),
                  output.f = here("results","tumor_purity","filterCommonGenes.gct"), id="GeneSymbol")
```

The next function computes stromal, immune, and ESTIMATE scores per sample using gene expression data. It also insists on writing files.

```{r}
estimateScore(input.ds = here("results","tumor_purity","filterCommonGenes.gct"),
              output.ds = here("results","tumor_purity","results_estimate_score.gct"),
              platform="illumina")
```

```{r}
scores <- read.delim(here("results","tumor_purity","results_estimate_score.gct"), skip=2, sep="\t")
head(scores)
```

## Purity score calculation

Note: Only validated for microarray data:
Tumour purity=cos (0.6049872018+0.0001467884 × ESTIMATE score)
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3826632/

```{r}
purity_score <- scores %>%dplyr::select(-Description) %>% rename(score=NAME) %>% gather(key="sample_name", -score, value=value) %>%
  mutate(sample_name=str_replace_all(sample_name, "\\.","-")) %>% spread(key=score,value=value) %>%
  mutate(tumor_purity = cos(0.6049872018+0.0001467884 * ESTIMATEScore))
```

## Visualizations

```{r}
scores %>% dplyr::select(-Description) %>% gather(key = "sample_name", -NAME, value="Score") %>%
  left_join(.,as.data.frame(colData(dds)), by="sample_name") %>% arrange(study_group) %>%
  ggplot(aes(x=factor(sample_name, levels = unique(sample_name)),y=Score, fill=study_group)) +
  geom_bar(stat="identity") + facet_wrap(~NAME, ncol=1) + theme(axis.text.x=element_blank()) +
  xlab("Samples")
``` 

```{r}
purity_score %>% gather(key=score, -sample_name, value=value) %>%left_join(.,as.data.frame(colData(dds)), by="sample_name") %>%
  arrange(study_group) %>%
  ggplot(aes(x=factor(sample_name, levels = unique(sample_name)),y=value, fill=study_group)) +
  geom_bar(stat="identity") + facet_wrap(~score, ncol=1, scales="free") + theme(axis.text.x=element_blank()) +
  xlab("Sample")
```

```{r}
coldata_purity <- purity_score %>% select(sample_name, tumor_purity) %>% left_join(.,as.data.frame(colData(dds)), by="sample_name")

coldata_purity %>% ggplot(aes(x=study_group, y=tumor_purity)) + geom_boxplot() + ggtitle("Tumor purity by study group")
```

# Save tumor purity for downstream analyses

```{r}
head(coldata_purity)
```

Add tumor purity to dds coldata

```{r}
stopifnot(colData(dds)$sample_name == coldata_purity$sample_name)
dds$tumor_purity = coldata_purity$tumor_purity
design(dds) <- ~PAM50 + tumor_purity + study_group
```


Write both the coldata and the new dds object

```{r}
write.csv(coldata_purity, file=here("data","metadata", "04_sample_annot_filtered_PAM50_purity.csv"), row.names = F)
saveRDS(dds, file=here("data","Rds","04_dds_PAM50_tumorpurity.Rds"))
```


```{r}
save.image(here("reports","04_tumor_purity.RData"))
```
