---
title: "Genewise reports"
Author: Kat Moore
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    theme: darkly
    highlight: breezedark
    canon: true
---

```{r, include=F}
rm(list = ls())
```

```{r, include=F}
library(DESeq2)
library(ComplexHeatmap)
library(tidyverse)
library(here)
```

# Load data {.tabset}

Count matrix from LRT. Counts will be the same between all comparisons.

```{r}
dds = readRDS(here("data/Rds/08_dds_ovr_inv_vs_rest.Rds"))
```

Gene annotation

```{r}
gx_annot <- read_tsv(here("data/metadata/01_tx_annot.tsv"))
gx_annot = gx_annot %>% select(ensembl_gene_id = gene_id, gene_name,
                               gene_type, description = gene_description) %>% distinct()
```

```{r}
source(here("src", "deseq_report_functions.R"))
```

## Variance stabilizing transformation

Renders the data homoskedastic, useful for visualization purposes.

Ensure that we can set "study group" to be one vs rest columns when desired.

```{r}
vsd = readRDS(here("data","Rds","08_vsd_ovr.Rds"))
```

## Sample annotation data

```{r}
sd = as.data.frame(colData(dds)) %>%
  select(sample_name, study_group, contains("rest"),
         PAM50, everything())
head(sd)
```

## Symbol conversion and de-deuplication

```{r}
geneEx = rownames_to_column(as.data.frame(assay(vsd)), "ensembl_gene_id")
geneEx = right_join(select(gx_annot, gene_name, ensembl_gene_id),
                    geneEx, by = "ensembl_gene_id") %>%
    select(-ensembl_gene_id)
geneEx = summarize_expression_duplicate_ids(geneEx, id_column = "gene_name", verbose = T)

```

## Colors

```{r}
cp = readRDS(here("data", "Rds", "color_palettes.Rds"))
study_colors = cp$study_colors
pam_colors = cp$pam_colors
gene_colors = cp$gene_colors
ppbc_colors = cp$ppbc_colors

sp = readRDS(here("data/Rds/survival_colors.Rds"))
```


## Diffex results

Likelihood ratio tests, pairwise comparisons, and one vs rest.

For the LRT, significant and all genes are kept in the same Excel output. Pairwise and one vs rest comparisons report significant genes separately. Here, we load all genes so that we can plot expression regardless of significance within a given comparison.

```{r, collapse=T}
lrt <- here("results", "diffex", "06_LRT_allgenes.xlsx") %>% 
  readxl::excel_sheets() %>% 
  set_names() %>% 
  map(readxl::read_excel, path = here("results", "diffex", "06_LRT_allgenes.xlsx"))
names(lrt)

pw <- here("results", "diffex", "07_pairwise_comparisons_allgenes.xlsx") %>% 
  readxl::excel_sheets() %>% 
  set_names() %>% 
  map(readxl::read_excel, path = here("results", "diffex", "07_pairwise_comparisons_allgenes.xlsx"))
names(pw)

ovr = here("results", "diffex", "08_one_vs_rest_allgenes.xlsx") %>% 
  readxl::excel_sheets() %>% 
  set_names() %>% 
  map(readxl::read_excel, path = here("results", "diffex", "08_one_vs_rest_allgenes.xlsx"))
names(ovr)

```

# Process and combine

Substitute lrt down to all genes and rename it for convenience.

```{r}
lrt = lrt["repLRT_all"]

names(lrt) = "LRT"
```

Combine all results into single list.

```{r}
res_list = append(lrt, pw)
res_list = append(res_list, ovr)
names(res_list) = str_remove_all(str_remove_all(names(res_list), "rep_"), "_all")
names(res_list) = str_replace(names(res_list), "_", "_vs_")
names(res_list)
```
"Quirky" genes

Genes marked with * downstream are highlighted for appearing to be more/less differentially expressed than they are reported. These include:
TACI, BCMA, NFKB1

# Differentially expressed genes

## Inv vs rest {.tabset}

### [JChain](https://www.genecards.org/cgi-bin/carddisp.pl?gene=JCHAIN&keywords=jchain)

Frequently reoccuring significant hit (inv vs rest)

>Serves to link two monomer units of either IgM or IgA. In the case of IgM, the J chain-joined dimer is a nucleating unit for the IgM pentamer, and in the case of IgA it induces larger polymers. It also help to bind these immunoglobulins to secretory component.

```{r}
gene_report(res_list,"ENSG00000132465")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000132465",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

### [POTEC](https://www.genecards.org/cgi-bin/carddisp.pl?gene=POTEC&keywords=potec)

Inv vs rest significant hit. Part of the the cancer/testis (CT) antigen family.

Cancer/testis (CT) genes represent a unique class of genes, which are expressed by germ cells, normally silenced in somatic cells, but activated in various cancers. We have previously reported that CTs are relatively commonly expressed in estrogen receptor (ER) negative, high risk carcinomas. In this study, we examined the expression of selected CT genes in ductal carcinoma in situ (DCIS), lobular carcinoma in situ (LCIS) and benign proliferative lesions of the breast. ER negative DCIS were found to be associated with significant CT gene expression together with HER2 positivity and a marked stromal immune response. See also [this article](https://www.ncbi.nlm.nih.gov/pubmed/19719775/).

```{r}
gene_report(res_list, "ENSG00000183206")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000183206",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

#### POTEC and ER relationship

Is POTEC more expressed in ER- breast cancers?
ESR1 = ENSG00000091831
ESR2 = ENSG00000140009

Here the opposite appears to be true : POTEC is more highly expressed in ESR postiive tumors.

```{r}
rbind(
  plotCounts(dds, gene="ENSG00000091831", intgroup="PPBC", normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000091831", gene_name = "ESR1"),
  #Expression of ESR2 is quite low
  #plotCounts(dds, gene="ENSG00000140009", intgroup=c("study_group","PAM50"), normalize=T, returnData=TRUE) %>%
   # rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000183206", gene_name = "ESR2"),
  plotCounts(dds, gene="ENSG00000183206", intgroup="PPBC", normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000183206", gene_name = "POTEC")
  ) %>% select(-ensembl_gene_id) %>%
  spread(key = gene_name, value=count) %>%
  ggplot(aes(x = ESR1, y= POTEC, color=PPBC)) + geom_jitter(height=0) +
  scale_y_log10() + scale_x_log10() +
  ggtitle("Relationship between POTEC and ESR1, study group") +
  theme_bw() +
  scale_color_manual(values = ppbc_colors)

rbind(
  plotCounts(dds, gene="ENSG00000091831", intgroup=c("study_group","PAM50"), normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000091831", gene_name = "ESR1"),
  plotCounts(dds, gene="ENSG00000183206", intgroup=c("study_group","PAM50"), normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000183206", gene_name = "POTEC")
  ) %>% select(-ensembl_gene_id) %>%
  spread(key = gene_name, value=count) %>%
  ggplot(aes(x = ESR1, y= POTEC, color=PAM50)) + geom_jitter(height=0) +
  scale_y_log10() + scale_x_log10() +
  ggtitle("Relationship between POTEC and ESR1, PAM50") +
  theme_bw() +
  scale_color_manual(values = pam_colors)
```

### [POTED](https://www.genecards.org/cgi-bin/carddisp.pl?gene=POTED&keywords=poted)

As POTEC above.

```{r}
gene_report(res_list, "ENSG00000166351")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000166351",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

#### POTED and ER relationship

```{r}
rbind(
  plotCounts(dds, gene="ENSG00000091831", intgroup="PPBC", normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000091831", gene_name = "ESR1"),
  #Expression of ESR2 is quite low
  #plotCounts(dds, gene="ENSG00000140009", intgroup=c("study_group","PAM50"), normalize=T, returnData=TRUE) %>%
   # rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000183206", gene_name = "ESR2"),
  plotCounts(dds, gene="ENSG00000166351", intgroup="PPBC", normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000166351", gene_name = "POTED")
  ) %>% select(-ensembl_gene_id) %>%
  spread(key = gene_name, value=count) %>%
  ggplot(aes(x = ESR1, y= POTED, color=PPBC)) + geom_jitter(height=0) +
  scale_y_log10() + scale_x_log10() +
  ggtitle("Relationship between POTED and ESR1, study group") +
  theme_bw() +
  scale_color_manual(values = ppbc_colors)

rbind(
  plotCounts(dds, gene="ENSG00000091831", intgroup=c("study_group","PAM50"), normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000091831", gene_name = "ESR1"),
  plotCounts(dds, gene="ENSG00000166351", intgroup=c("study_group","PAM50"), normalize=T, returnData=TRUE) %>%
    rownames_to_column("sample_name") %>% mutate(ensembl_gene_id = "ENSG00000166351", gene_name = "POTED")
  ) %>% select(-ensembl_gene_id) %>%
  spread(key = gene_name, value=count) %>%
  ggplot(aes(x = ESR1, y= POTED, color=PAM50)) + geom_jitter(height=0) +
  scale_y_log10() + scale_x_log10() +
  ggtitle("Relationship between POTED and ESR1, PAM50") +
  theme_bw() +
  scale_color_manual(values = pam_colors)
```

### [CXADRP3](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CXADRP3)

Inv vs rest significant hit. One of many genes typically expressed in the prostate or testis that is surprisingly found to be differentially expressed in ppbc.

Coxsackie Virus And Adenovirus Receptor Pseudogene 3 

```{r}
gene_report(res_list, "ENSG00000265766")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000265766",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

### [ESR2](https://www.genecards.org/cgi-bin/carddisp.pl?gene=ESR2&keywords=ESR2)

Estrogen receptor 2. Lowly expressed, but significant in inv vs rest.

Oddly, shows a similar expression pattern as BCMA, but is reported as much more significant. Possibly due to the presence of an outlier in the involution group?

```{r}
gene_report(res_list,"ENSG00000140009")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000140009",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000140009",],
                  intgroup = "batch", colorby="PPBC", color_vals = ppbc_colors,
                  title_string = "Involution vs rest")
```

## LRT {.tabset}

### [LALBA](https://www.genecards.org/cgi-bin/carddisp.pl?gene=LALBA&keywords=lalba)

Common milk protein, upregulated in pregnancy. Sometimes subject to Cook's distance filtering due to extreme outliers. Technically not diffex in LRT due to Cook's distance, but would be if that feature is disabled.

```{r}
gene_report(res_list, "ENSG00000167531")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000167531",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

### [GRN](https://www.genecards.org/cgi-bin/carddisp.pl?gene=GRN&keywords=GRN) 

Granulin precursor, mostly upregulated in pregnancy.

>Secreted proteins that act as key regulator of lysosomal function and as a growth factor involved in inflammation, wound healing and cell proliferation (PubMed:28541286, PubMed:28073925, PubMed:18378771, PubMed:28453791, PubMed:12526812). Functions as regulator of proteins trafficking to lysosome and activity of lysosomal enzymes (PubMed:28453791, PubMed:28541286). Facilitates also the acidification of lysosomes, causing degradation of mature CTSD by CTSB (PubMed:28073925). In addition, functions as wound-related growth factor that acts directly on dermal fibroblasts and endothelial cells to promote division, migration and the formation of capillary-like tubule structure (By similarity). Also promotes epithelial cell proliferation by blocking TNF-mediated neutrophil activation preventing release of oxidants and proteases (PubMed:12526812). Moreover, modulates inflammation in neuron by preserving neuron survival, axonal outgrowth and neuronal integrity (PubMed:18378771).

```{r}
gene_report(res_list, "ENSG00000030582")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000030582",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [IL18R1](https://www.genecards.org/cgi-bin/carddisp.pl?gene=IL18R1&keywords=ENSG00000115604)

Interleukin 18 receptor 1. Downregulated in pregnancy. Roughly equivalent in np and inv.

>FN-alpha and IL12 are reported to induce the expression of this receptor in NK and T cells. 

```{r}
gene_report(res_list, "ENSG00000115604")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000115604",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

# Immune cell markers

## B cells markers {.tabset}

### [CD20](https://www.genecards.org/cgi-bin/carddisp.pl?gene=MS4A1&keywords=CD20)

Official name of CD20 is MS4A1 ENSG00000156738

```{r}
gene_report(res_list,"ENSG00000156738")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000156738",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

```{r}
plot_gene_beehive(dds[,!is.na(dds$breastfeeding_duration)],
                  res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000156738",],
                  title_string = "Breastfeeding duration",
                  intgroup = "breastfeeding_duration",
                  colorby = "overall_survival") +
  ggtitle("Breastfeeding duration (CD20)")
```

```{r}
plot_gene_beehive(dds[,!is.na(dds$involution_duration)],
                  res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000156738",],
                  title_string = "Involution duration",
                  intgroup = "involution_duration",
                  colorby = "overall_survival") +
  ggtitle("Involution duration (CD20)")
```


### [IL10](https://www.genecards.org/cgi-bin/carddisp.pl?gene=IL10&keywords=IL-10)

B-reg marker IL10: ENSG00000136634

```{r}
gene_report(res_list,"ENSG00000136634")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000136634",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

### [SIRPA](https://www.genecards.org/cgi-bin/carddisp.pl?gene=SIRPA&keywords=SIRPA)

Previously significant in pregnancy vs rest, above threshold after removing tumor purity from design formula.

```{r}
gene_report(res_list,"ENSG00000198053")
```

```{r}
plot_gene_beehive(dds, res_list$prbc_vs_nonprbc[res_list$prbc_vs_nonprbc$ensembl_gene_id=="ENSG00000198053",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

Other plasma cells markers: CD138, CD98, TACI, BCMA
https://www.ncbi.nlm.nih.gov/pubmed/28787106

### [CD138](https://www.genecards.org/cgi-bin/carddisp.pl?gene=SDC1&keywords=CD138)

CD138 = SDC1 syndecan 1
[Plasma cell marker](https://www.ncbi.nlm.nih.gov/pubmed/28787106)

```{r}
gene_report(res_list,"ENSG00000115884")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000115884",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

```{r}
plot_gene_beehive(dds[,!is.na(dds$breastfeeding_duration)],
                  res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000115884",],
                  title_string = "Breastfeeding duration",
                  intgroup = "breastfeeding_duration",
                  colorby = "overall_survival") +
  ggtitle("Breastfeeding duration (SDC1)")
```

```{r}
plot_gene_beehive(dds[,!is.na(dds$involution_duration)],
                  res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000115884",],
                  title_string = "Involution duration",
                  intgroup = "involution_duration",
                  colorby = "overall_survival") +
  ggtitle("Involution duration (SDC1)")
```


### [CD98](https://www.genecards.org/cgi-bin/carddisp.pl?gene=SLC3A2&keywords=SLC3A2)

Official name SLC3A2 ENSG00000168003

>CD98 is an important amino acid transporter, also termed LAT1, which consists of heavy (SLC3A2) and light (SLC7A5) subunits. The increased expression of CD98 is most likely a reflection of the high metabolic requirements of ASCs (Antibody secreting cells).

```{r}
gene_report(res_list,"ENSG00000168003")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000168003",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

### [TACI](https://www.genecards.org/cgi-bin/carddisp.pl?gene=TNFRSF13B&keywords=taci)

Official name TNFRSF13B ENSG00000240505

```{r}
gene_report(res_list,"ENSG00000240505")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000240505",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

### *[BCMA](https://www.genecards.org/cgi-bin/carddisp.pl?gene=TNFRSF17&keywords=BCMA)

Official name TNFRSF17 ENSG00000048462

A good example of a gene which "looks" significant but is reported with a lower log2fc and higher p value than the graph would indicate. Similar to ESR2, which was reported with both higher fold changes and lower padj.

```{r}
gene_report(res_list,"ENSG00000048462")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000048462",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000048462",],
                  title_string = "Involution vs rest", intgroup = "inv_vs_rest",
                  colorby = "PPBC", color_vals = ppbc_colors)
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000048462",],
                  intgroup = "batch", colorby = "PPBC", color_vals = ppbc_colors, title_string = "Inv vs rest")
```

### [CD30](https://www.genecards.org/cgi-bin/carddisp.pl?gene=TNFRSF8&keywords=CD30)

[Marker for activation (anti-naive B cell marker)](https://www.abcam.com/primary-antibodies/b-cells-basic-immunophenotyping)

```{r}
gene_report(res_list,"ENSG00000120949")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000120949",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```


### [CD40](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CD40&keywords=cd40)

Memory B cell marker

```{r}
gene_report(res_list,"ENSG00000101017")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000101017",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```


### [MZB1](https://www.genecards.org/cgi-bin/carddisp.pl?gene=MZB1)

>Marginal Zone B And B1 Cell Specific Protein 
Associates with immunoglobulin M (IgM) heavy and light chains and promotes IgM assembly and secretion. May exert its effect by acting as a molecular chaperone or as an oxidoreductase as it displays a low level of oxidoreductase activity (By similarity). Isoform 2 may be involved in regulation of apoptosis. Helps to diversify peripheral B-cell functions by regulating Ca(2+) stores, antibody secretion and integrin activation.

```{r}
gene_report(res_list,"ENSG00000170476")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000170476",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000170476",],
                  intgroup = "batch", colorby = "PPBC", color_vals = ppbc_colors)

```

### BTLA[(https://www.genecards.org/cgi-bin/carddisp.pl?gene=BTLA&keywords=btla)]

>B- And T-Lymphocyte-Associated Protein This gene encodes a member of the immunoglobulin superfamily. The encoded protein contains a single immunoglobulin (Ig) domain and is a receptor that relays inhibitory signals to suppress the immune response. Alternative splicing results in multiple transcript variants. Polymorphisms in this gene have been associated with an increased risk of rheumatoid arthritis. 

```{r}
gene_report(res_list, "ENSG00000186265")
```

```{r}
plot_gene_beehive(dds, res_list$LRT[res_list$LRT$ensembl_gene_id=="ENSG00000186265",],
                  title_string = "LRT", intgroup = "PPBC", colorby = "batch")
plot_gene_beehive(dds, res_list$LRT[res_list$LRT$ensembl_gene_id=="ENSG00000186265",],
                  intgroup = "batch", colorby = "PPBC", color_vals = ppbc_colors)

```
 
## T cell markers {.tabset}

See [this overview](https://www.abcam.com/primary-antibodies/effector-t-cell-markers)

### [CD3G](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CD3G&keywords=CD3)

>Part of the TCR-CD3 complex present on T-lymphocyte cell surface that plays an essential role in adaptive immune response. When antigen presenting cells (APCs) activate T-cell receptor (TCR), TCR-mediated signals are transmitted across the cell membrane by the CD3 chains CD3D, CD3E, CD3G and CD3Z

```{r}
gene_report(res_list, "ENSG00000160654")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000160654",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [CD4](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CD4)

```{r}
gene_report(res_list, "ENSG00000010610")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000010610",],
                  title_string = "LRt", intgroup = "PPBC", color_vals = pam_colors)
```

### [CD8A](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CD8A&keywords=cd8)

```{r}
gene_report(res_list, "ENSG00000153563")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000153563",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [CD8B](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CD8B&keywords=cd8)

```{r}
gene_report(res_list, "ENSG00000172116")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000172116",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [CXCR4](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CXCR4&keywords=cxcr4)

```{r}
gene_report(res_list, "ENSG00000121966")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000121966",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [CD30](https://www.genecards.org/cgi-bin/carddisp.pl?gene=TNFRSF8&keywords=ENSG00000120949)

TNFRSF8 ENSG00000120949

>The protein encoded by this gene is a member of the TNF-receptor superfamily. This receptor is expressed by activated, but not by resting, T and B cells. TRAF2 and TRAF5 can interact with this receptor, and mediate the signal transduction that leads to the activation of NF-kappaB. This receptor is a positive regulator of apoptosis, and also has been shown to limit the proliferative potential of autoreactive CD8 effector T cells and protect the body against autoimmunity.

```{r}
gene_report(res_list, "ENSG00000120949")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000120949",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

## Macrophages {.tabset}

### [CD68](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CD68&keywords=cd68)

>CD68 (Cluster of Differentiation 68) is a protein highly expressed by cells in the monocyte lineage (e.g., monocytic phagocytes, osteoclasts), by circulating macrophages, and by tissue macrophages (e.g., Kupffer cells, microglia). This gene encodes a 110-kD transmembrane glycoprotein that is highly expressed by human monocytes and tissue macrophages. It is a member of the lysosomal/endosomal-associated membrane glycoprotein (LAMP) family. The protein primarily localizes to lysosomes and endosomes with a smaller fraction circulating to the cell surface.

Significant padj, but below the l2fc threshold.

```{r}
gene_report(res_list, "ENSG00000129226")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000129226",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [CD163](https://www.genecards.org/cgi-bin/carddisp.pl?gene=CD163&keywords=cd163)

>Compared to CD68 that is commonly used as a pan-macrophage marker (Holness & Simmons, 1993), CD163 is regarded as a specific monocyte/macrophage marker for M2 macrophages (Ambarus et al., 2012; Lau et al., 2004; Qian & Pollard, 2010). The presence of CD163+ macrophages was suggested to have a stronger association with less favorable clinicopathological features than CD68+ macrophages (Medrek, Ponten, Jirstrom, & Leandersson, 2012). 

See [this overview](https://www.sciencedirect.com/topics/immunology-and-microbiology/cd163)

```{r}
gene_report(res_list, "ENSG00000177575")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000177575",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

# Other genes of interest {.tabset}

## *[NFKB1](https://www.genecards.org/cgi-bin/carddisp.pl?gene=NFKB1&keywords=ENSG00000109320)

>NF-kappa-B is a pleiotropic transcription factor present in almost all cell types and is the endpoint of a series of signal transduction events that are initiated by a vast array of stimuli related to many biological processes such as inflammation, immunity, differentiation, cell growth, tumorigenesis and apoptosis.

Gene appears as though it should be more differentially expressed than it is.

```{r}
gene_report(res_list, "ENSG00000109320")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id == "ENSG00000109320",],
                  title_string = "Involution vs rest", intgroup = "PPBC", color_vals = pam_colors)
```

```{r}
plot_gene_beehive(dds, result_df = res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id == "ENSG00000109320",],
                  title_string = "Involution vs rest", intgroup = "inv_vs_rest", color_vals = pam_colors)
```


In particular, it looks as though this gene should be significant in inv vs rest. Did it maybe get replaced due to Cook's distance outliers?

```{r}
outliers_replaced_inv_vs_rest = mcols(dds)$replace %>% enframe("ensembl_gene_id", "outlier_replaced") %>% filter(outlier_replaced==T) %>%
  left_join(.,select(gx_annot, "ensembl_gene_id", "gene_name", "gene_type"), by = "ensembl_gene_id")

print(paste("Genes with outliers replaced by the trimmed mean: ", nrow(outliers_replaced_inv_vs_rest)))
```

Evidently not...

```{r}
outliers_replaced_inv_vs_rest %>% filter(ensembl_gene_id == "ENSG00000109320")
```

### [FOXP3](https://www.genecards.org/cgi-bin/carddisp.pl?gene=FOXP3&keywords=foxp3)

```{r}
gene_report(res_list,"ENSG00000049768")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000049768",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [Androgren receptor AR](https://www.genecards.org/cgi-bin/carddisp.pl?gene=AR&keywords=AR)

```{r}
gene_report(res_list,"ENSG00000049768")
```

```{r}
plot_gene_beehive(dds, res_list$inv_vs_rest[res_list$inv_vs_rest$ensembl_gene_id=="ENSG00000169083",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [NOS1](https://www.genecards.org/cgi-bin/carddisp.pl?gene=NOS1&keywords=nos1)

Nitric oxide synthase

```{r}
gene_report(res_list, ensembl_id = "ENSG00000089250")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000089250",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [HLA-DQA1](https://www.genecards.org/cgi-bin/carddisp.pl?gene=HLA-DQA1&keywords=HLA-DQA1)

>HLA-DQA1 belongs to the HLA class II alpha chain paralogues.

Reported as significant in inv vs prbc, but shows a bimodal distribution in most groups with several extreme outliers in each.

```{r}
gene_report(res_list, "ENSG00000228284")
```


```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000228284",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

### [IRS4](https://www.genecards.org/cgi-bin/carddisp.pl?gene=IRS4)

>Acts as an interface between multiple growth factor receptors possessing tyrosine kinase activity, such as insulin receptor, IGF1R and FGFR1, and a complex network of intracellular signaling molecules containing SH2 domains. Involved in the IGF1R mitogenic signaling pathway. Promotes the AKT1 signaling pathway and BAD phosphorylation during insulin stimulation without activation of RPS6KB1 or the inhibition of apoptosis. Interaction with GRB2 enhances insulin-stimulated mitogen-activated protein kinase activity.

Major Cooks outlier.

```{r}
gene_report(res_list, "ENSG00000133124")
```

```{r}
plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == "ENSG00000133124",],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors)
```

# Milk proteins vs IG genes

Ensure that we're not just looking at IG genes in breast milk.

```{r}
#res_list$LRT %>% filter(padj < 0.05 & abs(log2FoldChange) > 0.5) %>%
#  filter(str_detect(gene_name, "CSN"))
milk_genes <- c("CSN1S1", "CSN1S2AP", "CSN3", "LALBA")
IG_genes <- res_list$inv_vs_rest %>% filter(Type == "immunoglobulin") %>%
  filter(padj < 0.05 & abs(log2FoldChange) > 0.5) %>% pull(gene_name)

milk_IG_mat <- geneEx[rownames(geneEx) %in% c(milk_genes, IG_genes),]
milk_IG_mat
```

## Heatmap

Show the differentially expressed IG genes and the differentially expression milk genes in the involution vs rest comparison.

```{r, fig.height = 7, fig.width=10}
allgenes = readxl::read_excel(path = here("results", "diffex", "06_LRT_allgenes.xlsx"))
mat = scrime::rowScales(milk_IG_mat)
hlp = list(title="rowscaled vst")

#Ensure input is matrix
mat = as.matrix(mat)
  
#Heatmap annotation
ann_top = sd[,"PPBC", drop=F]
  
#Top column annotation
colTop <- HeatmapAnnotation(df=ann_top, which="col",
                            col = list(PPBC = ppbc_colors)
                            #annotation_legend_param = list(list(title = legend_title))
)

#Bottom column annotation
ann_bottom = sd[,"PAM50", drop=F]
colBottom <- HeatmapAnnotation(df=ann_bottom, which="col", col = list(PAM50 = pam_colors))
  
# Row annotation
#anno_rows = allgenes %>%
#    select(gene_name, Type) %>%
#    distinct() %>%
#    filter(!duplicated(gene_name)) %>%
#    column_to_rownames("gene_name")

anno_rows = data.frame(type = case_when(
  rownames(milk_IG_mat) %in% milk_genes ~ "milk protein",
  TRUE ~ "immunoglobulin"),
  row.names = rownames(milk_IG_mat))

#Essential that the order be the same!
anno_rows = anno_rows[match(rownames(mat),rownames(anno_rows)), ,drop=F]
  
rowAnno = HeatmapAnnotation(df=anno_rows, which="row",
                            col=list(type = c("milk protein" = "lightgray", "immunoglobulin" = "#E7298A"))
)
  
milk_hm = Heatmap(mat,
                  top_annotation = colTop,
                  bottom_annotation = colBottom,
                  left_annotation = rowAnno,
                  heatmap_legend_param = hlp,
                  show_row_names = T,
                  show_column_names = F,
                  cluster_rows = T,
                  row_names_gp = gpar(fontsize = 8),
                  column_names_gp = gpar(fontsize = 10),
                  column_title = "Milk genes vs IG genes")

milk_hm
```

```{r}
pdf(here("results", "diffex", "figs", "milk_vs_IG_genes.pdf"), width = 10, height = 7)

milk_hm

dev.off()
```

## Beehives

```{r}
ensembl_milk = gx_annot %>% filter(gene_name %in% milk_genes) %>% pull(ensembl_gene_id)
ensembl_milk = ensembl_milk[1:4] #Duplicate

for (m in ensembl_milk){
  #print(res_list$LRT[res_list$LRT$ensembl_gene_id == m,])
  plot_gene_beehive(dds, result_df = res_list$LRT[res_list$LRT$ensembl_gene_id == m,],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors) %>%
    print()
}

```

Potential todo? Check more milk proteins from [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3586783/)

>The proteins of human milk are divided into the whey and casein fractions or complexes, with each comprised by a remarkable array of specific proteins and peptides.14,15 The most abundant proteins are casein, α-lactalbumin, lactoferrin, secretory immunoglobulin IgA, lysozyme, and serum albumin.16,17 

# Session info

```{r}
sessionInfo()
```

