---
title: "Clustering within involution samples"
Author: Kat Moore
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    highlight: tango
    df_print: paged
---

```{r}
rm(list = ls())
```


```{r, include=F}
library(DESeq2)
library(apeglm)
library(ggrepel)
library(ComplexHeatmap)
library(openxlsx)
library(scrime)
#library(sva)
library(survival)
library(survminer)
library(RColorBrewer)
library(here)
library(tidyverse)

theme_set(theme_bw())
```

A k-means clustering strategy on DEG genes within the PPBC project.

# Setup and pre-processing {.tabset}

This data has been filtered on the basis of having a gene-wise non-zero count in at least a third of the samples. 

```{r}
dds = readRDS(here("data/Rds/08_dds_ovr_inv_vs_rest.Rds"))
```

We will be changing study_group to match the groups plotted, so save the original in its own column.

```{r}
table(dds$study_group)
dds$study_group_orig <- dds$study_group
```

```{r}
source(here("src", "deseq_report_functions.R"))
```

Subset for just the involution samples.

```{r}
inv.dds = dds[,colnames(dds)[dds$study_group == "ppbc_inv"]]
inv.dds$study_group = droplevels(inv.dds$study_group)
table(inv.dds$study_group)
```

## Gene metadata

```{r}
gx_annot <- read_tsv(here("data/metadata/01_tx_annot.tsv"))
gx_annot = gx_annot %>% select(ensembl_gene_id = gene_id, gene_name, gene_type, description = gene_description) %>% distinct()
```

## Variance stabilizing transformation

Renders the data homoskedastic, useful for visualization purposes.

Ensure that we can set "study group" to be one vs rest columns when desired.

```{r}
vsd = readRDS(here("data","Rds","08_vsd_ovr.Rds"))
```

## Sample annotation data

```{r}
sd = as.data.frame(colData(dds)) %>%
  select(sample_name, study_group, contains("rest"),
         PAM50, study_group_orig, everything())
head(sd)
```

## Symbol conversion and de-deuplication

```{r}
geneEx = rownames_to_column(as.data.frame(assay(vsd)), "ensembl_gene_id")
geneEx = right_join(select(gx_annot, gene_name, ensembl_gene_id),
                    geneEx, by = "ensembl_gene_id") %>%
    select(-ensembl_gene_id)
geneEx = summarize_expression_duplicate_ids(geneEx, id_column = "gene_name", verbose = T)

```

## Colors

```{r}
cp = readRDS(here("data", "Rds", "color_palettes.Rds"))
study_colors = cp$study_colors
pam_colors = cp$pam_colors
gene_colors = cp$gene_colors
ppbc_colors = cp$ppbc_colors

sp = readRDS(here("data/Rds/survival_colors.Rds"))
```

## Diffex results

Read in significant results from comparison involution vs non-prbc, and inv vs rest

```{r}
#openxlsx::getSheetNames(here("results", "diffex", "07_pairwise_comparisons_allgenes.xlsx"))
res_invp = openxlsx::read.xlsx(here("results", "diffex", "07_pairwise_comparisons_allgenes.xlsx"), sheet="rep_inv_nonprbc_all")
#head(res_invp)

#openxlsx::getSheetNames(here("results", "diffex", "07_pairwise_comparisons_sig_genes.xlsx"))
sig.invp = openxlsx::read.xlsx(here("results", "diffex", "07_pairwise_comparisons_sig_genes.xlsx"), sheet="sig_rep_inv_nonprbc")
#head(sig.invp)

#openxlsx::getSheetNames(here("results", "diffex", "08_one_vs_rest_allgenes.xlsx"))
res_invr = openxlsx::read.xlsx(here("results", "diffex", "08_one_vs_rest_allgenes.xlsx"), sheet="rep_inv_rest_all")
#head(res_invr)

#openxlsx::getSheetNames(here("results", "diffex", "08_one_vs_rest_sig_genes.xlsx"))
sig.invr = openxlsx::read.xlsx(here("results", "diffex", "08_one_vs_rest_sig_genes.xlsx"), sheet="sig_rep_inv_rest")
#head(sig.invp)
```

# Diffex genes

Thresholds applied:

```{r}
rbind(openxlsx::read.xlsx(here("results", "diffex", "07_pairwise_comparisons_sig_genes.xlsx"), sheet="thresholds") %>%
        filter(comparison == "rep_inv_nonprbc"),
      openxlsx::read.xlsx(here("results", "diffex", "08_one_vs_rest_sig_genes.xlsx"), sheet="thresholds") %>%
        filter(comparison == "rep_inv_rest"))
```

```{r, collapse=T}
print(paste("There are", nrow(sig.invp), "significant genes in the comparison involution vs non-prbc"))
print(paste("There are", nrow(sig.invr), "significant genes in the comparison involution vs rest"))
```

Top genes vs nulliparous

```{r}
sig.invr %>% head(30)
```

vs rest

```{r}
sig.invp %>% head(30)
```

# Two-way Heatmap

Depicted are genes significant in the comparison involution vs non-prbc using counts normalized via the variance stabilizing transformation.

## Dendrogram split

Inv vs np

```{r, include=F}
sd$study_group_orig %>% unique()
```

```{r, include=F}
geneEx[1:4,1:4]
```


```{r, fig.height=7, fig.width=10}
source(here("src", "deseq_report_functions.R"))

deseq_heatmap(mat = geneEx, sampledata = mutate(sd, study_group = study_group_orig),
              sig_results = sig.invp, title = "Involution vs nonprbc, Cutree split",
              groups_to_plot =c("ppbc_inv", "non_prbc"), maxn_rownames = Inf,
              column_split=3, row_split = 3)

```

```{r, fig.height=7, fig.width=10}
source(here("src", "deseq_report_functions.R"))

deseq_heatmap(mat = geneEx, sampledata = mutate(sd, study_group = study_group_orig),
              sig_results = sig.invr, title = "Involution vs rest, Cutree split",
              #groups_to_plot =c("ppbc_inv", "non_prbc"), 
              maxn_rownames = Inf,
              column_split=3, row_split = 3)

```

## K-means

```{r, fig.height=6, fig.width=10}
set.seed(1)

deseq_heatmap(mat = geneEx, sampledata = mutate(sd, study_group = study_group_orig),
              sig_results = sig.invp, title = "Involution vs nonprbc, K-means",
              groups_to_plot =c("ppbc_inv", "non_prbc"), maxn_rownames = Inf,
              column_km=3, row_km = 3)

```

```{r, fig.height=7, fig.width=10}
set.seed(1)

inv.hm = deseq_heatmap(mat = geneEx, sampledata = sd,#mutate(sd, study_group = inv_vs_rest),
              sig_results = sig.invr, title = "Significant DEGs from involution vs rest, K-means clustering",
              #groups_to_plot =c("ppbc_inv", "rest"),
              column_km=3, row_km = 3)

inv.hm
```

```{r}
resDir = file.path(here("results/clustering"))
dir.create(resDir, showWarnings = F)
pdf(file = file.path(resDir, "11_hm_clust_DEG_inv_vs_rest.pdf"))
inv.hm
dev.off()
```


### Summary column clusters

The column clusters can be split into a low (1), medium(2) and high(3) Ig-expressing groups

We can retrieve the gene/sample names either with row/column_order functions, which return indices:

```{r, fig.height = 7, fig.width= 10}
inv.hm = draw(inv.hm)
```


```{r, collapse=T}
#row_order(inv.hm)
#column_order(inv.hm) #Shows that `3` is high, `2` is medium and `1` is low
lapply(column_order(inv.hm), length)
#colnames(geneEx)[82] #Index of first sample in "high" group
#stopifnot(colnames(geneEx)[82] %in% as.hclust(coldend.hm$`3`)$labels) #Should be in the high group

col_clust = 
bind_rows(
  tibble(sample_name = colnames(geneEx)[column_order(inv.hm)$`1`],IG_col_cluster = "low", col_cluster=1),
  tibble(sample_name = colnames(geneEx)[column_order(inv.hm)$`2`],IG_col_cluster = "medium", col_cluster=2),
  tibble(sample_name = colnames(geneEx)[column_order(inv.hm)$`3`],IG_col_cluster = "high", col_cluster=3)
) %>% left_join(.,select(as.data.frame(colData(dds)), sample_name, study_group, PAM50), by="sample_name") %>%
  mutate(IG_col_cluster = factor(IG_col_cluster, levels = c("low", "medium", "high")))

addmargins(table(col_clust$study_group, col_clust$IG_col_cluster), 1)

nrow(col_clust)

#col_clust %>% arrange(sample_name)
```

...or by using the row/column_dend followed by as.hclust, after which the sample and gene names are accessible via the label operator.

Note that for this to work, you MUST create the heatmap by explicitly calling draw. If this this not done, the code will complete without errors, but the sample labels will be wrong!

```{r, eval=F}
stopifnot(class(inv.hm) == "HeatmapList") #Will be "HeatmapList" instead of Heatmap if draw has been called


test_clust =
bind_rows(
  tibble(sample_name = as.hclust(coldend.hm$`1`)$labels,IG_col_cluster = "low", col_cluster=1),
  tibble(sample_name = as.hclust(coldend.hm$`2`)$labels,IG_col_cluster = "medium", col_cluster=2),
  tibble(sample_name = as.hclust(coldend.hm$`3`)$labels,IG_col_cluster = "high", col_cluster=3)
) %>% left_join(.,select(as.data.frame(colData(dds)), sample_name, study_group, PAM50), by="sample_name") %>%
  mutate(IG_col_cluster = factor(IG_col_cluster, levels = c("low", "medium", "high")))

addmargins(table(test_clust$study_group, test_clust$IG_col_cluster), 1)

stopifnot(identical(arrange(col_clust, sample_name), arrange(test_clust, sample_name)))
```

Define colors for IG clusters

```{r}
IG_expression_colors = cp$tertile_colors
names(IG_expression_colors) = c("low", "medium", "high")
color_grid(IG_expression_colors)
```

### Bar plot IG clusters

```{r, fig.width=8}

study_ig = col_clust %>% ggplot(aes(x = IG_col_cluster, fill=study_group)) + geom_bar(position = "fill") +
  scale_fill_manual(values=study_colors) +
  ggtitle("Study group by IG expression cluster")


clust_ig = col_clust %>% ggplot(aes(fill = factor(IG_col_cluster, levels=c("high", "medium","low")), x=study_group)) +
  geom_bar(position = "fill") +
  #scale_fill_brewer(type="div", palette = 9, direction = 1) +
  scale_fill_manual(values = IG_expression_colors) +
  ggtitle("Division of IG expression clusters by study group") +
  labs(fill="IG_col_cluster")


study_ig
clust_ig

```


```{r}

pdf(file = file.path(resDir, "11_barplots_ig_clusters.pdf"),
    width = 8, height = 6)
study_ig
clust_ig
dev.off()

```

### Summary row clusters

Extract and store the row clusters:

```{r}
rowdend.hm = row_dend(inv.hm)

row_clust = 
bind_rows(
  tibble(gene_name = as.hclust(rowdend.hm$`1`)$labels,row_cluster = 1),
  tibble(gene_name = as.hclust(rowdend.hm$`2`)$labels,row_cluster = 2),
  tibble(gene_name = as.hclust(rowdend.hm$`3`)$labels,row_cluster = 3)
) %>%
  right_join(sig.invr,
             ., by = "gene_name") #%>% select(gene_name, immune_gene, row_cluster, description, everything())
```

```{r}
row_clust[row_clust$row_cluster==1,] %>% head()
row_clust[row_clust$row_cluster==2,] %>% head()
row_clust[row_clust$row_cluster==3,] %>% head()
```

```{r}
write.xlsx(list(row_clusters = row_clust, col_clusters = col_clust), file = file.path(resDir,"11_inv_clusters.xlsx"))
saveRDS(list("row_clust" = row_clust,
             "col_clust" = col_clust),
        here("data", "Rds", "11_ig_clusters.Rds"))
```

### Clusters as metagenes

Collapse clusters into single metagene by averaging experession, then test assocation between metagene and survival/subgroups

Remember that row cluster 3 is the IG genes.

```{r}
metagene_mat = 
  rbind(colMeans(geneEx[rownames(geneEx) %in% row_clust[row_clust$row_cluster==1,"gene_name"], ]),
        colMeans(geneEx[rownames(geneEx) %in% row_clust[row_clust$row_cluster==2,"gene_name"], ]),
        colMeans(geneEx[rownames(geneEx) %in% row_clust[row_clust$row_cluster==3,"gene_name"], ])
  )

rownames(metagene_mat) = c("metagene1", "metagene2","metageneIG")
metagene_mat = scrime::rowScales(metagene_mat)


#pheatmap::pheatmap(metagene_mat, show_colnames = F, cluster_rows = F,
#                   annotation_col = data.frame(row.names = sd$sample_name,
#                                               study_group = sd$inv_vs_rest))

  

#Top column annotation
colTop <- HeatmapAnnotation(df=data.frame(row.names = sd$sample_name,
                                          study_group = sd$PPBC),
                            which="col",
                            col = list(study_group = ppbc_colors),
                            annotation_legend_param = list(list(title = "PPBC")))

#Bottom column annotation

colBottom <- HeatmapAnnotation(df=data.frame(row.names = sd$sample_name,
                                             PAM50 = sd$PAM50),
                               which="col",
                               col = list(PAM50 = pam_colors))

set.seed(1)
Heatmap(metagene_mat,column_km = 3,
        top_annotation = colTop,
        bottom_annotation = colBottom,
        #left_annotation = rowAnno,
        heatmap_legend_param = list(title="rowscaled vst"),
        show_row_names = T,
        show_column_names = F,
        cluster_rows = F,
        #row_names_gp = gpar(fontsize = row_size),
        #column_names_gp = gpar(fontsize = col_size),
        #column_title = title
)
```
MetageneIG only
              
```{r, fig.width = 8}
set.seed(1)

Heatmap(metagene_mat[rownames(metagene_mat) == "metageneIG", , drop =F],column_km = 3,
        top_annotation = colTop,
        bottom_annotation = colBottom,
        #left_annotation = rowAnno,
        heatmap_legend_param = list(title="rowscaled vst"),
        show_row_names = T,
        show_column_names = F,
        cluster_rows = F,
        #row_names_gp = gpar(fontsize = row_size),
        #column_names_gp = gpar(fontsize = col_size),
        #column_title = title
)
```
              
#### KW

For inv vs rest

```{r}
stopifnot(all(colnames(metagene_mat) == sd$sample_name))

kruskal.test(x = metagene_mat[rownames(metagene_mat)== "metageneIG"], 
             g = sd$inv_vs_rest)
```


```{r}
kruskal.test(x = metagene_mat[rownames(metagene_mat)== "metageneIG"], 
             g = sd$study_group)
```


```{r}
kruskal.test(x = metagene_mat[rownames(metagene_mat)== "metageneIG"], 
             g = sd$PAM50)
```

```{r}
kruskal.test(x = sd$inv_vs_rest, 
             g = sd$PAM50)
```


# Clustering involution samples

We will attempt clustering both before and after removal of molecular subtype as a confounding factor. See if the samples also cluster based on immune high-medium-low categories.

```{r}
inv.dds$IG_cluster = col_clust[match(colData(inv.dds)$sample_name,
                                     col_clust$sample_name), "IG_col_cluster"][[1]]

sd$IG_cluster = col_clust[match(sd$sample_name,col_clust$sample_name), "IG_col_cluster"][[1]]

table(inv.dds$IG_cluster, inv.dds$study_group)
```


## Inv heatmaps

Subset down to inv and define heat map annotations

```{r}
#Subst to inv
sd.inv = as.data.frame(colData(inv.dds))
geneEx.inv = geneEx[,colnames(geneEx) %in% sd.inv$sample_name]
#head(select(sd.inv, sample_name, IG_cluster))


#Top column annotation
ann_top = select(sd.inv, IG_cluster, death = overall_survival)
#Ensure that the order matches
ann_top = ann_top[match(colnames(geneEx.inv),rownames(ann_top)), ,drop=F]
colTop <- HeatmapAnnotation(df=ann_top, which="col",
                              col = list(IG_cluster=IG_expression_colors,
                                         death = sp$overall_survival))

#Bottom column annotation
ann_bottom = sd.inv[,"PAM50", drop=F]
colBottom <- HeatmapAnnotation(df=ann_bottom, which="col",
                               col = list(PAM50 = pam_colors))

# Row annotation
anno_rows = res_invr %>%
    select(gene_name, Type) %>%
    distinct() %>%
    filter(!duplicated(gene_name)) %>%
    column_to_rownames("gene_name")

anno_rows = anno_rows[match(rownames(geneEx.inv),rownames(anno_rows)), ,drop=F]
anno_rows = anno_rows[rownames(anno_rows) %in% sig.invr$gene_name,,drop=F]

#Essential that the order be the same!
rowAnno = HeatmapAnnotation(df=anno_rows, which="row", col=list(Type = gene_colors))

```



### Significant genes in involution vs rest

```{r, fig.height=7, fig.width=10}
set.seed(1)

invclust =  Heatmap(as.matrix(geneEx.inv[rownames(geneEx.inv) %in% sig.invr$gene_name,]),
          top_annotation = colTop,
          bottom_annotation = colBottom,
          left_annotation = rowAnno,
          heatmap_legend_param = list(title="vst counts"),
          show_row_names = T,
          show_column_names = F,
          cluster_rows = T,
          row_names_gp = gpar(fontsize = 8),
          column_names_gp = gpar(fontsize = 8),
          column_title = "Clustering involution samples, sig genes",
          column_km = 3, column_km_repeats=50, row_km = 3, row_km_repeats = 50)

invclust.row =  Heatmap(scrime::rowScales(geneEx.inv[rownames(geneEx.inv) %in% sig.invr$gene_name,]) %>% as.matrix,
          top_annotation = colTop,
          bottom_annotation = colBottom,
          left_annotation = rowAnno,
          heatmap_legend_param = list(title="rowscaled vst"),
          show_row_names = T,
          show_column_names = F,
          cluster_rows = T,
          row_names_gp = gpar(fontsize = 8),
          column_names_gp = gpar(fontsize = 8),
          column_title = "Clustering involution samples, sig genes",
          column_km = 3, column_km_repeats=50, row_km = 3, row_km_repeats = 50)

invclust
invclust.row

```

### Involution: Top variant genes

Look for patterns within the involution group taking the top variant genes.

```{r}
rv = rowVars(as.matrix(geneEx.inv))
select = order(rv, decreasing = TRUE)[seq_len(min(500, 
            length(rv)))]
varEx.vst = geneEx.inv[select,]
vargenes = rownames(varEx.vst)
dim(varEx.vst)
```

```{r, fig.width=10}
set.seed(1)

# Row annotation
anno_rows = res_invr %>%
    select(gene_name, Type) %>%
    distinct() %>%
    filter(!duplicated(gene_name)) %>%
    column_to_rownames("gene_name")

anno_rows = anno_rows[match(rownames(varEx.vst),rownames(anno_rows)), ,drop=F]
#anno_rows = anno_rows[rownames(anno_rows) %in% sig.invr$gene_name,,drop=F]

#Essential that the order be the same!
rowAnno = HeatmapAnnotation(df=anno_rows, which="row", col=list(Type = gene_colors))
inv.var =  
Heatmap(as.matrix(varEx.vst),
          top_annotation = colTop,
          bottom_annotation = colBottom,
          left_annotation = rowAnno,
          heatmap_legend_param = list(title="vst counts"),
          show_row_names = F,
          show_column_names = F,
          cluster_rows = T,
          row_names_gp = gpar(fontsize = 8),
          column_names_gp = gpar(fontsize = 8),
          column_title = "Clustering involution samples, top 500 most variant genes vst",
          column_km = 3, column_km_repeats=50, row_km = 3, row_km_repeats = 50)

inv.var.row =  
Heatmap(as.matrix(scrime::rowScales(varEx.vst)),
          top_annotation = colTop,
          bottom_annotation = colBottom,
          left_annotation = rowAnno,
          heatmap_legend_param = list(title="vst counts"),
          show_row_names = F,
          show_column_names = F,
          cluster_rows = T,
          row_names_gp = gpar(fontsize = 8),
          column_names_gp = gpar(fontsize = 8),
          column_title = "Clustering involution samples, top 500 most variant genes vst rowscaled",
          column_km = 3, column_km_repeats=50, row_km = 3, row_km_repeats = 50)

inv.var
inv.var.row
```

We just see PAM50...

```{r}
pdf(file.path(resDir, "11_hm_involution_only_heatmaps.pdf"),width = 10, height = 7)
print(invclust)
print(invclust.row)
print(inv.var)
print(inv.var.row)
dev.off()
```

All samples

```{r}
rv = rowVars(as.matrix(geneEx))
select = order(rv, decreasing = TRUE)[seq_len(min(500, 
            length(rv)))]
varEx.vst = geneEx[select,]
vargenes = rownames(varEx.vst)
dim(varEx.vst)
```

## Fibrinogen and IG clusters

Do low B cell patients have a high fibrinogen signature (FGB/FGG/FGFBP1)? 
Could the fibrinogen be suppressing B cell activity, resulting in increased mortality?

```{r}
as.data.frame(t(geneEx.inv[rownames(geneEx.inv) %in% c("FGB", "FGG", "FGFBP1"),])) %>%
  rownames_to_column("sample_name") %>%
  gather(key = "gene", value = "vst", -sample_name) %>%
  left_join(., select(col_clust, sample_name, IG_col_cluster), by = "sample_name") %>%
  left_join(., select(sd, sample_name, death = overall_survival),
                                  by ="sample_name") %>%
  ggplot(aes(x = IG_col_cluster, y = vst, color = as.factor(death))) +
  geom_point() + 
  facet_wrap(~ gene)
```


```{r, fig.width = 8}
#nrow(col_clust)
#fbg_mat <- geneEx[rownames(geneEx) %in% c("FGB", "FGG", "FGFBP1"),]
fbg_mat <- geneEx.inv[rownames(geneEx.inv) %in% c("FGB", "FGG", "FGFBP1"),]
ann_df <- col_clust %>% left_join(., select(sd, sample_name, death = overall_survival),
                                  by ="sample_name") %>%
  mutate(death = as.character(death)) %>%
  column_to_rownames("sample_name") %>% select(IG_col_cluster, death)
pheatmap::pheatmap(fbg_mat,
                   show_colnames = F,
                   annotation_col = ann_df,
                   annotation_colors = list(#study_group = study_colors[names(study_colors) != "rest"],
                                            death = sp$overall_survival,
                                            IG_col_cluster = IG_expression_colors),
                   scale = "row")
#geneEx[rownames(geneEx) %in% c(ig_gene_clust$gene_name, "FGB", "FGG", "FGFBP1"),]

#col_clust %>% arrange(sample_name)
```

```{r, fig.width=8}
#The third row cluster is the IG genes
ig_gene_clust = row_clust[row_clust$row_cluster==3,]
#ig_gene_clust$Type %>% table()
#fbg_mat <- geneEx[rownames(geneEx) %in% c(ig_gene_clust$gene_name,"FGB", "FGG", "FGFBP1"),]
fbg_mat <- geneEx.inv[rownames(geneEx.inv) %in% c(ig_gene_clust$gene_name,"FGB", "FGG", "FGFBP1"),]
ann_df <- col_clust %>% left_join(., select(sd, sample_name, death = overall_survival),
                                  by ="sample_name") %>%
  mutate(death = as.character(death)) %>%
  column_to_rownames("sample_name") %>% select(IG_col_cluster, death)
pheatmap::pheatmap(fbg_mat,
                   show_colnames = F,
                   annotation_col = ann_df,
                   annotation_colors = list(death = sp$overall_survival,
                     #study_group = study_colors[names(study_colors) != "rest"],
                                            PAM50 = pam_colors,
                                            IG_col_cluster = IG_expression_colors),
                   scale = "row", cluster_rows = F)
```

# Survival analysis IG clusters

```{r}
surv <- readRDS(here("data/Rds/04_survdata.Rds"))

surv <- right_join(surv,
           select(col_clust, sample_name, IG = IG_col_cluster),
           by = "sample_name") %>%
  filter(!is.na(patient_ref))

saveRDS(surv, here("data/Rds/11_ig_survdata.Rds"))
```

## Overall survival

Baseline: Study group

```{r}
survdiff(survival::Surv(time =months_overall_survival, event = overall_survival) ~ study_group, data=surv)
```

### KM IG clusters

```{r}
ig_colors = cp$tertile_colors
names(ig_colors) = paste0("IG=", names(ig_colors))
ggsurvplot( 
    fit = survfit(Surv(time = months_overall_survival, event=overall_survival) ~ IG, data = surv), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    title = "Overall survival by IG cluster",
    palette = ig_colors,
    pval = T,
    ggtheme = theme_bw())
```

### KM IG clusters, facet by PAM50

Basal and Her2 are spread among them, but Lum A and B are almost universally low IG groups.

```{r}
table(sd$PAM50, sd$IG_cluster)
```

```{r, fig.width=10}

  ggsurvplot(
    fit = survfit(Surv(time=months_overall_survival, event=overall_survival) ~ PAM50 + IG, data = surv), 
    xlab = "Months", 
    ylab = "Overall survival probability",
    title = "IG clusters faceted by PAM50",
    palette = cp$tertile_colors,
    pval = T,
    ggtheme = theme_bw(), 
    facet.by = "PAM50")


```

### KM IG clusters, facet by study group

```{r, fig.width=8}
sg_ig <-  ggsurvplot(
    fit = survfit(Surv(time=months_overall_survival, event=overall_survival) ~ PPBC + IG, data = surv), 
    xlab = "Months", 
    ylab = "Overall survival probability",
    title = "IG cluster by study group",
    pval = T,
    palette = cp$tertile_colors,
    ggtheme = theme_bw(),
    facet.by = "PPBC")

sg_ig
```


```{r}
ggsave(file = here("results", "survival", "11_IG_clusters_by_study_group.pdf"), print(sg_ig), width = 8, height=6)
```

### KM study group, faceted by IG cluster

```{r, fig.width=8}
ggsurvplot(
    fit = survfit(Surv(time=months_overall_survival, event=overall_survival) ~ PPBC + IG, data = surv), 
    xlab = "Months", 
    ylab = "Overall survival probability",
    title = "Study group by IG cluster",
    pval = T, 
    facet.by = "IG",
    palette = ppbc_colors,
    ggtheme = theme_bw())


```

### Multivariate Cox

```{r}
ig_cox = 
coxph(
  Surv(time=months_overall_survival,
       event=overall_survival) ~ age_at_diagnosis + grade + stage + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + strata(PAM50) + IG + study_group,
  data = surv)

summary(ig_cox)
```

#### Balanced curves IG clusters

Here we have to avoid using strata for both HT and PAM50 or we end up with groups that are too small to balance.
Since ER status is the main factor here, use strata for HT.

```{r}
coxph(
  Surv(time=months_overall_survival,
       event=overall_survival) ~ age_at_diagnosis + grade + stage + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + PAM50 + IG + study_group,
  data = surv) %>%

ggadjustedcurves(., variable = "IG", palette = cp$tertile_colors,
                 ggtheme = theme_bw(),
                 data = as.data.frame(surv), method="conditional") +
  ggtitle("IG cluster survival curve adjusted for clinical covariates: Balanced")

```

#### Stratified curves IG clusters

Applying stratification to IG cluster shows a curve balanced for everything BUT IG cluster.

```{r}
coxph(
  Surv(time=months_overall_survival,
       event=overall_survival) ~ age_at_diagnosis + year_of_diagnosis + grade + stage + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + PAM50 + strata(IG) + study_group,
  data = surv) %>%
  ggadjustedcurves(., variable = "IG", data = as.data.frame(surv), method="conditional",
                   palette = cp$tertile_colors,
                 ggtheme = theme_bw()) +
  ggtitle("IG cluster survival curve adjusted for clinical covariates: Stratifed")

```

## Distant recurrence free survival

### Univariate

```{r}

survdiff(survival::Surv(time =months_to_drs, event = distant_recurrence) ~ IG, data=surv)

```

### KM DRS: IG clusters

```{r}
ig_colors = cp$tertile_colors
names(ig_colors) = paste0("IG=", names(ig_colors))

ggsurvplot(
    fit = survfit(Surv(time=months_to_drs, event=distant_recurrence) ~ IG, data = surv), 
    xlab = "Months", 
    ylab = "DRS",
    title= "IG clusters: DRS (univariate)",
    pval = T,
    palette = ig_colors,
    ggtheme = theme_bw())
```

### KM DRS IG clusters, facet by study group

```{r, fig.width=10}
ggsurvplot(
    fit = survfit(Surv(months_to_drs, event=distant_recurrence) ~ PPBC + IG, data = surv), 
    xlab = "Months", 
    ylab = "DRS probability",
    title = "DRS - IG clusters by PPBC",
    pval = T,
    facet.by = "PPBC",
    palette = cp$tertile_colors,
    ggtheme = theme_bw()
)

```


### KM DRS study group, faceted by IG cluster

```{r, fig.width=10}
ggsurvplot(
    fit = survfit(Surv(time=months_to_drs, event=distant_recurrence) ~ PPBC + IG, data = surv), 
    xlab = "Months", 
    ylab = "DRS probability",
    title = "DRS - PPBC by IG cluster",
    pval = T, facet.by = "IG",
    palette = ppbc_colors,
    ggtheme = theme_bw())


```

### Multivariate Cox DRS

```{r}
ig_drs = coxph(
  Surv(time=months_to_drs,
       event=distant_recurrence) ~ age_at_diagnosis + year_of_diagnosis + grade + stage + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + strata(PAM50) + IG + PPBC,
  data = surv)

summary(ig_drs)
```

#### DRS Balanced curves IG clusters

As before, if we stratify both HT and PAM50, the groups become too small.
We stratify HT as it best encodes ER status.

```{r}
coxph(
  Surv(time=months_to_drs,
       event=distant_recurrence) ~ age_at_diagnosis + year_of_diagnosis + grade + stage + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + PAM50 + IG + PPBC,
  data = surv) %>%
ggadjustedcurves(., variable = "IG", data = as.data.frame(surv), method="conditional",     
                 palette = cp$tertile_colors,
    ggtheme = theme_bw()) +
  ggtitle("IG cluster DRS curve adjusted for clinical covariates: Balanced")

```

#### DRS Stratified curves IG clusters

```{r}
coxph(
  Surv(time=months_to_drs,
       event=distant_recurrence) ~ age_at_diagnosis + year_of_diagnosis + grade + stage + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + PAM50 + strata(IG) + PPBC,
  data = surv) %>%
  ggadjustedcurves(., variable = "IG", data = as.data.frame(surv), method="conditional", 
                   palette = cp$tertile_colors,
    ggtheme = theme_bw()) +
  ggtitle("IG cluster DRS curve adjusted for clinical covariates: Stratifed")

```

# Session info

```{r}
sessionInfo()
```

```{r}
save.image(here("reports", "11_clustering_involution.RData"))
```

