---
title: "Involution:gene CoxPH models"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
   html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    highlight: tango
    df_print: paged
params:
  survival_results:  
    value: "/DATA/share/postpartumbc/data/Rds/13_multi_interaction_drs.Rds"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=F}
library(DESeq2)
library(survival)
library(survminer)
library(pheatmap)
library(RColorBrewer)
library(openxlsx)
library(here)
library(ggpubr)
library(tidyverse)

theme_set(theme_bw())
```


# Load data {.tabset}

Auto detection of analysis type:
  
```{r test}
params$survival_results
```
Results from copxh analysis:  

```{r survres}
res <- readRDS(params$survival_results)
```
  
Detect basefile:

```{r resstring}
s = params$survival_results
s = str_remove_all(s, "/DATA/share/postpartumbc/data/Rds/")
print(s)
```

Analysis type:

```{r}
call = if_else(str_detect(params$survival_results, "multi"),
               "Multivariate", "Univariate")

event = if_else(str_detect(params$survival_results, "_drs."),
               "distant recurrence","overall survival")

input = if_else(str_detect(params$survival_results, "inv"),
                "involution only", "all samples")

coef = if_else(str_detect(params$survival_results, "interaction"),
                  "inv:gene", "genewise")
analysis_type = paste(call, event, coef, paste0("(", input ,")"))
print(paste(s, ":", analysis_type))
```

Gene annotation:

```{r}
gx_annot <- read_tsv(here("data/metadata/01_tx_annot.tsv"))
gx_annot = gx_annot %>% select(ensembl_gene_id = gene_id, gene_name,
                               gene_type, description = gene_description) %>% distinct()
```

For pathway analysis:

```{r}
source(here("src/enrichment-analysis-functions.R"))
```

## Gene sets

For Fisher's exact tests

```{r}
gene_sets = list(
  go_bp = flexgsea::read_gmt(here("data","external","gmt","c5.bp.v7.0.symbols.gmt")),
  hallmark = flexgsea::read_gmt(here("data","external","gmt","h.all.v7.0.symbols.gmt")),
  c2_canon = flexgsea::read_gmt(here("data","external","gmt","c2.cp.v7.0.symbols.gmt")),
  c2_cgp = flexgsea::read_gmt(here("data","external","gmt","c2.cgp.v7.0.symbols.gmt"))
)
```

## Colors

```{r}
cp = readRDS(here("data", "Rds", "color_palettes.Rds"))
study_colors = cp$study_colors
pam_colors = cp$pam_colors
gene_colors = cp$gene_colors
ppbc_colors = cp$ppbc_colors

sp = readRDS(here("data/Rds/survival_colors.Rds"))
```

## Coxdata

A sample x feature matrix, where the first columns are clinical covariates and the rest are genes.

```{r}
input
```

```{r}
coxfile = if_else(input == "all samples",
          here("data/Rds/12_coxdata.Rds"),
          here("data/Rds/12_invdata.Rds"))

coxdata = readRDS(coxfile)

colnames(coxdata)[1:22]
```

## Gene expression matrix

Has been TMM normalized/log2 transformed.

```{r}
geneEx = select(coxdata, ENSG00000000003:ncol(coxdata))
geneEx = t(geneEx)
geneEx[1:4,1:4]
```

Convert IDs to symbols and average duplicates.

```{r}
source(here("src", "general_R_tools.R"))
geneEx = rownames_to_column(as.data.frame(geneEx), "ensembl_gene_id")
geneEx = right_join(select(gx_annot, gene_name, ensembl_gene_id),
                    geneEx, by = "ensembl_gene_id") %>%
  select(-ensembl_gene_id)
geneEx = summarize_expression_duplicate_ids(geneEx, id_column = "gene_name")
```

Format for pheatmap:

```{r}
rownames(geneEx) = NULL
geneEx = column_to_rownames(geneEx, "GeneSymbol")
geneEx = as.matrix(geneEx)
geneEx[1:4,1:4]
```


## Diffex results

```{r}
#Omit for now
#lrt <- here("results", "diffex", "06_LRT_allgenes.xlsx") %>% 
#  readxl::excel_sheets() %>% 
#  set_names() %>% 
#  map(readxl::read_excel, path = here("results", "diffex", "06_LRT_allgenes.xlsx"))
#names(lrt)

pw <- 
here("results", "diffex", "07_pairwise_comparisons_allgenes.xlsx") %>% 
  readxl::excel_sheets() %>% 
  purrr::set_names() %>% 
  purrr::map(readxl::read_excel, path = here("results", "diffex", "07_pairwise_comparisons_allgenes.xlsx"))
#names(pw)

ovr = here("results", "diffex", "08_one_vs_rest_allgenes.xlsx") %>% 
  readxl::excel_sheets() %>% 
  purrr::set_names() %>% 
  purrr::map(readxl::read_excel, path = here("results", "diffex", "08_one_vs_rest_allgenes.xlsx"))
#names(ovr)

diffex_list = append(pw, ovr)
names(diffex_list) = str_remove_all(str_remove_all(names(diffex_list), "rep_"), "_all")
names(diffex_list) = str_replace(names(diffex_list), "_", "_vs_")
#names(res_list)

diffex_list = bind_rows(diffex_list, .id="comparison")

basemeans = ovr$rep_inv_rest_all %>%
  select(ensembl_gene_id, baseMean, type=Type) %>%
  arrange(ensembl_gene_id) %>% distinct() %>%
  mutate(baseMean = round(baseMean, 2))

stopifnot(nrow(filter(basemeans, duplicated(ensembl_gene_id))$ensembl_gene_id)==0)
```

## Survival results

```{r}
head(res)
```

Add average expression, condensed biotypes...

```{r}
res = 
res %>% left_join(., basemeans, by="ensembl_gene_id") %>%
  select(gene_name:p.value, baseMean, type, everything())

```

...and selection involution padj and l2fcs to results:

```{r}
res = res %>% left_join(
  ., select(filter(diffex_list, comparison == "inv_vs_nonprbc" ),
       ensembl_gene_id, padj.inv_vs_nonprbc = padj,
       l2fc.inv_vs_nonprbc = log2FoldChange),
  by = "ensembl_gene_id"
) %>% left_join(
  ., select(filter(diffex_list, comparison == "inv_vs_rest" ),
       ensembl_gene_id, padj.inv_vs_rest = padj,
       l2fc.inv_vs_rest = log2FoldChange),
  by = "ensembl_gene_id"
) %>%
  select(gene_name:type, padj.inv_vs_rest, padj.inv_vs_nonprbc,
         l2fc.inv_vs_rest, l2fc.inv_vs_nonprbc, everything())

res %>% head()
```

# Overview

Example formula for coxph (one per gene):

```{r}
df_formulas = res %>% select(gene_name, ensembl_gene_id, call)
as.character(df_formulas$call[[1]])
res = res %>% select(-call)
```

```{r}
df_thresh = tibble(cutoff = seq(0.05, 0.3, by =0.05),
       hits = sapply(seq(0.05, 0.3, by =0.05), function(x){nrow(res[res$fdr <= x,])})
)

for(i in 1:nrow(df_thresh)){
  print(paste0(
    "Genes with fdr <= ", df_thresh$cutoff[i], ": ", df_thresh$hits[i]
  ))
}
  
```

Threshold selection.
Pick the minimum threshold that has at least 50 hits.

```{r}
pthresh = df_thresh %>% filter(hits >= 50) %>% dplyr::slice(1) %>% pull(cutoff)
if (length(pthresh) == 0){
  pthresh <- 0.3
}
print(paste("Fdr threshold:", pthresh))
```

# Top genes

Top 30 most significant results.

```{r}
res = res %>% arrange(fdr, p.value)

head(res, 30)
```

Protein coding genes below the threshold:

```{r}
res %>% filter(type == "protein coding" & fdr <= !!pthresh) %>%
  mutate(description = str_remove(description, "\\[.*\\]")) %>%
  mutate(hit = paste(gene_name, description, sep=": ")) %>% pull(hit)
```

# Deleterious genes

Beta > 0 : Higher gene expression = Likelier death

```{r, collapse=T}
print(paste0("Genes with beta > 0, fdr <= ", pthresh, ": ",
      nrow(res %>% filter(beta > 0 & fdr <= !!pthresh))))

```

```{r}
res %>% filter(beta > 0 & fdr <= !!pthresh) %>%
  mutate(description = str_remove(description, "\\[.*\\]")) %>%
  mutate(hit = paste0(gene_name, ", ", "fdr: ", signif(fdr, 2),
                      " beta: ", beta, ", ", description)) %>% pull(hit)
```

# Protective genes

Beta < 0 : Lower gene expression = Likelier death

```{r, collapse=T}
print(paste0("Genes with beta < 0, fdr <= ", pthresh, ": ",
      nrow(res %>% filter(beta < 0 & fdr <= !!pthresh))))

```

```{r}
res %>% filter(beta < 0 & fdr <= !!pthresh) %>%
  mutate(description = str_remove(description, "\\[.*\\]")) %>%
  mutate(hit = paste0(gene_name, ", ", "fdr: ", signif(fdr, 2),
                      " beta: ", beta, ", ", description)) %>%
  pull(hit) %>% head(50)
```

# Pathway overview: Fisher's exact {.tabset}

```{r}
source(here("src/enrichment-analysis-functions.R"))
paths = fisher_pathways(
  sig_genes = pull(filter(res, fdr < !!pthresh),gene_name),
      background_genes = res$gene_name,
      list_signatures = gene_sets,
      fdr_cutoff = 0.1,
      verbose = F,
      collapse_rows=F)

print("Sig pathways for all genes associated with survival, fdr < 0.1: ")

lapply(paths, nrow)
```

Top 20 pathways:

```{r}
bind_rows(paths, .id = "gene_set") %>%
  arrange(fdr, pval, fold_enrichment) %>% 
  select(- gene_set) %>%
  select(pathway, fdr, pval, fold_enrichment,
         members_in_sample,
         path_in_sample, path_in_pop, everything()) %>%
  head(20)

```

## Pathways for deleterious genes

```{r}
delpaths = fisher_pathways(
  sig_genes = pull(filter(res, fdr < !!pthresh & beta > 0), gene_name),
      background_genes = res$gene_name,
      list_signatures = gene_sets,
      fdr_cutoff = 0.1,
      verbose = F,
      collapse_rows=F)

print("Sig pathways fdr < 0.1: ")

lapply(delpaths, nrow)
```

Show the top 20 paths (if available)

```{r}
bind_rows(delpaths, .id = "gene_set") %>%
  arrange(fdr, pval, fold_enrichment) %>% 
  select( - gene_set) %>%
  select(pathway, fdr, pval, fold_enrichment,
         members_in_sample,
         path_in_sample, path_in_pop, everything()) %>%
  head(20)

```

## Pathways for protective genes

```{r}
goodpaths = fisher_pathways(
  sig_genes = pull(filter(res, fdr < !!pthresh & beta < 0), gene_name),
      background_genes = res$gene_name,
      list_signatures = gene_sets,
      fdr_cutoff = 0.1,
      verbose = F,
      collapse_rows=F)

print("Sig pathways fdr < 0.1: ")

lapply(goodpaths, nrow)
```

Show the top 20 paths (if available)

```{r}
bind_rows(goodpaths, .id = "gene_set") %>%
  arrange(fdr, pval, fold_enrichment) %>% 
  select( - gene_set) %>%
  select(pathway, fdr, pval, fold_enrichment,
         members_in_sample,
         path_in_sample, path_in_pop, everything()) %>%
  head(20)

```

# Heatmap

Using a quantile-based color scale to prevent outliers from bleaching the heatmaps.

```{r, fig.height=8, fig.width=7}
#Define column annotation
ann_df = coxdata[, c("sample_name","PPBC", "overall_survival")]
colnames(ann_df)[3] = "death"
ann_df = mutate(ann_df, death = as.factor(death))
ann_df = column_to_rownames(ann_df, "sample_name")

#Sort by survival
ann_df = ann_df[order(ann_df$death),]

#Subset geneEx to genes below threshold
if(nrow(filter(res, fdr <= !!pthresh)) < 5){
  mat = geneEx[rownames(geneEx) %in% head(res, 10)$gene_name, , drop=F]
  ttext = paste0("Top 10:\n ", analysis_type)
} else {
 mat = geneEx[rownames(geneEx) %in% filter(res, fdr <= !!pthresh)$gene_name, , drop=F] 
 ttext = paste0("Genewise fdr <", pthresh, ":\n ", analysis_type)
}
#dim(mat)

#Match matrix columns to annotation
mat = mat[,colnames(mat)[match(rownames(ann_df), colnames(mat))]]

#sp$overall_survival #white is too light
death_cols = sp$overall_survival
death_cols[2] = "lightgray"

#Quantile normalization
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}

mat_breaks <- quantile_breaks(mat, n = 11)



pheatmap::pheatmap(mat, show_colnames = F,
                   scale="row", annotation_col = ann_df,
         cluster_cols = T, drop_levels = T,
         breaks =  mat_breaks,
         color = colorRampPalette(rev(brewer.pal(n = 8,
                                      name = "RdYlBu")))(length(mat_breaks) - 1),
         annotation_colors = list(PPBC = ppbc_colors[-5],
                                  death = death_cols), #sp$overall_survival white is too light
         main = ttext
         ) 
```

Alternative visualization without column clustering

```{r, fig.height=8, fig.width=7}
pheatmap::pheatmap(mat, show_colnames = F,
                   scale="row", annotation_col = ann_df,
         cluster_cols = F, drop_levels = T,
         breaks =  mat_breaks,
         color = colorRampPalette(rev(brewer.pal(n = 8,
                                      name = "RdYlBu")))(length(mat_breaks) - 1),
         annotation_colors = list(PPBC = ppbc_colors[-5],
                                  death = death_cols), #sp$overall_survival white is too light
         main = ttext
         ) 
```


# Overlap with diffex

Focusing exclusively on those genes significant in inv vs nonprbc or inv vs rest.

Are there any genes that meet this criteria?

```{r}
res %>% filter(fdr <= !!pthresh) %>%
  filter(padj.inv_vs_rest <= 0.05 | padj.inv_vs_nonprbc <= 0.05)
```

Show genes that are at least below 0.25 in both.
Note that there's no l2fc cutoff here.

```{r}
res %>%
  filter(fdr < 0.25 & padj.inv_vs_rest < 0.25 ) %>%
  ggplot(aes(x = fdr, y = padj.inv_vs_rest,
             color = type)) +
  geom_point() +
  scale_color_manual(values = cp$gene_colors) +
  ggrepel::geom_label_repel(aes(label = gene_name),
                            show.legend = F) +
  ggtitle(paste("Genes with both cox fdr and inv vs rest padj < 0.25")) +
  xlab("cox fdr") +
  ylab("padj inv vs rest")

res %>%
  filter(fdr < 0.25 & padj.inv_vs_nonprbc < 0.25 ) %>%
  ggplot(aes(x = fdr, y = padj.inv_vs_nonprbc,
             color = type)) +
  geom_point() +
  scale_color_manual(values = cp$gene_colors) +
  ggrepel::geom_label_repel(aes(label = gene_name),
                            show.legend = F) +
  ggtitle(paste("Genes with both cox fdr and inv vs nonprbc padj < 0.25")) +
  xlab("cox fdr") +
  ylab("padj inv vs nonprbc")

```

# Session info

Also save data

```{r}
res = res %>%
  mutate(analysis = analysis_type)

write_csv(res,
          path = file.path("/DATA/share/postpartumbc/results/survival",
                           str_replace(
                             str_remove_all(params$survival_results, ".*\\/"),
                             "Rds", "csv")
                           )
          )
```

```{r, echo=F}
sessionInfo()
```


