---
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    fig_width: 7
    fig_height: 6
    df_print: paged
params:
  gene:
    value: "NOTCH4"
    label: "Gene name: "
    input: "text"
  show_functions:
    value: FALSE
    input: checkbox
    label: "Show function definitions:"
  show_code:
    value: TRUE
    label: "Show code:"
    input: checkbox  
title: "`r paste0('PPBC report for ', params$gene)`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = params$show_code)
library(DESeq2)
library(survival)
library(ggthemes)
library(survminer)
library(RColorBrewer)
library(openxlsx)
library(here)
library(ggpubr)
library(tidyverse)
theme_set(theme_bw())
#options(width = 80) #Only when pdf, and does not always work
Gene = params$gene
```

This report will summarize the differential expression and survival analysis results for `r Gene`.
The data from the analysis pipeline was aggregated in notebook 16.

```{r load env}
load(here("data/Rds/16_gene_report_environment.RData"))
```

The function documentation section explains how the results were calculated.
To jump straight to results, skip to Gene Biology.


# Function documentation

Display function definition is `r params$show_function`.
Enable to show function code.

## Gene lookup

Will retrieve the various ID forms when provided with an ensembl gene ID or gene name (symbol).

```{r fun gene lookup, include=params$show_functions}
gene_lookup <- function(gene, dict=bx_annot){
  
  if(gene %in% dict$gene_name){
    gene_name = gene
    ensembl_gene_id = unique(dict[dict$gene_name == gene_name,]$ensembl_gene_id)
  } else if(gene %in% dict$ensembl_gene_id) {
    ensembl_gene_id = gene
    gene_name = unique(dict[dict$ensembl_gene_id == ensembl_gene_id,]$gene_name)
  } else {
    stop("Provided gene is not a valid ensembl id or gene name")
  }
  
  uni=unique(dict[dict$ensembl_gene_id == ensembl_gene_id,]$uniprot_id)
  ent=unique(dict[dict$ensembl_gene_id == ensembl_gene_id,]$entrez_id)
  
  list(gene_name=gene_name,
       ensembl_gene_id = ensembl_gene_id,
       uniprot_id=uni,
       entrez_id=ent)
}

gene_lookup("ENSG00000156738")
gene_lookup("MARVELD2")
#gene_lookup("Steve")
```

## Genewise survival results

Shows a summary of results for genewise survival models of that gene when provided with a gene name or ensembl id.
Can also show interaction results if specified in the argument.
If more than one ensembl id is found and return_first is F, will combine all ensembl ids into a single data frame.
Otherwise, only the results from the first ensembl id is shown.

```{r fun , include=params$show_functions}
genewise_survival_results <- function(id, df = gw_surv, return_first = F){
  
  i = gene_lookup(id)$ensembl_gene_id
  
  df = df %>%
    select(-type) %>%
    select(type = gene_type, everything()) %>%
    select(ensembl_gene_id, model:baseMean, type)
  
  if(length(i) == 1){
    return(arrange(filter(df, ensembl_gene_id == i), fdr))
  }
  
  if (length(i > 1) & return_first == T){
    i = i[1]
    warning(paste0("Multiple ensembl IDs found for ",
                   id, "returning first in series: ", i))
    return(arrange(filter(df, ensembl_gene_id == i), fdr))
  } 
  
  if (length(i > 1 & return_first == F)){
    lapply(i, function(x){
      df[df$ensembl_gene_id == x, ]
    }) %>% bind_rows() %>%
      select(ensembl_gene_id, everything()) %>%
      arrange(fdr)
  }
  
}

genewise_survival_results("MARVELD2", return_first = T)
genewise_survival_results("MARVELD2", return_first = F)
#genewise_survival_results("MS4A1", return_first = F)
#genewise_survival_results("Steve", return_first = F)

#Can also be use to extract interaction results:
genewise_survival_results("MARVELD2", df = int_surv, return_first = F)
```

Because the genewise models run on involution samples only yielded no fdr-signifcant results, only those run on all samples are included.

Available genewise models:

```{r}
#TODO fix the "analysis" column - incorrect regex in notebook 12
#gw_surv$analysis %>% unique()
paste(gw_surv$formula, gw_surv$outcome) %>% unique()
```

Available interaction models:

```{r}
paste(int_surv$formula, int_surv$outcome) %>% unique()
```

## Diffex report

Creates a data frame displaying differential expression results for a given gene.
If subgroup = T, will show the subgroup analysis instead.

```{r, include=params$show_functions}
#TODO: Option to return all results when multiple ensembl ids are found
diffex_report = function(id, subgroup = F, list_reports = res_list, subgroup_reports = sub_reslist,
                         pthresh = 0.05, abslogfcthresh = 0.5){
  
  #print(paste("Adjusted p value threshold:", pthresh, ", abs(log2FoldChange) threshold:", abslogfcthresh))
  
  ensembl_id = gene_lookup(id)$ensembl_gene_id
  
  if(length(ensembl_id) > 1){
    ensembl_id = ensembl_id[1]
    warning(paste0("Multiple ensembl IDs found for ",
                   id, "returning first in series: ", ensembl_id))
  }
  
  if(subgroup == T){
    lr = subgroup_reports
  } else {
    lr = list_reports  
    }
  

  comp = lapply(lr,
                function(x) x[x$ensembl_gene_id==ensembl_id,]) %>%
    bind_rows(.id = "comparison")
  
  comp = comp %>% mutate(
    padj = signif(padj, 2),
    log2FoldChange = signif(log2FoldChange, 2),
    baseMean = round(baseMean, 1),
    sig = (padj <= !!pthresh & abs(log2FoldChange) > !!abslogfcthresh)
  ) %>% arrange(padj)

  
  if (subgroup == T){
    comp = comp %>% select(comparison, subgroup, sig,
                           padj, log2FoldChange, baseMean,
                           gene_name, ensembl_gene_id, gene_type)
  } else {
    comp = comp %>% select(comparison, sig,
                           padj, log2FoldChange, baseMean,
                           gene_name, ensembl_gene_id, gene_type)
    
    #L2fc threshold not meaningful for LRT
    comp[comp$comparison == "LRT", "sig"] <- comp$padj[comp$comparison == "LRT"] <= 0.05
  }
  
  comp  
}
diffex_report("MS4A1", subgroup = F)
diffex_report("MARVELD2", subgroup = T)
```

Available diffex comparisons:

```{r}
names(res_list)
```

Available subgroup comparisons:

```{r}
sapply(names(sub_reslist),
       function(x) paste(x, unique(sub_reslist[[1]]$subgroup))
)
```


## Boxplot expression

Creates two boxplots for a given gene: TMM-normalized/log2 tranformed, and size factor-normalized.
The boxplots are faceted by outcome (death or metastasis).

```{r, include=params$show_functions}
box_plots <- function(id, gEx = geneEx, d=dds, s = sample_data){
  
  #stopifnot(norm_type %in% c("size factor", "log2 TMM"))
  
  ensembl_id <- gene_lookup(id)$ensembl_gene_id
  
  if(length(ensembl_id) > 1){
    ensembl_id = ensembl_id[1]
    warning(paste0("Multiple ensembl IDs found for ",
                   id, ", returning first in series: ", ensembl_id))
  }
  
  #TMM
  df <- gEx[,c("sample_name",  ensembl_id)]
    colnames(df)[2] <- "geneEx"
  
  df <- left_join(df, s, by = "sample_name") %>%
    mutate(id = id, ensembl_id = ensembl_id)
  
  #return(df)
  
  df = df %>%
  select(sample_name, id, ensembl_id, geneEx, PPBC, PAM50,
         death = overall_survival,
         metastasis = distant_recurrence) %>%
  gather(key = "outcome", value = "value", death, metastasis) %>%
    mutate(value = as.factor(value))
  
  tmm = df %>%
    ggplot(aes(x = PPBC, y = geneEx)) +
    geom_boxplot(alpha=0, notch=F) +
    geom_jitter(width = 0.1, aes(color = value)) +
    facet_wrap(~outcome) +
    scale_color_colorblind() +
    ylab(paste0("TMM-log2 normalized expression")) +
    ggtitle(id)
  
  #size factor
  df <- plotCounts(d, gene = ensembl_id,
                   intgroup=c("PPBC", "PAM50",
                              "overall_survival",
                              "distant_recurrence"),
                   returnData = T, normalized = T)
  
  colnames(df)[colnames(df)=="count"] <- "geneEx"
  df <- rownames_to_column(df, "sample_name") %>%
    mutate(id = id, ensembl_id = ensembl_id)

  df = df %>%
  select(sample_name, id, ensembl_id, geneEx, PPBC, PAM50,
         death = overall_survival,
         metastasis = distant_recurrence) 
  df = df %>%
  gather(key = "outcome", value = "value", death, metastasis) %>%
    mutate(value = as.factor(value)) %>%
    na.omit()
  #return(df)
  sf = df %>%
    ggplot(aes(x = PPBC, y = geneEx)) +
    geom_boxplot(alpha=0, notch=F) +
    geom_jitter(width = 0.1, aes(color = value)) +
    facet_wrap(~outcome) +
    scale_color_colorblind() +
    ylab(paste0("Size-factor normalized counts")) +
    ggtitle(id)

  list(tmm = tmm, sf = sf)
}

box_plots("MARVELD2")
```

## Gene quantile

Assigns each sample a quantile based on a given gene's expression in the dataset.

```{r, include=params$show_functions}
gene_ntile <- function(gene_id, gEx = geneEx, gene_dict = gx_annot,
                       n = 3){
  
  stopifnot(n %in% c(2,3,4))
  
  #TODO update gene_lookup this wya
  id <- gene_lookup(gene_id)$ensembl_gene_id
  name <- gene_lookup(gene_id)$gene_name
  

  labels = switch(as.character(n),
                  "2" = c("low", "high"),
                  "3" = c("low", "medium", "high"),
                  "4" = c("Q1", "Q2", "Q3", "Q4")
  )
  samples <- geneEx[,"sample_name"]
  g <- geneEx[, colnames(gEx) %in% id , drop = F]
  g <- as.matrix(g)
  rownames(g) <- samples
  #return(g)
  if(ncol(g) == 0){
    stop("Gene not found in provided count matrix")
  }
  
  if(ncol(g) > 1){
    warning("Multiple ensembl ids mapped to single gene symbol. Averaging expression.")
    gene <- as.matrix(rowMeans(g))
    #return(gene)
  }
  
  #rownames(gene) <- name #Not strictly necessary, useful for debugging
  
  if(length(id) > 1){
    id = paste(id, collapse = ";")
  }
  
  gene_df <- tibble::tibble(sample_name = rownames(g),
                            gene_symbol = name,
                            ensembl_gene_id = id,
                            tmm = g[,1],
                            ntile = dplyr::ntile(g[,1], n),
                            labels = factor(dplyr::ntile(g[,1], n), labels = labels)
  )
  
  #gene_df <- dplyr::mutate(gene_df, labels = as.character(labels)) #Or survminer will complain
  
  return(gene_df)
  

}

gene_ntile("MARVELD2")%>%head()
gene_ntile("MS4A1")%>%head()

```

## Involution vs rest Kaplan Meiers

Creates 4 separate kaplan meiers based on quantiles of a given gene, based on `survival_type`.
(Either overall survival or distant recurrence is allowed.)
The KMs will be shown for involution samples or for all non-involuting samples separately.
Adjusted curves are shown alongside unadjusted curves, using the *conditional* method of `survminer::surv_adjustedcurves`.
Adjusted curves are corrected for the following covariates:

```{r}
covariates <- c("age_at_diagnosis", "grade",
                "stage", "surgery", "radiotherapy",
                "strata(hormonetherapy)",
                "chemotherapy", "herceptin", "PAM50")
```

Ideally, both PAM50 and hormonetherapy would be stratified.
Because stratifying both PAM50 and HT yields too many NAs, only HT is stratifed to compensate for the PH violation of TN-tumors.
For unadjusted curves, the logrank p value is shown.
For adjusted curves, the p value is from an anova between a model that includes the gene quantile and a model without the gene quantile.

```{r, include=params$show_functions, fig.height=7, fig.width=10}
km_ntiles_ovr <- function(gene_id, gene_dict = gx_annot, gEX = geneEx,
                          n = 3, ovr_column = "involution",
                          line_colors = viridis::scale_color_viridis(discrete = T),
                          sampledata = sample_data, survival_type = "os",
                          legend_positions = c("none","none", "bottom","bottom"),
                          p_method = "anova", return_list = F){
  
  stopifnot(survival_type %in% c("os", "drs"))
  stopifnot(n %in% c(2,3,4))
  stopifnot(p_method %in% c("anova", "coef"))
  stopifnot(ovr_column %in% colnames(sampledata))
  
  ntiles <- gene_ntile(gene_id = gene_id, gene_dict = gene_dict,
                       gEx = gEX, n = n)
  
  sd <- dplyr::left_join(ntiles, sampledata, by = "sample_name")
  #  return(sd)
  
  #Toggle for survival type
  if(survival_type == "os"){
    ti <- "months_overall_survival"
    ev <- "overall_survival"
    ylab <- "Overall survival probability"
    type <- "Overall survival"
  }
  
  if(survival_type == "drs"){
    ti <- "months_to_drs"
    ev <- "distant_recurrence"
    ylab <- "Distant recurrence probability"
    type <- "Distant recurrence"
  }
  
  covariates <- c("age_at_diagnosis", "grade",
                  "stage", "surgery", "radiotherapy",
                  "strata(hormonetherapy)",
                  "chemotherapy", "herceptin", "PAM50")
  
  cov_legend <- c("age", "year of diagnosis", "grade",
                  "stage", "treatment", "PAM50")
  
  cov <- paste(covariates, collapse = "+")
  svf = paste0('Surv(time=',ti,', event=',ev,')~')
  facet_formula <- as.formula(paste0(svf, "labels"))
  #ntile_formula <- as.formula(paste0(svf, "ntile"))
  reduced =  as.formula(paste0(svf, cov))
  formula = as.formula(paste0(svf, paste0(cov, "+ntile")))
  
  reslist <- list()
  gn <- unique(sd$gene_symbol)
  
  #Split sample data into those samples belonging to the group of interest (or not)
  sd1 = sd[sd[[ovr_column]] == 1,]
  sd0 = sd[sd[[ovr_column]] == 0,]
  
  
  curv1 <- survminer::ggsurvplot(
    fit = survminer::surv_fit(facet_formula, data = as.data.frame(sd1)), 
    xlab = "Months", 
    ylab = ylab,
    legend = legend_positions[1],
    #title = paste(type, "(unadjusted), with", gn, "\nexpression in", ovr_column, "samples"),
    title = paste("Unadjusted curve for", ovr_column, "samples"),
    pval = T, #Logrank method
    )
  
  curv1 = curv1$plot #Otherwise we get a ggsurvplot object, which cannot be used with ggarrange
  
  curv1 <- curv1 +
    theme(plot.title = element_text(size = 12),
          axis.title.x=element_blank(),
          axis.title.y=element_blank()
    ) +
    line_colors
  
  reslist$curv1 <- curv1
  
  #Unadajusted kaplan meier plot in other samples
  curv0 <- survminer::ggsurvplot(
    fit = survminer::surv_fit(facet_formula, data = as.data.frame(sd0)), 
    xlab = "Months", 
    ylab = ylab,
    legend = legend_positions[2],
    #title = paste(type, "(unadjusted), with", gn, "\nexpression in other samples"),
    title = paste("Unadjusted curve for other samples"),
    pval = T, #Logrank method
    ) 
  
  curv0 <- curv0$plot #Otherwise we get a ggsurvplot object, which cannot be used with ggarrange
  
  curv0 <- curv0 +
    theme(plot.title = element_text(size = 12),
          axis.title.x=element_blank(),
          axis.title.y=element_blank()
    ) +
    line_colors
  
  reslist$curv0 <- curv0
  
  #return(reslist)
  
  #Cox models for involution group, adjusted for all covariates and ntile gene expression
  c <- survival::coxph(formula, data = sd1)
  #print(c)
  
  #Store p value associated with the coefficient of the ntile
  fit_ntile <- as.data.frame(summary(c)$coefficients)
  p_ntile <- signif(fit_ntile[rownames(fit_ntile) == "ntile", ]$`Pr(>|z|)`, 2)
  if (p_ntile < 0.001){
    p_ntile <- "< 0.001"
  }
  
  #print(p_ntile)
  #Anova p value vs reduced model
  d <- survival::coxph(reduced, data = as.data.frame(sd1))
  aod <- anova(d,c) #Conventional to list models from smallest to largest but order does not matter
  aod_p <- signif(aod[4][2,], 2)
  if (aod_p < 0.001){
    aod_p <- "< 0.001"
  }
  
  if (p_method == "anova"){
    p <- aod_p
  } else {
    p <- p_ntile
  }
  
  #Adjusted survival curve for samples belonging to group of interest
  #Use surv_adjustedcurves to get data and the then manually plot for more control over appearance
  adj_curv1 <- survminer::surv_adjustedcurves(fit = c, variable = "ntile", data = as.data.frame(sd1), method = "conditional") %>%
    left_join(select(mutate(sd1, ntile = as.factor(ntile)), ntile, labels), by = c("variable" = "ntile")) %>%
    ggplot(., aes(x = time, y = surv, color = labels)) + 
    geom_step(size = 1) +
    theme_survminer() +
    line_colors + 
    scale_y_continuous(limits = c(0,1)) +
    #scale_x_continuous(limits = c(0,250)) +
    ylab(label =  ylab) +
    labs(color = paste(gn, "expression")) +
    #ggtitle(paste(type, "for gene", gn, "in", ovr_column, "\nsamples, adjusted for clinical covariates")) +
    ggtitle(paste("Adjusted curve for", ovr_column,"samples")) +
    #annotate(geom = "text", label = paste0("Coef pval for ntile: ", p_ntile), x= 25, y =0.1) +
    #annotate(geom = "text", label = paste0("Anova pval vs reduced: ", aod_p), x= 25, y =0.2) +
    annotate(geom = "text", label = paste("p =", p), x= 25, y =0.1) +
    theme(plot.title = element_text(size = 12),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.position = legend_positions[3]
    )
  
  reslist$adj_curv1 <- adj_curv1
  
  #Cox models for rest group, adjusted for all covariates and ntile gene expression
  c <- survival::coxph(formula, data = sd0)
  #print(c)
  
  #Store p value associated with the coefficient of the ntile
  fit_ntile <- as.data.frame(summary(c)$coefficients)
  p_ntile <- signif(fit_ntile[rownames(fit_ntile) == "ntile", ]$`Pr(>|z|)`, 2)
  if (p_ntile < 0.001){
    p_ntile <- "< 0.001"
  }
  
  #Anova p value vs reduced model
  d <- survival::coxph(reduced, data = as.data.frame(sd0))
  aod <- anova(d,c) #Conventional to list models from smallest to largest but order does not matter
  aod_p <- signif(aod[4][2,], 2)
  if (aod_p < 0.001){
    aod_p <- "< 0.001"
  }
  
  if (p_method == "anova"){
    p <- aod_p
  } else {
    p <- p_ntile
  }
  
  #Adjusted survival curve for samples belonging to group of interest
  #Use surv_adjustedcurves to get data and the then manually plot for more control over appearance
  adj_curv0 <- survminer::surv_adjustedcurves(fit = c, variable = "ntile", data = as.data.frame(sd0), method = "conditional") %>%
    left_join(select(mutate(sd0, ntile = as.factor(ntile)), ntile, labels), by = c("variable" = "ntile")) %>%
    ggplot(., aes(x = time, y = surv, color = labels)) + 
    geom_step(size = 1) + 
    theme_survminer() + 
    line_colors +
    scale_y_continuous(limits = c(0,1)) + 
    #scale_x_continuous(limits = c(0,250)) +
    ylab(label =  ylab) +
    labs(color = paste(gn, "expression")) +
    #ggtitle(paste(type, "for gene", gn, "\nin other samples, adjusted for clinical covariates")) +
    ggtitle(paste("Adjusted curve for other samples")) +
    #annotate(geom = "text", label = paste0("Coef pval for ntile: ", p_ntile), x= 25, y =0.1) +
    #annotate(geom = "text", label = paste0("Anova pval vs reduced: ", aod_p), x= 25, y =0.2) +
    annotate(geom = "text", label = paste("p =", p), x= 25, y =0.1) +
    theme(plot.title = element_text(size = 12),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.position = legend_positions[4]
          )
  
  reslist$adj_curv0 <- adj_curv0
  
  if(return_list == T){
    return(reslist)  
  }
  
  # Arrange them all together
  arplot <- ggpubr::ggarrange(plotlist = reslist, nrow = 2, ncol = 2)
  
  arplot <- annotate_figure(arplot,
                  top = text_grob(paste(type, "for gene", gn), face = "bold", size = 14),
                  #bottom = text_grob("Time (Months)", size = 12),
                  bottom = text_grob(paste0("Time (Months)\n\nClinical covariates: ",
                                            paste(cov_legend, collapse=", ")),
                                     size = 12),
                  left = text_grob(ylab, rot = 90),
                  #right = "I'm done, thanks :-)!",
                  #fig.lab = "Figure 1", fig.lab.face = "bold"
  )
  
  return(arplot)
  
}

km_ntiles_ovr("MARVELD2")

```

## Retrieve gene summaries

Webscrape the entrez and uniprot summaries for a gene.

```{r, include=params$show_functions}
get_entrez_summary <- function(id, dict){
  
  ent <- unique(bx_annot[bx_annot$gene_name == id, ]$entrez_id)
  
  if(length(ent) == 0){
    print("No ID found")
    return(NA)
  }
  
  gene_summary <- rentrez::entrez_summary(db = "gene", id = ent)
  
  res <- list()
  
  res$name <- rentrez::extract_from_esummary(gene_summary, "name")
  res$description <- rentrez::extract_from_esummary(gene_summary, "description")
  res$otheraliases <- rentrez::extract_from_esummary(gene_summary, "otheraliases")
  #res$otherdesignations <- rentrez::extract_from_esummary(gene_summary, "otherdesignations")
  res$entrez_summary <- rentrez::extract_from_esummary(gene_summary, "summary")
  return(res)
}

get_entrez_summary("MARVELD2")
```

Alternative for uniprot: Use proteinatlas.org.
Also shows known prognostics.

```{r, include=params$show_functions}
#TODO Make html_text work better somehow, or maybe use proteinatlas.org instead

get_uniprot_summary <- function(id){
  
  uni <- unique(bx_annot[bx_annot$gene_name == id, ]$uniprot_id)
  
  if(length(uni) == 0){
    print("No ID found")
    return(NA)
  }
  
  base_url = "https://www.uniprot.org/uniprot/"
  url = paste0(base_url, uni)
  
  uniprot_summary <- tryCatch(
  xml2::read_html(url) %>%
    rvest::html_node(xpath='/html/body/main/div/div[3]/div[3]/div[1]') %>%
    rvest::html_text(),
    error = function(e){NA} 
  )
  
  uniprot_summary <- uniprot_summary %>% str_split("[:digit:] Publications")
  uniprot_summary <- uniprot_summary[[1]][[1]]
  return(uniprot_summary)
}

get_uniprot_summary("MS4A1")
```

## Elastic net Cox features

Show whether a gene was included as a potential feature by virtue of being in the top 1000 or top 5000 most variable genes.
If a gene as multiple ensembl IDs, multiple columns will be returned.

```{r, include=params$show_functions}
gene_enet_candidate <- function(id){
  
  gn = gene_lookup(id)$gene_name
  en = gene_lookup(id)$ensembl_gene_id
  
  top5000var = names(head(sort(apply(ens_mat, 1, var), decreasing = T), 5000))
  top1000var = names(head(sort(apply(ens_mat, 1, var), decreasing = T), 1000))
  
  cand = rbind(
    sym = gn,
    intop1000 = en %in% top1000var,
    intop5000 = en %in% top5000var
    )
  colnames(cand) = en
  cand = as.data.frame(cand)
  return(cand)
  
}
gene_enet_candidate("MS4A1")
gene_enet_candidate("MARVELD2")
```

If a gene has been selected as a feature, show the stability and effect size.

```{r, include=params$show_functions}
gene_enet_feature <- function(id){
  
  en <- gene_lookup(id)$ensembl_gene_id
  gn <- gene_lookup(id)$gene_name
  
  res <- glm_df %>%
    filter(feature == !! gn)
  
  if(nrow(res) == 0){
    return(paste(gn, "not selected as predictive feature"))
  }
  
  res %>%
    arrange(desc(n), desc(abs(mean_beta))) %>%
    mutate(mean_beta = signif(mean_beta, 2),
           sd_beta = signif(sd_beta, 2)) %>%
    select(-description:-gene_input)
}

gene_enet_feature("MS4A1")
gene_enet_feature("MARVELD2")
```



# Gene biology

Entrez gene summary:
Text wrapping is WIP.

```{r}
#TODO fix text wrapping.
get_entrez_summary(Gene)
```

Uniprot gene summary (html cleanup is WIP):

```{r}
get_uniprot_summary(Gene)
```


# Survival summary

Depicting data frame results for genewise and interaction models.

## Genewise models

For univariate models, the formula is: Surv(time, event)~gene.

For multivariate models, the formula is:
Surv(time, event)~age + stage + grade + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + strata(PAM50) + gene

```{r}
genewise_survival_results(Gene, df = gw_surv, return_first = F)
```

## Interaction models

Currently, all interaction models are multivariate.
The formula is:

Surv(time, event)~ age + stage + grade + surgery + radiotherapy + strata(hormonetherapy) + chemotherapy + herceptin + strata(PAM50) + involution + gene + involution * gene

```{r}
genewise_survival_results(Gene, df = int_surv, return_first = F)
```



# Differential expression

Likelihood ratio test, pairwise comparisons, and one vs rest comparisons:

```{r}
diffex_report(Gene, subgroup = F)
```

Subgroup analysis:

```{r}
diffex_report(Gene, subgroup = T)
```



# Counts by study group

Log2-TMM normalized

```{r, fig.width=8}
box_plots(Gene)$tmm
```

Size-factor normalized:

```{r, fig.width=8}
box_plots(Gene)$sf
```



# Kaplan-Meiers

Involution vs rest survival depictions

## Overall survival

```{r, warning=F, fig.height=7, fig.width=10}
km_ntiles_ovr(Gene, ovr_column = "involution", survival_type = "os")
```

## Distant recurrence

```{r, warning=F, fig.height=7, fig.width=10}
km_ntiles_ovr(Gene, ovr_column = "involution", survival_type = "drs")
```


# Elastic net predictive power

Elastic net Cox models predict outcome (death or metastasis) based on:

* top 1000 most variable genes
* top 5000 most variable genes
* all genes that pass the minimum diffex count threshold (~30K)

Show whether this gene was among the top 1000 or 5000 most variable.

```{r}
gene_enet_candidate(Gene)
```

If this gene was selected as a feature, show the feature stability (n, sd beta) and effect size (mean beta).

```{r}
gene_enet_feature(Gene)
```


