---
title: "Experiments with limma and combat"
output: html_notebook
---


```{r, include=F}
library(DESeq2)
library(limma)
library(edgeR)
library(sva)
library(here)
library(tidyverse)
```

DESeq can't be run with transformed data, and ComBat requires that the input matrix be cleaned and normalized beforehand. 

We'll therefore need to use limma to perform differential expression analysis on the ComBat-corrected counts.

# Load and filter

Our pipeline has been in DESeq so far.

```{r}
dds = readRDS(here("data", "Rds", "05_dds_PAM50_tumorpurity_batch.Rds"))
gx_annot <- read_tsv(here("data/metadata/01_tx_annot.tsv"))
gx_annot = gx_annot %>% select(ensembl_gene_id = gene_id, gene_name, gene_type, description = gene_description) %>% distinct()
sd = as.data.frame(colData(dds))
```

Set up one vs rest comparisons


```{r}
dds$inv_vs_rest  = colData(dds) %>% as.data.frame() %>% mutate(inv_vs_rest = if_else(study_group == "ppbc_inv", "ppbc_inv", "rest")) %>%
  pull(inv_vs_rest) %>% factor(levels=c("rest", "ppbc_inv"))

dds$prbc_vs_rest  = colData(dds) %>% as.data.frame() %>% mutate(prbc_vs_rest = if_else(study_group == "prbc", "prbc", "rest")) %>%
  pull(prbc_vs_rest) %>% factor(levels=c("rest", "prbc"))

dds$lac_vs_rest  = colData(dds) %>% as.data.frame() %>% mutate(lac_vs_rest = if_else(study_group == "ppbc_lac", "ppbc_lac", "rest")) %>%
  pull(lac_vs_rest) %>% factor(levels=c("rest", "ppbc_lac"))

dds$nonprbc_vs_rest  = colData(dds) %>% as.data.frame() %>% mutate(nonprbc_vs_rest = if_else(study_group == "non_prbc", "non_prbc", "rest")) %>%
  pull(nonprbc_vs_rest) %>% factor(levels=c("rest", "non_prbc"))

table(dds$inv_vs_rest, dds$study_group)
table(dds$prbc_vs_rest, dds$study_group)
table(dds$lac_vs_rest, dds$study_group)
table(dds$nonprbc_vs_rest, dds$study_group)
```

Discard genes that don't have a cpm of at at least 1 in n samples, where n is the size of the smallest study group.

```{r}
print(paste("Started with", nrow(dds), "genes"))
genEx = counts(dds, normalize=F)[rowSums(fpm(dds, robust = F) > 1) >= (sd %>% group_by(study_group) %>%
                                        summarise(n=n()) %>% arrange(n) %>%
                                        pull(n) %>% head(1)),]
print(paste("After filtering,", nrow(genEx), "genes"))
```

# Combat

Set up DGEList

```{r}
dge = DGEList(counts = genEx, samples = sd, group = sd$study_group,
              genes = gx_annot[gx_annot$ensembl_gene_id %in% rownames(genEx),])
```

Normalize and voom transform

```{r}
dge  = calcNormFactors(dge, method="TMM")

vmodel = model.matrix(~ PAM50 + tumor_purity + study_group, data = dge$samples)

v = voom(dge, design = vmodel, log=T, normalized.lib.sizes=T, plot = T)

```

Apply ComBat to voom-transformed data.

```{r}
modcombat = model.matrix(~1, data=v$targets)

combat_counts <- sva::ComBat(
  dat = v$E, batch = v$targets$batch, mod = modcombat, 
  par.prior = TRUE, prior.plots = FALSE
  )
```


# Diffex with Limma

# Pairwise

Our design matrix:

```{r}
head(vmodel)
```

Basic approach, with non-prbc as reference level:

```{r}
fit = lmFit(combat_counts, design=vmodel)
efit = eBayes(fit)
```

Add gene annotation

```{r}
efit$genes = v$genes
```


## Inv vs non-prbc

```{r}
print(paste("Significant genes with adjusted p val < 0.05:", nrow(topTable(efit, coef = 7, p.value = 0.05, number = Inf))))
topTable(efit, coef = 7, p.value = 0.05, number = Inf)
```

## Lac vs non-prbc

```{r}
print(paste("Significant genes with adjusted p val < 0.05:", nrow(topTable(efit, coef = 8, p.value = 0.05, number = Inf))))
topTable(efit, coef = 8, p.value = 0.05, number = Inf)
```

## Prbc vs non-prbc

```{r}
print(paste("Significant genes with adjusted p val < 0.05:", nrow(topTable(efit, coef = 9, p.value = 0.05, number = Inf))))
topTable(efit, coef = 9, p.value = 0.05, number = Inf) 
```

# One vs rest


```{r}
model.matrix(~ PAM50 + tumor_purity + inv_vs_rest, data = colData(dds)) %>% head()
```

Basic approach, with non-prbc as reference level:

```{r}
fit = lmFit(combat_counts, design=model.matrix(~ PAM50 + tumor_purity + inv_vs_rest, data = colData(dds)))
efit = eBayes(fit)
```

Add gene annotation

```{r}
efit$genes = v$genes
```


## Inv vs rest

```{r}
print(paste("Significant genes with adjusted p val < 0.05:", nrow(topTable(efit, coef = 7, p.value = 0.05, number = Inf))))
topTable(efit, coef = 7, p.value = 0.05, number = Inf)
```

## Prbc vs rest


```{r}
fit = lmFit(combat_counts, design=model.matrix(~ PAM50 + tumor_purity + prbc_vs_rest, data = colData(dds)))
efit = eBayes(fit)
```

Add gene annotation

```{r}
efit$genes = v$genes
```


```{r}
print(paste("Significant genes with adjusted p val < 0.05:", nrow(topTable(efit, coef = 7, p.value = 0.05, number = Inf))))
topTable(efit, coef = 7, p.value = 0.05, number = Inf)
```

## Non-prbc vs rest

```{r}
fit = lmFit(combat_counts, design=model.matrix(~ PAM50 + tumor_purity + nonprbc_vs_rest, data = colData(dds)))
efit = eBayes(fit)

efit$genes = v$genes

print(paste("Significant genes with adjusted p val < 0.05:", nrow(topTable(efit, coef = 7, p.value = 0.05, number = Inf))))
topTable(efit, coef = 7, p.value = 0.05, number = Inf)
```


## Lac vs rest

```{r}
fit = lmFit(combat_counts, design=model.matrix(~ PAM50 + tumor_purity + lac_vs_rest, data = colData(dds)))
efit = eBayes(fit)

efit$genes = v$genes

print(paste("Significant genes with adjusted p val < 0.05:", nrow(topTable(efit, coef = 7, p.value = 0.05, number = Inf))))
topTable(efit, coef = 7, p.value = 0.05, number = Inf)
```


```{r}
save.image(here("reports", "experiments_limma_combat.RData"))
```



