---
title: "RNA metadata"
date: "`r Sys.Date()`"
author: "Kat Moore"
output: 
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    highlight: kate
    df_print: paged
params:
  sampledata: "data/external/sample_data.tsv"
  patientdata: "data/external/patient_data.tsv"
  salmondir: "data/rnaseq/salmon"
  pre_excluded_samples: "data/external/pre_excluded_samples.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(here)
library(tidyverse)
```

Match the RNAseq fastq-derived IDs with their corresponding identifiers and read the salmon quantifications with tximport.

Some fastqs were pre-excluded:

```{r}
pre_excluded_samples <- read_csv(here(params$pre_excluded_samples), show_col_types = F)

pre_excluded_samples
```

## Sample data

```{r}
sampledataRaw <- read_tsv(here(params$sampledata))
```

```{r}
glimpse(sampledataRaw)
```

General overview all sample and platforms

```{r}
sampledataRaw %>% group_by(sample_type, experimental_platform, Included) %>% count()
```

Batch HALO is only relevant for spatial data (slide type).

```{r}
sampledataRaw %>% 
  mutate(batch_Leuven = ifelse(!is.na(batch_Leuven), "exists", "missing")) %>%
  mutate(batch_HALO = ifelse(!is.na(batch_HALO), "exists", "missing")) %>% 
  group_by(sample_type, batch_Leuven, batch_HALO) %>% count()
```

Duplicate patient IDs may exist where a sample has been sequenced more than once. Column typing is therefore not set until later.

```{r}
source(here("src/rnaseq/read_rna_sampledata.R"))
sampledata <- read_rna_sampledata(here(params$sampledata), verbose = T)
```

```{r}
head(sampledata)
```

## Patient data

```{r}
patientdataRaw <- read_tsv(here(params$patientdata))
glimpse(patientdataRaw)
```

Sample availability by study group

```{r}
patientdataRaw %>%
  group_by(study_group, PPBC) %>%
  count()
```

Clinical subtyping by study group

```{r}
patientdataRaw %>%
  group_by(study_group, clin_subtype) %>%
  count() %>%
  pivot_wider(names_from = study_group, values_from = n, values_fill = 0)
```

A number of columns contain median or tertile calculation that were made manually in Excel, without a formula. These columns pertain to the TAPC, TIL, and CD38 scoring. 

The raw scoring from pathologists is kept, the median/tertiles will be recalculated downstream.

```{r}
source(here("src/utils/read_patient_data.R"))
patientdata <- read_patient_data(here(params$patientdata), verbose = T)
```

```{r}
head(patientdata)
```

## RNA metadata

Combine patient and sample data.

```{r}
rnaMeta <- left_join(sampledata, patientdata, by = "patient_ID")

rnaMeta
```

## Check missing files

```{r}
salmonfiles <- list.files(here(params$salmondir))
```

Samples that have no corresponding fastq.

```{r}
no_fastq <- sampledata %>%
  filter(!str_remove(sampledata$fastq, ".fastq.gz") %in% salmonfiles)

no_fastq

filecheck <- function(df, fn){
  
  if(file.exists(fn)){file.remove(here(fn))}
  
  if(nrow(df) > 0){write_csv(df, fn)}
  
}

filecheck(no_fastq, fn = here("data/rnaseq/metadata/01_nofastq.csv"))
```

Salmon output with no link to metadata file.

```{r}
no_metadata <- salmonfiles[!salmonfiles %in% str_remove(sampledataRaw$sample_ID,
                                                        ".fastq.gz")]
no_metadata
```

Are these samples among those pre-excluded?

```{r}
no_metadata_excluded <- pre_excluded_samples %>%
  filter(fastq_file %in% no_metadata)
no_metadata_excluded
```

Show those which cannot be found in the pre-excluded list.

```{r}
missing_metadata <- pre_excluded_samples %>%
  filter(!fastq_file %in% no_metadata)

missing_metadata

filecheck(missing_metadata, fn = here("data/rnaseq/metadata/01_missing_metdata.csv"))
```

## Duplicate summary

These patients have duplicate samples:

```{r}
duplicate_samples <- rnaMeta %>%
  filter(patient_ID %in% filter(rnaMeta, duplicated(patient_ID))$patient_ID) %>%
  arrange(patient_ID) %>% select(patient_ID, fastq, batch)

duplicate_samples
```

Duplicates per sample:

```{r}
duplicate_samples %>% group_by(patient_ID) %>% count() %>%
  group_by(n) %>% count(name = "number of samples with n duplicates")
```

## Write metadata

```{r}
saveRDS(rnaMeta, here("data/rnaseq/metadata/01_rnaMeta.Rds"))
```

## Session info

```{r}
sessionInfo()
```

