---
title: "Color Palettes"
date: "`r Sys.Date()`"
author: "Kat Moore"
output: 
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 4
    highlight: kate
    df_print: paged
params:
  dds: "data/rnaseq/interim/03_dds_PAM50.Rds"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(tidyverse)
library(here)
library(RColorBrewer)
library(DESeq2)
```

This notebook defines color palettes to be used consistently in plots throughout the pipeline.

```{r}
dds <- readRDS(here(params$dds))
```

```{r}
color_grid = function (colours, labels = T, names = T, borders = NULL, cex_label = 1)
{
  
  #Adapted from scales' show_colors
  #Visualizes colors chosen for each study group
  
  n <- length(colours)
  ncol <- ceiling(sqrt(n))
  nrow <- ceiling(n/ncol)
  
  #Also add option to plot group labels
  colornames = c(names(colours), rep(NA, nrow * ncol - length(names(colours))))
  colornames <- matrix(colornames, ncol = ncol, byrow = TRUE)
  
  #Original scales code
  colours <- c(colours, rep(NA, nrow * ncol - length(colours)))
  colours <- matrix(colours, ncol = ncol, byrow = TRUE)
  
  
  old <- par(pty = "s", mar = c(0, 0, 0, 0))
  on.exit(par(old))
  size <- max(dim(colours))
  plot(c(0, size), c(0, -size), type = "n", xlab = "", ylab = "",
       axes = FALSE)
  rect(col(colours) - 1, -row(colours) + 1, col(colours), -row(colours),
       col = colours, border = borders)
  
  
  
  if (labels) {
    text(col(colours) - 0.5, -row(colours) + 0.3, colours,
         cex = cex_label)
  }
  
  #Plot the group labels
  if (names) {
    text(col(colornames) - 0.5, -row(colornames) + 0.5, colornames,
         cex = cex_label)
  }
}
```

## Define PPBC colors

Hanne has indicated her preferred palette for study group.
Demonstrate palette for both the encoded and non-encoded version.

Encoded version:

```{r}
levels(dds$study_group)
```

Full version:

```{r}
levels(dds$PPBC)
```

```{r}
study_colors = c(
  ppbcpw = "#46CDCF",
  ppbcdl= "#112D4E",
  prbc = "#AA96DA",
  npbc = "#521262",
  rest = "#F73859"
)
color_grid(study_colors)

ppbc_colors = c(
  involuting = "#46CDCF",
  lactating = "#112D4E",
  pregnant = "#AA96DA",
  nulliparous = "#521262",
  rest = "#F73859"
)

color_grid(ppbc_colors)
```

## Define PAM50 colors

```{r}
levels(dds$PAM50)
```

```{r}
pam_colors <- suppressWarnings(
  colorRampPalette(
    brewer.pal(length(unique(dds$PAM50)),"Paired"))(length(unique(dds$PAM50)))
  )

names(pam_colors) = c(
  "Basal",
  "Normal",
  "LumA",
  "LumB",
  "Her2"
)

color_grid(pam_colors)
```

## Define gene biotype colors

Reduce the possible types of genes down to those defined in deseq_report_functons::annotate_results()

```{r}
gene_types <- c("other noncoding", "immune protein coding",
                "protein coding", "immunoglobulin", "T cell receptor")
gene_colors <- suppressWarnings(
  colorRampPalette(
    brewer.pal(length(unique(gene_types)),"Dark2"))(length(unique(gene_types)))
  )
names(gene_colors) <- gene_types
color_grid(gene_colors)
```

## Tertile colors

```{r}
tertile_colors <- colorRampPalette(brewer.pal(3,"YlOrRd"))(3)
names(tertile_colors) <- c("low", "medium", "high")
color_grid(tertile_colors)
```

```{r}
saveRDS(list(study_colors = study_colors,
             ppbc_colors = ppbc_colors,
             pam_colors = pam_colors,
             gene_colors = gene_colors,
             tertile_colors = tertile_colors
             ),
        here("data", "rnaseq", "interim", "color_palettes.Rds"))
```

## Session info

```{r}
sessionInfo()
```
