---
title: "Minimum count threshold"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    theme: darkly
    highlight: breezedark
    canon: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DESeq2)
library(here)
library(tidyverse)

```

Load count matrix:

```{r}
dds <- readRDS(file = here("data/Rds/05_dds_PAM50_batch.Rds"))

dim(dds)
```

Minimum count threshold: remove any features (rows) without at least 1 read per million in n of the samples, where n is the size of the smallest group of replicates [@anders2013].

```{r}
table(dds$PPBC)
```
In our case, one group (lactating) is much smaller than the rest.
We therefore choose the lowest n in all groups except lactating.

Minimum n:

```{r, eval=F, include=F}
#This is how it would be if we chose lactating as the minimum group
min.n <- min(table(dds$PPBC))

min.n
```

```{r}
min.n <- min(table(dds$PPBC)[names(table(dds$PPBC))!="lactating"])

min.n
```

Create index for which genes to keep:
at least 1 cpm in `r min.n` of the samples

```{r}
keep <- rowSums(fpm(dds, robust = F)) >= min.n
table(keep)
```

The old threshold was as follows:
Require non-zero counts in 1/3 of the dataset.

```{r}
old.keep <- rowSums(counts(dds)!=0) >= ceiling(ncol(dds)/3)
table(old.keep)
```

Subset the dds and save:

```{r}
dds <- dds[keep,]
dim(dds)

saveRDS(dds, here("data/Rds/05b_dds_filtered.Rds"))
```


