---
title: "GSEA transcription factors"
Author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    highlight: kate
params:
  dds: "data/rnaseq/processed/08_dds_ovr_inv_vs_rest.Rds"
  ovr: "results/rnaseq/diffex/08_one_vs_rest_allgenes.xlsx"
  gmt: "data/external/gmt/c3.tft.v2022.1.Hs.symbols.gmt"
  gx_annot: "data/rnaseq/processed/bx_annot.Rds"
  dedup: "src/rnaseq/summarize_duplicate_ids.R"
  testing: FALSE
  overwrite: FALSE
---

## Dependencies

```{r}
library(here)
library(DESeq2)
library(flexgsea)
library(tidyverse)
library(readxl)
library(circlize)
library(ComplexHeatmap)

theme_set(theme_bw())
```

## Load data and params

Count matrix.

```{r load_dds}
dds <- readRDS(here(params$dds))
```

GMT for TFs.

```{r load_gmt}
gmt <- flexgsea::read_gmt(here(params$gmt))
gmt[1]

names(gmt)[1:10]
```

Gene dictionary.

```{r load_gx}
gx_annot <- readRDS(here(params$gx_annot))
glimpse(gx_annot)
```

DESeq2 inv vs rest results.

```{r load_ivr}
ivr <- readxl::read_excel(here(params$ovr), sheet = "rep_inv_rest_all") %>%
  mutate(sig = ifelse(padj < 0.05 & abs(log2FoldChange) > 0.5, "*", NA))

ivr$padj[is.na(ivr$padj)] <- 1 # for visualization
glimpse(ivr)
```

Load gene symbol summarization function.

```{r}
source(here(params$dedup))
```

Set logical class for inherited params.

```{r}
testing <- as.logical(params$testing)
testing
```

Number of permutations

```{r}
nperm <- ifelse(testing, 3, 1000)
nperm
```

Rerun analysis if run before?

```{r}
overwrite <- as.logical(params$overwrite)
overwrite
```

```{r}
study_colors <- c("npbc"="#521262", "prbc"="#AA96DA",
                  "ppbcdl"="#112D4E", "ppbcpw"="#46CDCF")
```

## Convert to gene symbol

Extract raw counts, using fewer genes if testing.

```{r}
if(testing){dds <- head(dds,500)}
counts <- assay(dds)
```

Convert ensembl ID to gene symbol

```{r}
df <- right_join(
  gx_annot[,c("ensembl_gene_id","gene_name")],
  rownames_to_column(as.data.frame(counts),"ensembl_gene_id"),
  by="ensembl_gene_id"
)

dim(df)
```

Remove genes with no gene symbol.

```{r}
print(paste("Number of ensembl IDs with no corresponding gene symbol:", sum(is.na(df$gene_name))))
df <- df[!is.na(df$gene_name),]
```

De-duplicate genes. When dealing with duplicated gene names, take the mean expression

```{r}
geneEx <- df %>% select(-ensembl_gene_id)
```

```{r}
geneEx <- summarize_expression_duplicate_ids(mat = geneEx, id_column = "gene_name", f=colMeans)
```

## Gene scoring function

The scoring function ranks genes by DESeq2 p value using a formula derived from the column names of `y`.
In our case, this is `~batch + PAM50 + inv_vs_rest`. 
It will return the p value of the last covariate in the design formula when `coef_of_interest` is set to NULL. This is most convenient when the covariate is a binary factor.
When the covariate has multiple levels, such as `study_group`, using `coef_of_interest` will return all such levels at once. 
This saves quite a bit of time.
Regardless, the p value is multiplied by the sign of l2fc so that enriched and depleted genes are ranked accordingly.

```{r}
make_score_genes <- function(coef_of_interest="study_group") {
  score_genes_deseq <- function(x, y, abs) {
    
    stopifnot(coef_of_interest %in% colnames(y))
    
    #Required for permutation test
    rownames(x) <- rownames(y)
    genes <- colnames(x)
    
    #Create dds with formula derived from column names
    dds <- DESeq2::DESeqDataSetFromMatrix(t(x), y, as.formula(paste("~", paste(colnames(y), collapse="+"))))
    #print(design(dds))
    dds <- DESeq2::DESeq(dds)
    
    # Extract result names
    coefficients <- DESeq2::resultsNames(dds) %>%
      discard(~ . == 'Intercept') %>%
      set_names(., .)
    
    #Subset down to column(s) of interest
    if (!is.null(coef_of_interest)){
      
      #Retrieve the levels from each of the columns of interest
      ourgroups = as.character(unlist(apply(y[,coef_of_interest, drop=F], 2, unique)))
      
      # Subset results names by pattern matching to levels in columns of interest
      # Mapping returns multiple non-unique matches as levels will
      # appear more than once in the design matrix
      # Use ~keep instead of ~detect
      # map_chr expects a single match and will crash when multiple hits are found
      # Solution: Use map instead of map_chr and unlist
      coefficients =
        unique(
          unlist(map(ourgroups, ~ keep(coefficients, str_detect, .x)))
        )
      #coefficients
    } else {
      coefficients = tail(coefficients, 1) #Last one in group
    }
    
    #Shrink coefficients and extract pval and l2fc
    #Include l2fc so we can adjust by sign
    res <- coefficients %>%
      map_dfc(~ as.data.frame(
        lfcShrink(dds, coef=., type='apeglm', apeMethod='nbinomC')
      )[,c("pvalue","log2FoldChange")]) 
    rownames(res) <- genes
    
    #Separate p val
    p <- res %>% select(contains("pval"))
    colnames(p) <- coefficients
    p <- data.matrix(p)
    p[is.na(p)] <- 1.0
    
    #Create data frame indicating whether l2fc is above or below 0
    l2fc <- res %>% select(contains("log2FoldChange"))
    colnames(l2fc) <- coefficients
    l2fc <- data.matrix(l2fc)
    sign <- ifelse(l2fc > 0, 1, -1)
    
    if (abs) {
      return(1-p)
    } else {
      return((1-p) * sign )} #Multiply pval by -1 if l2fc < 0
  }}
```

## Flexgsea

Use inv vs rest comparison.

```{r}
print(levels(dds$inv_vs_rest))
print(table(dds$study_group, dds$inv_vs_rest))
comp <- "inv_vs_rest"
```

Create a matrix, convert any decimals created by rounding to nearest integer (rounding down) and tranpose.

```{r, matrix_integers}
geneEx <- as.matrix(column_to_rownames(remove_rownames(geneEx), "GeneSymbol"))

x <- t(geneEx)
x <- apply(x, 2, as.integer)

dim(x)
```

Select the relevant sample data to generate the design formula.

```{r}
y <- as.data.frame(colData(dds)) %>% select(batch, PAM50, !!comp)
y %>% glimpse()
stopifnot(nrow(x)==nrow(y))
```

```{r flexgsea}
print(paste("Starting FlexGSEA for comparison", comp,
            "with gene signature", basename(params$gmt)))

#Test whether analysis has been run before
outFile <- here("data/rnaseq/processed/TF_flexgsea_invVsRest.Rds")


#To save time, we design the scoring function to return pairwise comparisons simultaneously
#For one vs rest, selects the last column only (which corresponds to one vs rest)
if (comp == "study_group"){
  this.comp = "study_group"
} else {
  this.comp = NULL
}

if(!file.exists(outFile) | overwrite == T){
  start <- Sys.time()
  flexres <- flexgsea(x = x, y = y,
                      gene.sets = gmt,
                      gene.score.fn = make_score_genes(coef_of_interest=this.comp),
                      es.fn = flexgsea_weighted_ks, #flexgsea_maxmean
                      sig.fun=flexgsea_calc_sig,
                      nperm=nperm, parallel = T,
                      block.size = 1,
                      abs = FALSE,
                      return_values = c("gene_name","leading_edge",'running_es_pos',
                                        'running_es_neg', "running_es_at"))
  saveRDS(flexres, outFile)
  end <- Sys.time()
  print(end - start)
} else {
  flexres <- readRDS(outFile)
}

```

## Results

Available results:

```{r}
names(flexres)
```

Table of sorted results.

```{r}
flexres$table$inv_vs_rest_ppbcpw_vs_rest %>%
  arrange(fdr)
```

TFs below fdr 0.05:

```{r}
flextable <- flexres$table$inv_vs_rest_ppbcpw_vs_rest %>%
  mutate(sig = ifelse(fdr < 0.05, "sig", "n.s.")) %>%
  mutate(comparison = "PPBCpw vs rest")

flextable %>%
  pull(sig) %>% table()
```

Names of significant TFs:

```{r}
flextable %>%
  filter(fdr < 0.05) %>%
  pull(GeneSet)
```

## NES bar plot

Show normalized enrichment score of significant pathways.

```{r}
plot_top_pathway <- function(df, thresh, top_n, comparison){
  
  stopifnot(comparison %in% unique(df$comparison))
  comp <- df[df$comparison == comparison, ]
  comp_thresh <- comp[comp$fdr < thresh, ]
  top_comp <- comp_thresh %>%
    arrange(fdr, p) %>%
    head(top_n)
  
  if(nrow(top_comp) <1 ){
    return(paste("No sig pathways for", comparison,"at FDR threshold", thresh))
  }
  
  top_comp %>%
    #truncate long terms
    mutate(GeneSet = substr(GeneSet,1,50)) %>%
    ggplot(aes(x = GeneSet, y = nes, fill = fdr)) +
    geom_bar(stat = "identity") + coord_flip() +
    ggtitle(paste("Top", top_n, comparison, "TF gene sets at FDR", thresh,
                  paste0("\n(", nrow(comp_thresh), " total hits)")))
  
}

plot_top_pathway(flextable, thresh = 0.05, top_n = 10, comparison = "PPBCpw vs rest")
```

## Heatmap of TF targets

Show only significant pathways.

```{r}
sigtfs <- flextable %>% filter(fdr < 0.05)
sigtfs

show_heatmaps <- nrow(sigtfs) > 0
```

Show the heatmaps only if some significant TFs are found.

```{r}
show_heatmaps
```


```{r heatmaps, fig.height=8, eval=show_heatmaps, echo=show_heatmaps}
for(i in 1:nrow(sigtfs)){
  tf <- sigtfs$GeneSet[i]
  tfFDR <- signif(sigtfs$fdr[i], 2)
  tfNES <- signif(sigtfs$nes[i], 2)
  
  htitle <- paste(str_replace_all(tf, "_", " "), "\nGSEA FDR:", tfFDR, "NES:", tfNES)
  
  tfmat <- geneEx[rownames(geneEx) %in% gmt[[tf]],]
  
  mat <- varianceStabilizingTransformation(tfmat)
  
  rowscalemat <- t(scale(t(mat)))
  
  pvalue_col_fun = colorRamp2(c(0, 2, 3), c("green", "white", "red"))
  
  # Reduce annotations to those contained in matrix, and discard duplicate symbols
  anno <- ivr[ivr$hgnc_symbol %in% rownames(rowscalemat),]
  anno <- anno[!duplicated(anno$hgnc_symbol),]
  
  anno$hgnc_symbol[duplicated(anno$hgnc_symbol)]
  
  stopifnot(all(sort(rownames(rowscalemat)) == sort(anno$hgnc_symbol)))
  
  hm <- Heatmap(rowscalemat, column_title = htitle,
                top_annotation = ComplexHeatmap::HeatmapAnnotation(
                  df=as.data.frame(colData(dds))[,c("study_group"), drop=F],
                  which = "col", show_legend = T,
                  col=list(study_group = study_colors),
                  annotation_legend_param = list(study_group = list(direction = "horizontal"))),
                right_annotation = HeatmapAnnotation(
                  pvalue = anno_simple(-log10(anno$pvalue),
                                       col = pvalue_col_fun, pch = anno$sig),
                  which = "row"),
                show_column_names = F,
                row_names_gp = gpar(fontsize = 8),
                heatmap_legend_param = list(
                  title = "row Z vst", direction = "horizontal"
                ))
  
  lgd_pvalue <- Legend(title = "p-value Deseq2", col_fun = pvalue_col_fun, at = c(0, 1, 2, 3), 
                       labels = c("1", "0.1", "0.01", "0.001"), direction = "horizontal")
  
  lgd_sig <- Legend(pch = "*", type = "points", labels = "DESeq2 FDR < 0.05 & abs(l2fc) > 0.5")
  draw(hm, annotation_legend_list = list(lgd_pvalue, lgd_sig),
       merge_legend = TRUE, heatmap_legend_side = "bottom", 
       annotation_legend_side = "bottom") 
}

```


## Session info

```{r}
sessionInfo()
```
