---
title: "Aggregation genewise survival"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    highlight: kate
    canon: true
params:
   resdir: "results/rnaseq/survival"
---

Create an overlap file of miscellaneous survival analyses.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(UpSetR)
library(RColorBrewer)
library(openxlsx)
library(here)
library(pheatmap)
library(tidyverse)

theme_set(theme_bw())
```

# Aggregate results

Files within results directory.

```{r}
filelist <- list.files(here(params$resdir), pattern=".csv", full.names = T)
filelist
```

Load all data frames related to genewise survival

```{r}
res_list <- purrr::map(filelist, read_csv, show_col_types = F)

names(res_list) <- str_remove(str_remove_all(basename(filelist), ".csv"),"12_")
```

Example input

```{r}
head(res_list$multi_genewise_os)
```

The number of genes/rows should be the same everywhere.

```{r}
stopifnot(all(unlist(lapply(res_list, function(x) nrow(x)==nrow(res_list$multi_genewise_os)))))
```

Aggregate Excel sheet.

```{r}
outFile <- here("results", "rnaseq", "survival", "12_cox_allgenes.xlsx")
openxlsx::write.xlsx(res_list, outFile)
```

The Excel sheet contains one tab with the same name and content as each element in `res_list`.

```{r}
stopifnot(all(readxl::excel_sheets(outFile) == names(res_list)))
```

Unified df:

```{r}
res = bind_rows(res_list) %>%
  arrange(fdr, p.value) %>%
  select(gene_name, fdr, p.value, baseMean, beta, `HR (95% CI for HR)`,
         analysis, everything())

head(res)
```

Sanity check: Number of genes in each comparison.

```{r}
unique(sapply(res_list, nrow))
```

# Global overview

What global significance level should we choose?
There are generally more hits in univariate than multivariate models (not strange).
Interestingly, there are more hits with drs models than os models.
Models run on involution samples only have hardly any hits, probably due to insufficient sample size.

```{r}
pt = list()
for (i in seq(0.05, 0.3, by = 0.05)){
  td = enframe(sapply(res_list, function(x) nrow(dplyr::filter(x, fdr <= i))))
  td = td[, "value", drop=F]
  colnames(td)[1] = as.character(i)
  pt[[length(pt)+1]] = td
}

pt = pt %>% bind_cols()
pt = t(as.matrix(pt))
colnames(pt) = str_remove(names(res_list), "genewise_")
pt
```

Overview analysis type

```{r}
ann_df = data.frame(row.names = colnames(pt),
                    input = if_else(str_detect(colnames(pt), "inv"),
                                    "involution samples", "all samples"),
                    formula = if_else(str_detect(colnames(pt), "multi"),
                                      "multivariate", "univariate"),
                    event = if_else(str_detect(colnames(pt), "os"),
                                    "overall survival", "distant recurrence")
)

ann_df
```

```{r}
quantile_breaks <- function(xs, n = 10) {
  breaks <- quantile(xs, probs = seq(0, 1, length.out = n))
  breaks[!duplicated(breaks)]
}

mat_breaks <- quantile_breaks(pt, n = 11)

pheatmap(pt, annotation_col = ann_df,
         main = "Significant hits at various fdr cutoffs",
         cluster_rows = F, cluster_cols = F,
         breaks =  mat_breaks,
         color = colorRampPalette(rev(brewer.pal(n = 8,
                                                 name = "RdYlBu")))(length(mat_breaks) - 1),
         display_numbers = T, number_format = "%0.f")
```

# Upset plots

Show overlap between hits at fdr cutoffs 0.05, 0.1, 0.15 and 0.2.

```{r, fig.width=8}
for(i in seq(0.05, 0.2, by = 0.05)){
 sig_genes = lapply(res_list, function(x) x[x$fdr <= i,]$ensembl_gene_id)
 
 #In a few cases, genes for which p value or fdr was NA will return NA
 #Remove these
 sig_genes <- lapply(sig_genes, function(x) x[!is.na(x)])

 #Remove items for which nothing is significant at the current threshold
 sig_genes <- sig_genes[unlist(lapply(sig_genes, function(x) length(x) > 0))]
 
 #We need at least 2 left in order to make an upset plot.
 if(length(sig_genes) < 2){next}
 
 print(upset(fromList(sig_genes), order.by = "freq", nsets = length(sig_genes),
       #empty.intersections = "on",
       #nsets = length(sig_genes), 
       set_size.show = TRUE,
       mainbar.y.label = paste("Cox FDR cutoff", i),
       #sets.bar.color = "blue", 
       #main.bar.color = "black", 
       set_size.scale_max = max(unlist(lapply(sig_genes, length))) + 125,
       sets.x.label = "Sig genes one vs rest comparisons") )
}
```

# Top 20 for every list:

```{r}
lapply(res_list, function(x) head(dplyr::select(x, gene_name:type), 20))
```

# Session info

```{r}
sessionInfo()
```

