---
title: "Genewise reports"
Author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
    highlight: kate
    canon: true
params:
  dds: "data/rnaseq/processed/08_dds_ovr_inv_vs_rest.Rds"
  tools: "src/rnaseq/deseq_report_functions.R"
  lrt: "results/rnaseq/diffex/06_LRT_allgenes.xlsx"
  vsd: "data/rnaseq/interim/08_vsd_ovr.Rds"
  gx_annot: "data/rnaseq/metadata/01_gene_annot.tsv"
  cp: "data/rnaseq/interim/color_palettes.Rds"
  sp: "data/rnaseq/interim/survival_colors.Rds"
  ivr: "results/rnaseq/diffex/08_one_vs_rest_allgenes.xlsx"
---

Check to see whether the IG gene signature is correlated with milk gene expression.

```{r, message=F}
library(DESeq2)
library(ComplexHeatmap)
library(tidyverse)
library(here)
```

## Load data {.tabset}

### DeseqDataSet

```{r}
dds = readRDS(here(params$dds))
```

### Sample annotation data

Derived from dds.

```{r}
sd = as.data.frame(colData(dds)) %>%
  select(sample_name, study_group, contains("rest"),
         PAM50, everything())
head(sd)
```

### Gene annotation

```{r}
gx_annot <- read_tsv(here(params$gx), show_col_types = F)
```

### Utility functions

```{r}
source(here(params$tools))
```

### Variance stabilizing transformation

```{r}
vsd = readRDS(here(params$vsd))
```

### Colors

```{r}
cp = readRDS(here(params$cp))
study_colors = cp$study_colors
pam_colors = cp$pam_colors
gene_colors = cp$gene_colors
ppbc_colors = cp$ppbc_colors

sp = readRDS(here(params$sp))
```

### Diffex results

```{r}
ivr <- readxl::read_excel(here(params$ivr))
head(ivr)
```

### LRT

Likelihood ratio test results, for DEGs from at least one study group.

```{r}
lrt <- readxl::read_excel(here(params$lrt))
head(lrt)
```

## Symbol conversion and de-deuplication

```{r}
geneEx = rownames_to_column(as.data.frame(assay(vsd)), "ensembl_gene_id")
geneEx = right_join(distinct(select(gx_annot, gene_name, ensembl_gene_id=gene_id)),
                    geneEx, by = "ensembl_gene_id") %>%
    select(-ensembl_gene_id)
geneEx = summarize_expression_duplicate_ids(geneEx, id_column = "gene_name", verbose = T)

```

## Milk proteins vs IG genes

Manual selection of some milk-associated genes, based on volcano plots of diffex genes and Cook's outliers.

```{r}
milk_genes <- c("CSN1S1", "CSN1S2AP", "CSN3", "LALBA")
```

Select differentially expressed IGs from the involution vs rest comparison.

```{r}
IG_genes <- ivr %>% filter(Type == "immunoglobulin") %>%
  filter(padj < 0.05 & abs(log2FoldChange) > 0.5) %>% pull(gene_name)
```

Subset the gene expression matrix to just milk and IG genes.

```{r}
milk_IG_mat <- geneEx[rownames(geneEx) %in% c(milk_genes, IG_genes),]
milk_IG_mat[1:6,1:6]
```

## Heatmap

Show the differentially expressed IG genes and the differentially expression milk genes in the involution vs rest comparison.

```{r, fig.height = 7, fig.width=10}
mat = scrime::rowScales(milk_IG_mat)
hlp = list(title="rowscaled vst")

#Ensure input is matrix
mat = as.matrix(mat)
  
#Heatmap annotation
ann_top = sd[,"PPBC", drop=F]
  
#Top column annotation
colTop <- HeatmapAnnotation(df=ann_top, which="col",
                            col = list(PPBC = ppbc_colors)
)

#Bottom column annotation
ann_bottom = sd[,"PAM50", drop=F]
colBottom <- HeatmapAnnotation(df=ann_bottom, which="col",
                               col = list(PAM50 = pam_colors))

anno_rows = data.frame(type = case_when(
  rownames(milk_IG_mat) %in% milk_genes ~ "milk protein",
  TRUE ~ "immunoglobulin"),
  row.names = rownames(milk_IG_mat))

#Essential that the order be the same!
anno_rows = anno_rows[match(rownames(mat),rownames(anno_rows)), ,drop=F]
  
rowAnno = HeatmapAnnotation(df=anno_rows, which="row",
                            col=list(type = c("milk protein" = "lightgray",
                                              "immunoglobulin" = "#E7298A"))
)
  
milk_hm = Heatmap(mat,
                  top_annotation = colTop,
                  bottom_annotation = colBottom,
                  left_annotation = rowAnno,
                  heatmap_legend_param = hlp,
                  show_row_names = T,
                  show_column_names = F,
                  cluster_rows = T,
                  row_names_gp = gpar(fontsize = 8),
                  column_names_gp = gpar(fontsize = 10),
                  column_title = "Milk genes vs IG genes")

milk_hm
```

```{r}
pdf(here("results", "rnaseq", "diffex", "milk_vs_IG_genes.pdf"),
    width = 10, height = 7)

milk_hm

dev.off()
```

## Beehives

```{r}
ensembl_milk <- gx_annot %>%
  filter(gene_name %in% milk_genes) %>%
  select(ensembl_gene_id=gene_id, gene_name) %>%
  distinct() %>%
  pull(ensembl_gene_id)

ensembl_milk <- ensembl_milk[ensembl_milk %in% lrt$ensembl_gene_id]

for (m in ensembl_milk){
  #print(res_list$LRT[res_list$LRT$ensembl_gene_id == m,])
  plot_gene_beehive(dds, result_df = lrt[lrt$ensembl_gene_id == m,],
                  title_string = "LRT", intgroup = "PPBC", color_vals = pam_colors) %>%
    print()
}

```

Potential todo? Check more milk proteins from [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3586783/)

## Session info

```{r}
sessionInfo()
```
