---
title: "Miscellaneous figures"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: yes
    toc_depth: 4
    highlight: tango
    df_print: tibble
---

```{r, include=F}
library(DESeq2)
library(ComplexHeatmap)
library(here)
library(tidyverse)
library("chisq.posthoc.test")
theme_set(theme_bw())
```

# Heatmap with involution split

Request: DEG heatmap inv versus rest also subdifferentiate inv in <6 months and >6months involution

## Gene expression and annotation

```{r}
dds = readRDS(here("data/Rds/08_dds_ovr_inv_vs_rest.Rds"))
vsd = readRDS(here("data", "Rds", "08_vsd_ovr.Rds"))
gx_annot <- read_tsv(here("data/metadata/01_tx_annot.tsv"))
#head(gx_annot)
gx_annot = gx_annot %>%
  select(ensembl_gene_id = gene_id, gene_name, hgnc_symbol,
         gene_type, description = gene_description) %>%
  distinct()
```

## ID conversion and deduplication

```{r}
source(here("src/general_R_tools.R"))
geneEx = rownames_to_column(as.data.frame(assay(vsd)), "ensembl_gene_id")
geneEx = right_join(select(gx_annot, gene_name, ensembl_gene_id),
                    geneEx, by = "ensembl_gene_id") %>%
    select(-ensembl_gene_id)
geneEx = summarize_expression_duplicate_ids(geneEx, id_column = "gene_name")
rownames(geneEx) = NULL
geneEx = column_to_rownames(geneEx, "GeneSymbol")
geneEx = as.matrix(geneEx)
geneEx[1:3, 1:3]
```

## Diffex results

```{r}
readxl::excel_sheets(here("results/diffex/08_one_vs_rest_sig_genes.xlsx"))
res_invr = openxlsx::read.xlsx(here("results", "diffex", "08_one_vs_rest_sig_genes.xlsx"), sheet="sig_rep_inv_rest")

dim(res_invr)
```

## Colors

```{r}
cp = readRDS(here("data", "Rds", "color_palettes.Rds"))
study_colors = cp$study_colors
pam_colors = cp$pam_colors
gene_colors = cp$gene_colors
ppbc_colors = cp$ppbc_colors

sp = readRDS(here("data/Rds/survival_colors.Rds"))
```

## Heatmaps

Row scaling

```{r}
mat = scrime::rowScales(geneEx)
mat[1:4, 1:4]
```

```{r}
sig_results = res_invr
maxn_genes = 1000

#Reduce genes to significant only
sig_results = sig_results %>% arrange(padj)
genestoplot = sig_results$gene_name[1:maxn_genes]
mat = mat[rownames(mat) %in% genestoplot, ]

dim(mat)
```

Top column annotation

```{r}
sampledata = as.data.frame(colData(dds))

#inv_nas = sampledata %>%
#  filter(is.na(involution_duration) & PPBC == "involuting")
#sampledata$involution_duration
ann_top = 
sampledata %>%
  mutate(involution_duration = as.character(involution_duration),
         PPBC = as.character(PPBC),
         overall_survival = as.character(overall_survival)) %>%
  mutate(PPBC.old = PPBC,
         PPBC = case_when(
           PPBC.old == "involuting" & is.na(involution_duration) ~ "inv unknown",
           PPBC.old == "involuting" & involution_duration == "<= 6 months" ~ "inv <= 6 months",
           PPBC.old == "involuting" & involution_duration == "> 6 months" ~ "inv > 6 months",
           TRUE ~ PPBC
         )) %>% select(PPBC, death = overall_survival)

newcols = c(ppbc_colors[-5], sp$involution_duration)
names(newcols)[1] = "inv unknown"
names(newcols)[5:6] = paste("inv", names(newcols)[5:6])
#Top column annotation
colTop <- HeatmapAnnotation(df=ann_top, which="col",
                            col = list(PPBC = newcols,
                                       death = sp$overall_survival)
)

```

Bottom column annotation

```{r}
ann_bottom = sampledata[,"PAM50", drop=F]
colBottom <- HeatmapAnnotation(df=ann_bottom, which="col", col = list(PAM50 = pam_colors))

```

Row annotation

```{r}
anno_rows = sig_results %>%
  select(gene_name, Type) %>%
  distinct() %>%
  filter(!duplicated(gene_name)) %>%
  column_to_rownames("gene_name")

#Essential that the order be the same!
anno_rows = anno_rows[match(rownames(mat),rownames(anno_rows)), ,drop=F]

rowAnno = HeatmapAnnotation(df=anno_rows, which="row",
                            col=list(Type = cp$gene_colors)
)

```

Draw heatmap

```{r, fig.height=7, fig.width=10}
hlp = list(title="rowscaled vst")

Heatmap(mat,
        top_annotation = colTop,
        bottom_annotation = colBottom,
        left_annotation = rowAnno,
        heatmap_legend_param = hlp,
        show_row_names = T,
        show_column_names = F,
        cluster_rows = T,
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        column_title = "DEGs: Involution vs rest")
```
Top column annotation, tweaked

```{r}
sampledata = as.data.frame(colData(dds))

#inv_nas = sampledata %>%
#  filter(is.na(involution_duration) & PPBC == "involuting")
#sampledata$involution_duration
ann_top = 
sampledata %>%
  mutate(involution_duration = as.character(involution_duration),
         PPBC = as.character(PPBC),
         overall_survival = as.character(overall_survival)) %>%
  mutate(PPBC.old = PPBC,
         PPBC = case_when(
           PPBC.old == "involuting" & is.na(involution_duration) ~ "inv unknown",
           PPBC.old == "involuting" & involution_duration == "<= 6 months" ~ "inv <= 6 months",
           PPBC.old == "involuting" & involution_duration == "> 6 months" ~ "inv > 6 months",
           TRUE ~ "rest"
         )) %>% select(PPBC, death = overall_survival)

newcols = c(ppbc_colors[-2:-4], sp$involution_duration)
names(newcols)[1] = "inv unknown"
names(newcols)[3:4] = paste("inv", names(newcols)[3:4])
#Top column annotation
colTop <- HeatmapAnnotation(df=ann_top, which="col",
                            col = list(PPBC = newcols,
                                       death = sp$overall_survival)
)

```

```{r}
scales::show_col(newcols)
```


```{r, fig.height=7, fig.width=10}
hlp = list(title="rowscaled vst")

Heatmap(mat,
        top_annotation = colTop,
        bottom_annotation = colBottom,
        left_annotation = rowAnno,
        heatmap_legend_param = hlp,
        show_row_names = T,
        show_column_names = F,
        cluster_rows = T,
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        column_title = "DEGs: Involution vs rest")
```

## With k-means

```{r, fig.height=7, fig.width=10}
set.seed(1)

ann_top = 
sampledata %>%
  mutate(involution_duration = as.character(involution_duration),
         PPBC = as.character(PPBC),
         overall_survival = as.character(overall_survival)) %>%
  mutate(PPBC.old = PPBC,
         PPBC = case_when(
           PPBC.old == "involuting" & is.na(involution_duration) ~ "inv unknown",
           PPBC.old == "involuting" & involution_duration == "<= 6 months" ~ "inv <= 6 months",
           PPBC.old == "involuting" & involution_duration == "> 6 months" ~ "inv > 6 months",
           TRUE ~ PPBC
         )) %>% select(PPBC, death = overall_survival)

newcols = c(ppbc_colors[-5], sp$involution_duration)
names(newcols)[1] = "inv unknown"
names(newcols)[5:6] = paste("inv", names(newcols)[5:6])
#Top column annotation
colTop <- HeatmapAnnotation(df=ann_top, which="col",
                            col = list(PPBC = newcols,
                                       death = sp$overall_survival)
)



Heatmap(mat,
        top_annotation = colTop,
        bottom_annotation = colBottom,
        left_annotation = rowAnno,
        heatmap_legend_param = hlp,
        show_row_names = T,
        show_column_names = F,
        cluster_rows = T,
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        column_title = "DEGs: Involution vs rest",
        column_km=3, row_km = 3)
```


```{r, fig.height=7, fig.width=10}
set.seed(1)

ann_top = 
sampledata %>%
  mutate(involution_duration = as.character(involution_duration),
         PPBC = as.character(PPBC),
         overall_survival = as.character(overall_survival)) %>%
  mutate(PPBC.old = PPBC,
         PPBC = case_when(
           PPBC.old == "involuting" & is.na(involution_duration) ~ "inv unknown",
           PPBC.old == "involuting" & involution_duration == "<= 6 months" ~ "inv <= 6 months",
           PPBC.old == "involuting" & involution_duration == "> 6 months" ~ "inv > 6 months",
           TRUE ~ "rest"
         )) %>% select(PPBC, death = overall_survival)

newcols = c(ppbc_colors[-2:-4], sp$involution_duration)
names(newcols)[1] = "inv unknown"
names(newcols)[3:4] = paste("inv", names(newcols)[3:4])
#Top column annotation
colTop <- HeatmapAnnotation(df=ann_top, which="col",
                            col = list(PPBC = newcols,
                                       death = sp$overall_survival)
)

Heatmap(mat,
        top_annotation = colTop,
        bottom_annotation = colBottom,
        left_annotation = rowAnno,
        heatmap_legend_param = hlp,
        show_row_names = T,
        show_column_names = F,
        cluster_rows = T,
        row_names_gp = gpar(fontsize = 8),
        column_names_gp = gpar(fontsize = 8),
        column_title = "DEGs: Involution vs rest",
        column_km=3, row_km = 3)
```

# Sig test involution duration vs IG clusters

With IG as integer:

```{r}
ig_clusts = readRDS(here("data", "Rds", "11_ig_clusters.Rds"))

sd = sampledata %>%
  mutate(involution_duration = as.character(involution_duration),
         PPBC = as.character(PPBC),
         overall_survival = as.character(overall_survival)) %>%
  mutate(PPBC.old = PPBC,
         PPBC = case_when(
           PPBC.old == "involuting" & is.na(involution_duration) ~ "inv unknown",
           PPBC.old == "involuting" & involution_duration == "<= 6 months" ~ "inv <= 6 months",
           PPBC.old == "involuting" & involution_duration == "> 6 months" ~ "inv > 6 months",
           TRUE ~ "rest"
         )) %>%
  right_join(select(ig_clusts$col_clust, sample_name, IG_fctr = IG_col_cluster, IG_int = col_cluster), .,
             by = "sample_name")

  kruskal.test(x = sd$PPBC, 
             g = sd$IG_int)  
```

With IG as factor:

```{r}
kruskal.test(x = sd$PPBC, 
             g = sd$IG_fctr)  

```

Chi square:

Individual contributions:

```{r}
ct = chisq.test(sd$PPBC, sd$IG_fctr)
  

ct
```


```{r}
ct$stdres
```



```{r}
chisq.posthoc.test::chisq.posthoc.test(table(sd$PPBC, sd$IG_fctr)) %>%
  filter(Value == "p values")
```


