---
title: "QC Checks"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(car)
library(ggpubr)
library(ggrepel)

theme_set(theme_bw())
```

```{r}
filedict <- read_csv(here("data/metadata/spatial/00_file_location_dictionary.csv"))
```

## Summary files

Combine all summary files together for each panel.

MPIF26 = CD3_CD20_CD27_FoxP3_PanCK
MPIF27 = CD3_CD8_CD138_CD20_PanCK

```{r}
read_fn <- function(file, trim_string = here("data/vectra/raw/summary/")) {
  read_csv(file, col_types=cols(
      .default = col_double(),
      `Image Location` = col_character(),
      `Image Tag` = col_character(),
      `Algorithm Name` = col_character(),
      `Start Time` = col_character(),
      `Analysis Region` = col_character(),
      `Analysis Inputs` = col_character())) %>% 
    mutate(filename = str_remove(file, trim_string))
}

summary1 <-
    list.files(here("data/vectra/raw/summary"), 
               full.names = T, pattern = "MPIF26") %>% 
    map_df(~read_fn(.))

summary2 <-
    list.files(here("data/vectra/raw/summary"), 
               full.names = T, pattern = "MPIF27") %>% 
    map_df(~read_fn(.))
```

Markers unique to summary1: CD27 (early activation B and T cells), FOXP3
Markers unique to summary2: CD138 (plasma B), CD8

## Process summaries

`rename_all(make.names)` will result in nonsensical output, we'll need to do this stepwise.

```{r, echo=F}
summary_name_repair <- function(cols){
  cols %>%
    tolower() %>%
    str_replace_all("\\%", "percent") %>%
    str_remove("\\(opal.*\\) ") %>%
    str_remove("\\:") %>%
    str_replace("\\(μm²\\)", "um2") %>%
    str_replace("\\(μm\\)", "um") %>%
    str_replace("\\(mm²\\)", "mm2") %>%
    str_replace_all(" ", "_")
} 

colnames(summary1) <- summary_name_repair(colnames(summary1))
colnames(summary2) <- summary_name_repair(colnames(summary2))
```

Column names after cleaning (it's a lot)

```{r, echo=F}
colnames(summary1)
```

```{r}
#temporary fix, unclear why these columns are getting duplicated
summary1 <- summary1[,colnames(summary1)[-which(duplicated(colnames(summary1)))]]
summary2 <- summary2[,colnames(summary2)[-which(duplicated(colnames(summary2)))]]
```


Clean t-number column, remove unnecessary image locations etc.

```{r}
clean_summary <- function(df){
  df %>%
    mutate(t_number=str_extract(filename, "^T[0-9]+-[0-9]+(_[A-Z][0-9]+)?")) %>%
    relocate(t_number, .before = everything()) %>%
    select(-image_location:-start_time)
} 

summary1 <- clean_summary(summary1)
summary2 <- clean_summary(summary2)

head(summary1)
```

Extract batch information from the file name.

```{r}
extract_batch <- function(df){
  df %>%
    mutate(temp = str_remove(str_remove(filename, "_summary_results.csv"), "_I1|_II1")) %>%
    tidyr::separate(temp, into = c(NA, "panel", "batch"), sep="_") %>%
    relocate(panel, batch, .after=t_number) %>%
    mutate(batch = factor(batch, levels = sort(unique(batch))))
}

summary1 <- extract_batch(summary1)
summary2 <- extract_batch(summary2)
```


## Common fields

Some fields, such as tumor area etc, should be pretty similar in both panels. Show Spearman correlation and R squared. Label samples that are outliers based on the Bonferroni outlier test. See `?car::outlierTest()` for details. (Alternatively, plot the highest Cook's distance?)

## Tumor area

```{r}
plot_common_columns <- function(panel1 = summary1, panel2 = summary2,
                                panel1_name = "MPIF26", panel2_name = "MPIF27",
                                col, returndata = F, returnoutliers = F, ...){
  
  df <- inner_join(
    select(panel1, t_number, all_of(col)),
    select(panel2, t_number, all_of(col)),
    by = "t_number"
  )
  
  colnames(df)[2:3] <- c("x", "y")

  if(returndata){return(df)}
  
  batches <- bind_rows(
    distinct(select(panel1, t_number, batch)),
    distinct(select(panel2, t_number, batch))
  ) %>% distinct()
  stopifnot(nrow(filter(batches, duplicated(t_number)))==0)
  
  #Make the basic ggplot
  g <- df %>%
    left_join(., batches, by = "t_number") %>%
    ggplot(aes(x, y)) +
    #geom_text(aes(label = t_number)) +
    geom_point(aes(color = batch)) +
    ggtitle(paste("Correlation:", str_replace_all(col, "_", " "))) +
    xlab(paste("Panel 1:", panel1_name)) +
    ylab(paste("Panel 2:", panel2_name)) +
    #p value in stat_cor comes from `cor.test`
    ggpubr::stat_cor(method = "spearman",
                     aes(label = paste(..rr.label.., ..r.label.., sep = "~`,`~")))
  
  #Test for outliers
  df <- as.data.frame(df)
  rownames(df) <- df$t_number #necessary for proper labelling
  fit <- lm(y ~ x, data = df)
  outliers <- names(car::outlierTest(fit, cutoff = 0.05, n = Inf)$bonf.p)
  if(returnoutliers){return(outliers)}
  
  #Label outliers, if any
  g +
    ggrepel::geom_label_repel(data = filter(df, t_number %in% outliers),
             aes(label = t_number), ...)
}

plot_common_columns(col = "tumor_area_mm2", returndata = F)
```

## Stroma area

```{r}
plot_common_columns(col = "stroma_area_mm2")
```

## Glass area

```{r}
plot_common_columns(col = "glass_area_mm2")
```

## Total cells

```{r}
plot_common_columns(col = "total_cells")
```

## DAPI positive cells

Looks the same as total cells

```{r}
plot_common_columns(col = "dapi_positive_cells")
```

## PanCK +

```{r}
plot_common_columns(col = "panck_positive_cells")
```

```{r}
plot_common_columns(col = "cd3_positive_cells")
```

```{r}
plot_common_columns(col = "cd20_positive_cells")
```

```{r}
sessionInfo()
```


                         


