---
author: "Tycho Bismeijer & Kat Moore"
date: "`r Sys.Date()`"
params:
  batch: 
    label: "Select batch:"
    value: "batch3"
    input: select
    choices: ["batch1", "batch2", "batch3", "batch4", "batch5", "batch6", "batch7"]
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
requireNamespace("ggplot2")
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r libraries}
library(here)
library(fs)
library(tidyverse)
```

```{r,echo=F}
#params$batch cannot be passed directly to title
#separate yaml blocks are possible, see below
batch <- params$batch
```

---
title: '`r paste('Object QC for:', batch)`'
---

In this notebook, we perform a variety of sample-wise sanity checks regarding marker positivity cutoffs, localization, and correlation.

## Marker intensity by sample

For every sample, check each marker to ensure that it's localized where it's supposed to be, and ensure that the classifier cutoffs look reasonable based on marker intensity. For markers common to both panels, ensure that both the cell counts and the marker intensities are reasonably well correlated between both panels. 

(In the case of cell counts, this SHOULD be the same as the analysis performed on the summary QC files in the previous notebook. But, check to make sure that the reality within the data frame matches the summary.)

Produce separate dfs per panel. This is done by batch.

```{r detect_batch_files}
inDir <- "data/vectra/interim/objects/"
inFiles26 <- list.files(here(inDir), pattern = paste0("MPIF26_", batch, ".nc"))
inFiles26
inFiles27 <- list.files(here(inDir), pattern = paste0("MPIF27_", batch, ".nc"))
inFiles27
```

Load the relevant objects:

```{r read_26}
source(here("src/spatial/read_cells.R"))

start <- Sys.time()
mpif26 <- rlang::set_names(here(file.path(inDir, inFiles26))) %>%
    purrr::map_dfr(~ read_cells(., intensity=T)) %>%
    dplyr::relocate(t_number, panel, .before = everything())
end <- Sys.time()

end-start
```

```{r read_27}
start <- Sys.time()
mpif27 <- rlang::set_names(here(file.path(inDir, inFiles27))) %>%
    purrr::map_dfr(~ read_cells(., intensity=T)) %>%
    dplyr::relocate(t_number, panel, .before = everything())
end <- Sys.time()


end-start
```


## DAPI

All present cells are DAPI+.

```{r}
stopifnot(nrow(filter(mpif26, DAPI_positive == F))==0 &
            nrow(filter(mpif27, DAPI_positive == F))==0)
```

### Localization

DAPI should be nucleus only.

```{r summarize_markers_fn}
summarise_marker <- function(marker, panel1=mpif26, panel2=mpif27, bysample = F){
  pos <- paste0(marker, "_positive")
  nuc <- paste0(marker, "_nucleus_intensity")
  cyt <- paste0(marker, "_cytoplasm_intensity")
  mem <- paste0(marker, "_membrane_intensity")
  
  if(pos %in% colnames(panel1) & pos %in% colnames(panel2)){
    #print("marker in both")
    df <- bind_rows(
      select(panel1, t_number, panel, classifier_label, {{pos}}, {{nuc}}, {{cyt}}, {{mem}}),
      select(panel2, t_number, panel, classifier_label, {{pos}}, {{nuc}}, {{cyt}}, {{mem}})
    )
  } else if (!pos %in% colnames(panel2)) {
    #print("marker in panel1")
    df <- select(panel1, t_number, panel, classifier_label, {{pos}}, {{nuc}}, {{cyt}}, {{mem}})
  } else {
    #print("marker in panel2")
    df <- select(panel2, t_number, panel, classifier_label, {{pos}}, {{nuc}}, {{cyt}}, {{mem}})
  }
  #return(df)
  if(bysample == F){
    df %>%
      group_by(panel, classifier_label, .data[[pos]]) %>%
      summarise(
        count = n(),
        mean_cytoplasm = mean(.data[[cyt]], na.rm=T),
        mean_nuclear = mean(.data[[nuc]], na.rm=T),
        mean_membrane = mean(.data[[mem]], na.rm=T),
        .groups="drop"
        )
  } else {
    df %>%
      group_by(panel,classifier_label, t_number, .data[[pos]]) %>%
      summarise(
        count = n(),
        mean_cytoplasm = mean(.data[[cyt]], na.rm=T),
        mean_nuclear = mean(.data[[nuc]], na.rm=T),
        mean_membrane = mean(.data[[mem]], na.rm=T),
        .groups="drop"
        )
  }
}
```

```{r summarize_dapi}
summarise_marker("DAPI")
```

### Samplewise summary

```{r summarize_dapi_per_sample}
summarise_marker("DAPI", bysample=T) %>%
  select(-mean_cytoplasm, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_nuclear),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Count scatterplot

```{r plot_count}
summarise_marker("DAPI", bysample=T) %>%
  select(-mean_cytoplasm, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_nuclear),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_count, y = MPIF27_count, color = DAPI_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("DAPI count by sample")
  
```

### Average intensity

```{r plot_intensity}
summarise_marker("DAPI", bysample=T) %>%
  select(-mean_cytoplasm, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_nuclear),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_mean_nuclear, y = MPIF27_mean_nuclear, color = DAPI_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("DAPI mean intensity by sample")
  
```

```{r hist_intensity, fig.height=6, include=F, eval=F}
##### Intensity histogram
#Is this useful? Probably not when the marker is present in both panels
mpif26 %>%
  ggplot(aes(x = DAPI_nucleus_intensity, fill = DAPI_positive)) +
  geom_histogram(bins = 30) +
  facet_wrap(~t_number) +
  ggtitle("MPIF26")

mpif27 %>%
  ggplot(aes(x = DAPI_nucleus_intensity, fill = DAPI_positive)) +
  geom_histogram(bins = 30) +
  facet_wrap(~t_number) +
  ggtitle("MPIF27")
```

### Barplot positive cells

```{r barplot_counts}
summarise_marker("DAPI", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(DAPI_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("DAPI positive cells")
```

## PanCK

### Localization

PanCK is cytoplasm-only.

```{r}
summarise_marker(marker = "PanCK")
```

### Samplewise summary

```{r}
summarise_marker("PanCK", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Count dotplot

```{r}
summarise_marker("PanCK", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_count, y = MPIF27_count, color = PanCK_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("PanCK count by sample")
  
```

### Average intensity

```{r}
summarise_marker("PanCK", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_mean_cytoplasm, y = MPIF27_mean_cytoplasm, color = PanCK_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("PanCK mean intensity by sample")
  
```

### Barplot positive cells

```{r}
summarise_marker("PanCK", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(PanCK_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("PanCK positive cells")
```

## CD3

### Localization

CD3 is cytoplasm-only.

```{r}
summarise_marker(marker = "CD3")
```

### Samplewise summary

```{r}
summarise_marker("CD3", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Count dotplot

```{r}
summarise_marker("CD3", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_count, y = MPIF27_count, color = CD3_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("CD3 count by sample")
  
```

### Average intensity

```{r}
summarise_marker("CD3", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_mean_cytoplasm, y = MPIF27_mean_cytoplasm, color = CD3_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("CD3 mean intensity by panel")
  
```

### Barplot positive cells

```{r}
summarise_marker("CD3", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(CD3_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("CD3 positive cells")
```

## CD20

### Localization

CD20 is cytoplasm-only.

```{r}
summarise_marker(marker = "CD20")
```

### Samplewise summary

```{r}
summarise_marker("CD20", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Count dotplot

```{r}
summarise_marker("CD20", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_count, y = MPIF27_count, color = CD20_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("CD20 count by sample")
  
```

### Average intensity

```{r}
summarise_marker("CD20", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_mean_cytoplasm, y = MPIF27_mean_cytoplasm, color = CD20_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("CD20 mean intensity by panel")
  
```

### Barplot positive cells

```{r}
summarise_marker("CD20", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(CD20_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("CD20 positive cells")
```

## CD27

### Localization

CD27 is cytoplasm-only, and is exclusive to MPIF26.

```{r}
summarise_marker(marker = "CD27")
```

### Samplewise summary

```{r}
summarise_marker("CD27", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Barplot positive cells

```{r}
summarise_marker("CD27", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(CD27_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("CD27 positive cells")
```

```{r,include=F,eval=F}
#...go to bed and try again
summarise_marker("CD27", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{.value}"
  ) %>%
  ggplot(aes(x = MPIF26_count, y = MPIF26_mean_cytoplasm, color = CD27_positive)) +
  geom_point() +
  facet_wrap(~classifier_label) +
  ggtitle("CD27 count by sample")
  
```

## CD8

### Localization

CD8 is cytoplasm-only, and is exclusive to MPIF27.

```{r}
summarise_marker(marker = "CD8")
```

### Samplewise summary

```{r}
summarise_marker("CD8", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Barplot positive cells

```{r}
summarise_marker("CD8", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(CD8_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("CD8 positive cells")
```

## FoxP3

### Localization

CD27 is nucleus-only, and is exclusive to MPIF26.

```{r}
summarise_marker(marker = "FoxP3")
```

### Samplewise summary

```{r}
summarise_marker("FoxP3", bysample=T) %>%
  select(-mean_cytoplasm, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_nuclear),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Barplot positive cells

```{r}
summarise_marker("FoxP3", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(FoxP3_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("FoxP3 positive cells")
```

## CD138

### Localization

CD138 is cytoplasm-only, and is exclusive to MPIF27.

```{r}
summarise_marker(marker = "CD138")
```

### Samplewise summary

```{r}
summarise_marker("CD138", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  pivot_wider(names_from = c(panel,classifier_label),
              values_from = c(count, mean_cytoplasm),
              names_glue = "{panel}_{classifier_label}_{.value}"
              )
```

### Barplot positive cells

```{r}
summarise_marker("CD138", bysample=T) %>%
  select(-mean_nuclear, -mean_membrane) %>%
  filter(CD138_positive == T) %>%
  ggplot(aes(x = t_number, y = count, fill = classifier_label)) +
  geom_bar(stat = "identity") +
  ggpubr::rotate_x_text(angle = 45) +
  ggtitle("CD138 positive cells")
```
