---
title: 'Aggregate marker combinations'
author: "Kat Moore and Tycho Bismeijer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(here)
library(tidyverse)
library(ggrepel)
theme_set(theme_bw())
```

## Aggregate counts

```{r}
source(here("src/spatial/aggregate_counts.R"))
sample_aggregate <- sum_cell_counts()

head(sample_aggregate)
```

```{r}
sample_aggregate %>%
  pull(marker_pos) %>% unique() %>% sort()
```

Marker counts across the whole dataset.

```{r}
marker_df <- sample_aggregate %>%
  rename(region = classifier_label) %>%
  group_by(marker_pos, region) %>%
  summarise(n = sum(n), .groups = "drop") %>%
  arrange(desc(n))

marker_region <- marker_df %>%
  pivot_wider(names_from = region, values_from = n, values_fill = 0) %>%
  mutate(Total = Tumor + Stroma) %>%
  arrange(desc(Total)) %>%
  mutate(marker_pos = ifelse(marker_pos == "DAPI+", "DAPI_only", marker_pos)) %>%
  mutate(marker_pos = str_remove(str_remove(marker_pos, "DAPI\\+_"),"_DAPI\\+"))

marker_region
```

```{r}
write_csv(marker_region, here("results/spatial/marker_combos_by_region.csv"))
```


## Problematic marker combinations

We should not observe:

* CD3+ CD20+
* CD8+ CD20+
* FOXP3+ CD20+

In principle, we should also not observe double positivity of PanCK and immune markers, nor CD138 and T cell markers, but these combinations may occur due to the high abundance of these markers, which causes spillover to adjacent cells.

Recall that:

* MPIF26 exclusive markers:  CD27, FoxP3
* MPIF27 exclusive markers: CD138, CD8 
* Shared markers: CD20, CD3, PanCK, DAPI

```{r}
panel_markers <- sample_aggregate %>%
  rename(region = classifier_label) %>%
  group_by(marker_pos, panel, region) %>%
  summarise(n = sum(n), .groups = "drop") %>%
  pivot_wider(names_from = region, values_from = n, values_fill = 0) %>%
  mutate(marker_pos = ifelse(marker_pos == "DAPI+", "DAPI_only", marker_pos)) %>%
  mutate(marker_pos = str_remove(str_remove(marker_pos, "DAPI\\+_"),"_DAPI\\+")) %>%
  mutate(Total = Stroma + Tumor) %>%
  arrange(desc(Total))

panel_markers
```

PanCK/immune marker combinations:

```{r}
panel_markers %>%
  filter(str_detect(marker_pos, "PanCK") & str_detect(marker_pos, "CD"))
```

All possible T-cell combos in MPIF27

```{r}
panel_markers %>%
  filter(str_detect(panel, "MPIF27")) %>%
  filter(str_detect(marker_pos, "CD[83]"))
```

Also address:

* CD8+CD3- (MPIF27)
* CD27+CD3- (MPIF26)
* CD20+_CD3+_CD8+ (MPIF27)
* CD138+_CD8+ (MPIF27)
* CD20+_CD8+

NK cells may sometimes express [CD8](https://journals.asm.org/doi/10.1128/JVI.01420-14) and [CD27](https://pubmed.ncbi.nlm.nih.gov/18326863/).

```{r}
panel_markers %>%
  filter(marker_pos %in% c(
    "CD20+_CD3+",
    "CD8+",
    "CD27+",
    "CD20+_CD3+_CD8+",
    "CD20+_CD8+"
  )) %>%
  mutate(Total = Stroma + Tumor)
```

Additional impossible combinations exist within the data, but in relatively low numbers.

## Missing and rare markers

Show samples with more than a few oddball combinations.

```{r}
plot_abun <- function(df, pt, label_top){
  df %>% ggplot(aes(x = classifier_label, y = n, shape = panel, color = batch)) +
      geom_jitter(width = 0.2, height = 0) +
      ggtitle(pt) +
      ggrepel::geom_label_repel(data = head(df, label_top),show.legend = F,
                                aes(label = t_number))
}
combo_abundance <- function(df = sample_aggregate, mno=NULL, mno2 = NULL, myes = NULL, myes2=NULL,
                            min_count = 2000, label_top = 5, returndata=F){
  
  nsamples <- length(unique(df$t_number))
  #Show a combination where a marker that should be present is missing
  if(!is.null(mno2)){
 
    df <- df %>% filter(
        !str_detect(marker_pos, fixed(mno)) & !str_detect(marker_pos, fixed(mno2)),
        str_detect(marker_pos, fixed(myes)),
        n > min_count) %>%
      group_by(t_number, panel, batch, classifier_label) %>%
      summarise(n = sum(n), .groups = "drop") %>%
      arrange(desc(n)) 
    
    if(returndata){return(df)}
    pt <- paste(paste0(length(unique(df$t_number)), "/", nsamples),
                "samples have",
                paste0(myes,str_replace(mno, "\\+", "-"),str_replace(mno2, "\\+", "-")),
                "count over", min_count)
    plot_abun(df, pt, label_top)
  #Show a combination that indicates coexpression that shouldn't exist
  } else if(!is.null(myes2)) {

    df <- df %>%
      filter(
        str_detect(marker_pos, fixed(myes)),
        str_detect(marker_pos, fixed(myes2)),
        n > min_count
        ) %>%
      group_by(t_number, panel, batch, classifier_label) %>%
      summarise(n = sum(n), .groups = "drop") %>%
      arrange(desc(n)) 
    
    #if(returndata){return(df)}
    pt <- paste(paste0(length(unique(df$t_number)), "/", nsamples),
                "samples have",
                paste0(myes,myes2),
                "count over", min_count)
    
    plot_abun(df, pt, label_top)
  
  #Show a combination where one of two expected markers is missing    
  } else if(!is.null(mno)){
      df <- df %>%
      filter(
        !str_detect(marker_pos, fixed(mno)),
        str_detect(marker_pos, fixed(myes)),
        n > min_count) %>%
      group_by(t_number, panel, batch, classifier_label) %>%
      summarise(n = sum(n), .groups = "drop") %>%
      arrange(desc(n)) 
    
    if(returndata){return(df)}
    pt <- paste(paste0(length(unique(df$t_number)), "/", nsamples),
                "samples have",
                paste0(str_replace(mno, "\\+", "-"),myes),
                "count over", min_count)
    
    plot_abun(df, pt, label_top)
  
  } else {
    stop("Provide a valid marker combo")
  }
}
```


### CD3-CD8+

```{r samples-that-miss-cd3-but-have-cd8}
combo_abundance(mno='CD3+', myes='CD8+')
```

### CD3-FoxP3+

```{r samples-that-miss-cd3-but-have-foxp3}
combo_abundance(mno='CD3+', myes='FoxP3+', min_count = 1000)
```

### CD20+CD138+

```{r cd20poscd138pos-samples}
combo_abundance(myes='CD20+', myes2='CD138+', min_count = 1000)
```

### CD27+CD3-CD20-

```{r}
combo_abundance(myes='CD27+', mno='CD3+', mno2='CD20+', min_count = 1000)
```


```{r}
sessionInfo()
```
