---
title: "Report cell types"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(here)
library(tidyverse)

theme_set(theme_bw())
```

We broadly define the following cell types:

* Cytotoxic T-cell (CD8+, CD3+)
* T-reg (CD3+, FoxP3+)
* Helper T-cell (CD3+, CD8-, FoxP3-)
* B-cell (CD20+, CD138-)
* Plasma B-cell (CD20+, CD138+)
* NK cell (CD8+, CD3- or CD27+, CD3-)
* Epithelial cell (PanCK+ and negative for immune markers)

Here we are assuming that CD3+CD8-FoxP3- are *probably* CD4+, since double-negative T-cells represent a [small (1-5%) populatoin of lymphocytes](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3269300/). One may also argue that not all CD8+ T-cells are by definition cytotoxic, but we lack the data to make such distinctions here.

CD27 is an early immune activation marker and is highly associated with a positive outcome in the PPBC cohort, especially among involuting patients. We therefore consider it as an additional characteristic of immune cells. 

CD138 in combination with CD20 is a plasma B marker, but it is also [expressed by T cells](https://www.biorxiv.org/content/10.1101/2020.03.25.008995v1) and [malignant cells](https://pubmed.ncbi.nlm.nih.gov/14983940/).

## Load data

Select panel specific files.

```{r}
inDir <- "data/vectra/processed/objects/"
inFiles26 <- list.files(here(inDir), pattern = paste0(".*MPIF26.*.Rds"))
#inFiles26
inFiles27 <- list.files(here(inDir), pattern = paste0(".*MPIF27.*.Rds"))
#inFiles27
```

Aggregate cell function

```{r}
agg_cells <- function(filename){
  info <- tibble(filename)
  #handle "T20-62169_I1"  "T20-62169_II1"
  info <- info %>%
    mutate(filename = str_replace(filename, "_I", "-I"))
  #extract sample info from filename
  info <- info %>%
    tidyr::separate(col = filename, into = c("t_number", "panel", "batch"),
                    sep = "_") %>%
    mutate(batch = str_remove(batch, ".Rds"))
  
  bind_cols(info, 
            #Only the necessary columsn for reporting
            readRDS(here(file.path(inDir, filename))) %>%
              select(cell_type, object_id, contains("positive"))
            )
}
```

Load panels separately

```{r, load_26}
start <- Sys.time()
mpif26 <- lapply(inFiles26, agg_cells) %>% bind_rows()
end <- Sys.time()
print(end - start)
```

```{r, load_27}
start <- Sys.time()
mpif27 <- lapply(inFiles27, agg_cells) %>% bind_rows()
end <- Sys.time()
print(end - start)
```


## Cell definitions

FoxP3 positivity in the absence of CD3 positivity can occur. These cells will be classified as T-regs.

```{r summarise-markers, cache=F}
summarise_markers <- function (objects) {
  #Summarise markers as strings for downstream sanity checks
  marker_cols <- colnames(objects)[str_detect(colnames(objects), "_positive")]
  #Except DAPI
  marker_cols <- marker_cols[-which(marker_cols == "DAPI_positive")]
  markers_str <- str_replace(marker_cols, "_positive", "+")
  positive <- as.matrix(objects[, marker_cols])
  objects$markers <- apply(positive, 1, function (pos) {
        paste0(markers_str[pos], collapse="_")
  })
  
  objects %>%
    relocate(markers, .after=panel) %>%
    mutate(marker = ifelse(markers == "", "DAPI_only", markers))
}

mpif26 <- summarise_markers(mpif26)
mpif27 <- summarise_markers(mpif27)
```

MPIF26 exclusive markers:  CD27, FoxP3
MPIF27 exclusive markers: CD138, CD8 
Shared markers: CD20, CD3, PanCK, DAPI

```{r}
mpif26 %>% 
  group_by(cell_type, markers) %>%
  count() %>%
  arrange(desc(n)) %>%
  split(.$cell_type)
```

```{r}
mpif27 %>% 
  group_by(cell_type, markers) %>%
  count() %>%
  arrange(desc(n)) %>%
  split(.$cell_type)
```

```{r}
sessionInfo()
```
