---
title: "Report cell types"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=F}
library(here)
library(tidyverse)

theme_set(theme_bw())
```

We broadly define the following cell types. Parenthetical cell identities are putative. 

CD27 positivity is tracked for downstream analysis given that it was highly associated with positive outcome for involuting patients within the RNA expression data. It may be found on both B and T cells. 

CD138 in combination with CD20 is a plasma B marker, but it is also [expressed by T cells](https://www.biorxiv.org/content/10.1101/2020.03.25.008995v1) and [malignant cells](https://pubmed.ncbi.nlm.nih.gov/14983940/). Due to its ubiquitous expression both on cells and in the ECM, it is not used in cell type calling.

We assume that CD3+CD8-FoxP3- are *probably* CD4+, since double-negative T-cells represent a [small (1-5%) population of lymphocytes](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3269300/).

See `src/spatial/call_cell_types.R` for exact cell definitions.

MPIF26: 

* FoxP3+ (Treg, assumed to be CD3 positive)
* CD3+CD27+FoxP3- (T cell, CD4 or CD8)
* CD3+CD27-FoxP3- (T cell, CD4 or CD8)
* CD20+CD27+ (B cell)
* CD20+CD27- (B cell)
* CD27+CD20-CD3- (possibly NK cells, or B/T with failed CD3/CD20 staining)
* PanCK+ (cancer cell)
* Other

MPIF27:

* CD3+CD8+ (CD8+ T cell, with non-zero CD3 intensity)
* CD3+CD8- (CD4+ T cell)
* CD3-CD8+ (NK cell or T cell with completely failed CD3 staining)
* CD20+ (B cell)
* PanCK+ (cancer cell)
* Other

## Load data

Select panel specific files.

```{r}
inDir <- "data/vectra/processed/objects/"
inFiles26 <- list.files(here(inDir), pattern = paste0(".*MPIF26.*.Rds"))
#inFiles26
inFiles27 <- list.files(here(inDir), pattern = paste0(".*MPIF27.*.Rds"))
#inFiles27
```

Aggregate cell function

```{r}
agg_cells <- function(filename){
  info <- tibble(filename)
  #handle "T20-62169_I1"  "T20-62169_II1"
  info <- info %>%
    mutate(filename = str_replace(filename, "_I", "-I"))
  #extract sample info from filename
  info <- info %>%
    tidyr::separate(col = filename, into = c("t_number", "panel", "batch"),
                    sep = "_") %>%
    mutate(batch = str_remove(batch, ".Rds"))
  
  bind_cols(info, 
            #Only the necessary columsn for reporting
            readRDS(here(file.path(inDir, filename))) %>%
              select(cell_type, object_id, contains("positive"))
            )
}
```

Load panels separately

```{r, load_26}
start <- Sys.time()
mpif26 <- lapply(inFiles26, agg_cells) %>% bind_rows()
end <- Sys.time()
print(end - start)
```

```{r, load_27}
start <- Sys.time()
mpif27 <- lapply(inFiles27, agg_cells) %>% bind_rows()
end <- Sys.time()
print(end - start)
```


## Cell definitions

FoxP3 positivity in the absence of CD3 positivity can occur. These cells will be classified as T-regs.

```{r summarise-markers, cache=F}
summarise_markers <- function (objects) {
  #Summarise markers as strings for downstream sanity checks
  marker_cols <- colnames(objects)[str_detect(colnames(objects), "_positive")]
  #Except DAPI
  marker_cols <- marker_cols[-which(marker_cols == "DAPI_positive")]
  markers_str <- str_replace(marker_cols, "_positive", "+")
  positive <- as.matrix(objects[, marker_cols])
  objects$markers <- apply(positive, 1, function (pos) {
        paste0(markers_str[pos], collapse="_")
  })
  
  objects %>%
    relocate(markers, .after=panel) %>%
    mutate(marker = ifelse(markers == "", "DAPI_only", markers))
}

mpif26 <- summarise_markers(mpif26)
mpif27 <- summarise_markers(mpif27)
```

MPIF26 exclusive markers:  CD27, FoxP3
MPIF27 exclusive markers: CD138, CD8 
Shared markers: CD20, CD3, PanCK, DAPI

### MPIF26 counts

```{r}
mpif26 %>% 
  group_by(cell_type) %>%
  count() %>%
  arrange(desc(n))
```


```{r}
mpif26 %>% 
  group_by(cell_type, markers) %>%
  count() %>%
  arrange(desc(n)) %>%
  split(.$cell_type)
```

### MPIF27 counts

```{r}
mpif27 %>% 
  group_by(cell_type, markers) %>%
  count() %>%
  arrange(desc(n)) %>%
  split(.$cell_type)
```

```{r}
mpif27 %>% 
  group_by(cell_type) %>%
  count() %>%
  arrange(desc(n))
```

```{r}
sessionInfo()
```
