---
title: "Link cell density to outcome"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
```

Prepare cell densities for survival analysis by linking them to PPBC subtypes and clinical outcomes.

## Load data

```{r}
density <- read_tsv(here("results/spatial/density.tsv"), show_col_types = F)
#head(density)

density %>%
  group_by(panel) %>%
  count()
```

Fix classifier labels

```{r}
#density %>% group_by(classifier_label) %>% count()

density <- density %>%
  mutate(classifier_label = ifelse(classifier_label == "tumor", "Tumor",
                                   classifier_label))
```


The metadata is separated into three sheets.

```{r}
read_sheets_to_list <- function(xlsfile){
  readxl::excel_sheets(xlsfile) %>% 
    set_names() %>% 
    map(readxl::read_excel, path = xlsfile)
}

meta <- read_sheets_to_list(here("data/metadata/PPBC_metadata.xlsx"))
```

Here the t_number can be mapped to the patient_ID (via sample_ID).

```{r}
head(meta$sample_data)
```

Survival data can be found here.

```{r}
head(meta$patient_data)
```

Explanations are in the codebook.

```{r}
head(meta$codebook)
```

## Survival outcomes in PPBC

Join t_numbers with survival, link to density.

```{r}
dens <- meta$sample_data %>%
  filter(experimental_platform != "RNAseq" & !is.na(sample_ID)) %>%
  select(t_number = sample_ID, patient_ID) %>% distinct() %>%
  left_join(., meta$patient_data, by = "patient_ID") %>%
  left_join(density, ., by = "t_number")

stopifnot(nrow(filter(dens, is.na(death)))==0)

head(dens)
```

Four possible possible PPBC categories exist : 

* nulliparous breast cancer(non_prbc)
* pregnant breast cancer (prbc)
* post-partum breast cancer: lactating (ppbc_lac)
* post-partum breast cancer: involuting (ppbc_inv)

```{r}
unique(dens$study_group)
```

Format PPBC and survival categories.

```{r}
#Grade is sometimes missing
#dens %>% group_by(study_group, database, grade) %>% count()

dens <- dens %>%
  mutate(classifier_label = factor(classifier_label,
                                   levels = c("Stroma", "Tumor")),
         panel = factor(panel, levels = c("MPIF26", "MPIF27")),
         #Shorter group names, similar to study_group but clearer 
         #and with no potential delimiters
         group = case_when(
           study_group == "ppbc_inv" ~ "inv",
           study_group == "non_prbc" ~ "nonprbc",
           study_group == "ppbc_lac" ~ "lac",
           TRUE ~ study_group) 
         ) %>%
  mutate(group = factor(group, levels = c("nonprbc", "prbc", "lac", "inv"))) %>%
  relocate(group, .after = study_group) %>%
  mutate(stage = factor(stage, levels = c("stage I", "stage II",
                                          "stage III", "stage IV")))

dens %>%
  select(study_group, group) %>%
  distinct()
```

## Write data

```{r}
saveRDS(dens, here("data/vectra/processed/density_ppbc.Rds"))
```


```{r}
sessionInfo()
```

