---
title: "Kruskal-Wallis and box plots of cell densities"
author: "Kat Moore"
date: "`r Sys.Date()`"
params:
  min_cell_count: 20000
  density_outcome: "data/vectra/processed/density_ppbc.Rds"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries, echo=F}
library(tidyverse)
library(here)
library(ggpubr)

theme_set(theme_bw())
```

Pairwise Kruskal-Wallis tests are made for the association between cell type density and PPBC study group. Box plots visualize the relative densities of cell types between groups.

## Overview & setup

Load data.

```{r}
dens <- readRDS(here(params$density_outcome))
head(dens)
```

Perform analyses on total density.

```{r}
dens <- dens %>%
  filter(classifier_label %in% c("Total"))

dens$classifier_label %>% unique()
```

### Cell counts by panel

```{r}
dens %>%
  group_by(cell_type, panel) %>%
  summarise(n = sum(n), .groups="drop") %>%
  pivot_wider(names_from = panel, values_from = n, values_fill = 0) %>%
  arrange(desc(MPIF27+MPIF26))
```

### Cell groups excluded

The minimum number of cells of a given type that must be present across the whole dataset is:

```{r}
min_cell_count <- as.integer(params$min_cell_count)
min_cell_count
```

Cell types which fail this threshold:

```{r}
excluded_cells <- dens %>%
  group_by(cell_type, panel) %>%
  summarise(n = sum(n), .groups="drop") %>%
  filter(n < min_cell_count) %>%
  pull(cell_type)

excluded_cells
```

Remaining cell groups:

```{r}
dens <- dens %>% filter(!cell_type %in% excluded_cells)
unique(dens$cell_type)
```

### Samples per group

Number of samples per study group:

```{r}
dens %>%
  select(t_number, panel, study_group) %>%
  distinct() %>%
  group_by(panel, study_group) %>%
  count() %>%
  pivot_wider(names_from = panel, values_from = n)
```

## Kruskal-wallis tests

The majority of samples come from the nulliparous and involuting groups: the primary contrast of interest. The relatively few number of samples from pregnant and lactating cancer patients can be used exploratively.

```{r}
source(here("src/spatial/kruskal_density.R"))

# Testing
stat_density(df = dens, panel = "MPIF26",
             cell_type = "CD3+FoxP3-", tidied = T)
```

```{r}
relevant_cells <- unique(dens$cell_type)
relevant_cells <- relevant_cells[!relevant_cells %in% c("PanCK+", "Other")]

density_by_ppbc <- lapply(relevant_cells,
                          function(x){multi_dens(dens, x)}) %>%
  bind_rows() %>%
  group_by(panel) %>%
  mutate(kruskal_bh = p.adjust(kruskal_p), .after = kruskal_p) %>%
  arrange(kruskal_bh, kruskal_p)

density_by_ppbc

```

### Sig results

Show significant comparisons (FDR 0.1) based on multiple testing correction. Only look at pairwise Wilcoxon results if the Kruskal-Wallis test is significant.

```{r}
sigdens <- density_by_ppbc %>%
  select(-kruskal_p) %>%
  filter(kruskal_bh < 0.1) %>%
  pivot_longer(cols = kruskal_bh:ppbcpw_vs_ppbcdl_bh, names_to = "comparison",
               values_to = "padjust") %>%
  filter(padjust < 0.1) %>%
  arrange(padjust)

sigdens
```

## Beehive plots

The p values in these plots do not include multiple testing correction. Kruskal-Wallis for all PPBC groups alongside pairwise Wilcoxon tests are shown. The y axis is square-root-transformed.

### Overall survival

Color param:

```{r}
cb <- "death"
```

#### MPIF26

```{r}
cells_26 <- unique(filter(dens, panel == "MPIF26")$cell_type)
cells_26 <- cells_26[!cells_26 %in% c("PanCK+", "Other")]

lapply(cells_26,
       function(x) {
         plot_density(panel = "MPIF26", cell_type = x, colorby = cb) +
           scale_y_sqrt() +
           scale_color_viridis_d()
       })
```

#### MPIF27

```{r}
cells_27 <- unique(filter(dens, panel == "MPIF27")$cell_type)
cells_27 <- cells_27[!cells_27 %in% c("PanCK+", "Other")]

lapply(cells_27,
       function(x) {
         plot_density(panel = "MPIF27", cell_type = x, colorby = cb) +
           scale_y_sqrt() + scale_color_viridis_d()
       })
```

### Distant recurrence

Color param:

```{r}
cb <- "distant_recurrence"
```

#### MPIF26

```{r}
lapply(cells_26,
       function(x) {
         plot_density(panel = "MPIF26", cell_type = x, colorby = cb) +
           scale_y_sqrt() +
           scale_color_viridis_d()
       })
```

#### MPIF27

```{r}
lapply(cells_27,
       function(x) {
         plot_density(panel = "MPIF27", cell_type = x, colorby = cb) +
           scale_y_sqrt() + scale_color_viridis_d()
       })
```

### Clinical subtype

Color param:

```{r}
cb <- "clin_subtype"
```

#### MPIF26

```{r}
lapply(cells_26,
       function(x) {
         plot_density(panel = "MPIF26", cell_type = x, colorby = cb) +
           scale_y_sqrt() +
           scale_color_viridis_d()
       })
```

#### MPIF27

```{r}
lapply(cells_27,
       function(x) {
         plot_density(panel = "MPIF27", cell_type = x, colorby = cb) +
           scale_y_sqrt() + scale_color_viridis_d()
       })
```

```{r}
sessionInfo()
```
