---
title: "Involution duration and immune density"
author: "Kat Moore"
date: "`r Sys.Date()`"
params:
  min_cell_count:
    value: 20000
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

Explore the relationship between involution or breastfeeding duration and immune cell density within PPBC-inv patients.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries, echo=F}
library(tidyverse)
library(here)
library(ggpubr)
library(broom)

theme_set(theme_bw())
```

Pairwise Kruskal-Wallis tests are made for the association between cell type density and PPBC study group. Box plots visualize the relative densities of cell types between groups.

## Set up

### Select regions

Load data.

```{r}
dens <- readRDS(here("data/vectra/processed/density_ppbc.Rds"))
head(dens)
```

Examine total density only.

```{r}
dens <- dens %>%
  filter(classifier_label %in% c("Total"))

dens$classifier_label %>% unique()
```

### Min cell threshold

The minimum number of cells of a given type that must be present across the whole dataset is:

```{r}
min_cell_count <- as.integer(params$min_cell_count)
min_cell_count
```

Cell types which fail this threshold:

```{r}
excluded_cells <- dens %>%
  group_by(cell_type, panel) %>%
  summarise(n = sum(n), .groups="drop") %>%
  filter(n < min_cell_count) %>%
  pull(cell_type)

excluded_cells
```

Remaining cell groups:

```{r}
dens <- dens %>% filter(!cell_type %in% excluded_cells)
unique(dens$cell_type)
```

### Subset involution samples

Samples remaining by panel

```{r}
inv_dens <- dens %>%
  filter(group == "inv") %>%
  #This subset leaves only numeric values and NAs
  mutate(months_involution = as.numeric(months_involution),
         months_breastfeeding = as.numeric(months_breastfeeding))

inv_dens %>%
  select(t_number, panel, study_group, group) %>%
  distinct() %>%
  group_by(panel, group) %>%
  count() %>%
  pivot_wider(names_from = panel, values_from = n)
```

## Months involution

### Dot plot

A few samples have missing values for duration involution or breastfeeding.

```{r}
duration_density_plot <- function(df, time, cell, colorby){
  
  df <- filter(df, cell_type == cell) %>%
    mutate(death = factor(death, levels = c(0,1))) %>%
    mutate(distant_recurrence = factor(distant_recurrence, levels = c(0,1)))
  
  time_title <- str_to_sentence(str_replace(time, "_", " "))
  cell_title <- paste(cell, "density")
  colortitle <- str_to_sentence(str_replace(colorby, "_", " "))
  
  df %>%
    ggplot(aes(x = get(time), y = density)) +
    geom_point(aes(color = get(colorby))) +
    geom_smooth(method = lm, se = FALSE) +
    facet_wrap(~classifier_label) +
    xlab(time_title) + ylab(cell_title) + labs(color = colortitle) +
    ggtitle(paste(time_title, "and", cell_title)) +
    ggpubr::stat_cor(method = "spearman",
                     aes(label = paste(..rr.label.., ..r.label.., sep = "~`,`~")))
  
}

cells_to_plot <- unique(inv_dens$cell_type)
cells_to_plot <- cells_to_plot[!cells_to_plot %in% c("Other", "PanCK+")]

lapply(cells_to_plot, function(x){
  duration_density_plot(inv_dens, time = "months_involution",
                      cell = x, colorby = "clin_subtype")
  })
```

### Linear regression

How well do the regression coefficients line up in the whole dataset?

```{r}
reg_time_dens <- function(df, time, cell){
   df <- filter(df, cell_type == cell) %>%
     select(t_number, panel, time = !!time, cell_type, density)
   #return(df)
   fit <- glm(time ~ density, family = "poisson", data = df)
   broom::tidy(fit) %>%
     filter(term != "(Intercept)") %>%
     select(-term) %>%
     mutate(cell_density = cell, predictor = time, .before = everything()) %>%
     relocate(p.value, .after = predictor)
}

lapply(cells_to_plot, function(x){
  reg_time_dens(inv_dens, time = "months_involution", cell = x)}) %>%
  bind_rows() %>%
  mutate(fdr = p.adjust(p.value, method = "fdr"), .before = p.value) %>%
  arrange(fdr) %>%
  mutate(sig = fdr < 0.05, .before = fdr)
```

## Months breastfeeding

### Dot plot

```{r}
lapply(cells_to_plot, function(x){
  duration_density_plot(inv_dens, time = "months_breastfeeding",
                      cell = x, colorby = "clin_subtype") +
    scale_x_log10()
  })
```

### Linear regression

```{r}
lapply(cells_to_plot, function(x){
  reg_time_dens(inv_dens, time = "months_breastfeeding", cell = x)}) %>%
  bind_rows() %>%
  mutate(fdr = p.adjust(p.value, method = "fdr"), .before = p.value) %>%
  arrange(fdr) %>%
  mutate(sig = fdr < 0.05, .before = fdr)
```

## Session info

```{r}
sessionInfo()
```
