---
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
params:
  outcome:
    label: "Survival outcome type"
    value: "OS"
    input: "select"
    choice: ["OS", "DRS"]
  min_cell_count: 20000
  density_outcome: "data/vectra/processed/density_ppbc.Rds"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, libraries}
library(tidyverse)
library(here)
library(ggpubr)
library(survival)
library(survminer)
library(broom)

theme_set(theme_bw())
```

This notebook reports on the relationship between cell densities and the following clinical outcome:

```{r}
outcome <- params$outcome
ptitle <- ifelse(outcome == "OS", "Overall survival", "Distant recurrence")
print(ptitle)
```

---
title: '`r paste('Cox regression with total density:', ptitle)`'
---

## Setup

### Select regions

Load data

```{r}
dens <- readRDS(here(params$density_outcome))
```

Use total density for association with outcome.

```{r}
dens <- dens %>%
  filter(classifier_label %in% c("Total"))

dens$classifier_label %>% unique()
```

### Cell count threshold

Counts per panel at start:

```{r}
dens %>%
  group_by(cell_type, panel) %>%
  summarise(n = sum(n), .groups="drop") %>%
  pivot_wider(names_from = panel, values_from = n, values_fill = 0) %>%
  arrange(desc(MPIF27+MPIF26))
```

### Cell count threshold

The minimum number of cells of a given type that must be present across the whole dataset is:

```{r}
min_cell_count <- as.integer(params$min_cell_count)
min_cell_count
```

Cell types which fail this threshold:

```{r}
excluded_cells <- dens %>%
  group_by(cell_type, panel) %>%
  summarise(n = sum(n), .groups="drop") %>%
  filter(n < min_cell_count) %>%
  pull(cell_type)

excluded_cells
```

Remaining cell groups:

```{r}
dens <- dens %>% filter(!cell_type %in% excluded_cells)
unique(dens$cell_type)
```

### Sample count threshold

Number of samples per panel:

```{r}
dens %>%
  select(t_number, panel, study_group) %>%
  distinct() %>%
  group_by(panel, study_group) %>%
  count() %>%
  pivot_wider(names_from = panel, values_from = n)
```

Per stage:

```{r}
dens %>%
  select(t_number, stage, study_group) %>%
  distinct() %>%
  group_by(stage, study_group) %>%
  count() %>%
  pivot_wider(names_from = stage, values_from = n, values_fill = 0)
```

Not enough samples to include prbc and ppbcdl, or stage IV samples.

```{r}
cox_dens <- dens %>%
  filter(!study_group %in% c("prbc", "ppbcdl")) %>%
  filter(stage != "stage IV")

cox_dens <- droplevels(cox_dens)

cox_dens %>%
  select(t_number, stage, study_group) %>%
  distinct() %>%
  group_by(stage, study_group) %>%
  count() %>%
  pivot_wider(names_from = stage, values_from = n)
```

### Clinical outcomes

Detect the appropriate time and clinical outcomes for Cox regressions.

```{r}
if(params$outcome == "OS"){
  timevar <- "time_OS_months"
  eventvar <- "death"
} else if (params$outcome == "DRS"){
  timevar <- "time_DRS_months"
  eventvar <- "distant_recurrence"
} else {
  stop("Outcome must be OS or DRS")
}

if(!timevar %in% colnames(cox_dens)){
  stop(paste("Time variable", timevar,"not found in input data"))
} else if (!eventvar %in% colnames(cox_dens)){
  stop(paste("Event variable", eventvar,"not found in input data"))
}

print(paste0(ptitle, " (", outcome, ")"))
print(paste("Time variable:", timevar))
print(paste("Event variable:", eventvar))
```

Set up the left side of the Cox formula.

```{r}
survres <- paste0("Surv(time = ", timevar,", event = ", eventvar, ")")
survres
```

### Cox functions

```{r}
source(here("src/spatial/cox_spatial.R"))
```

## Cox: all samples

Overall impact of density independent on involution and nulliparous groups combined.

Fit separate models for each cell_type/tumor vs stroma/panel. P value is calculated by a likelihood ratio test (LRT) that compares the fit of the Cox model with the fit of a reduced model that does not contain the covariate of interest. Hazard ratios and confidence intervals are obtained from the covariate of interest in the full model.

### Relevant cells

We are only interested in the association of putative immune cells and clinical outcome.

```{r}
relevant_cells <- unique(cox_dens$cell_type)
relevant_cells <- relevant_cells[!relevant_cells %in% c("PanCK+", "Other")]
relevant_cells
```

### Univariate

Use the intercept as the reduced formula for univarate analysis

```{r}
univ_form <- paste0(survres, " ~ 1")
univ_form
```

Example Cox output for single cell type.

```{r}
cox_density(cox_dens, "MPIF27", "CD3+CD8+",
            f = univ_form, tidied = T)
```

#### Results

Loop over all cell types, while auto-detecting which panel to use.

```{r}
univ_cox <- lapply(relevant_cells, function(x){
  all_cox_density(cox_dens, x,
                  f = univ_form)
}) %>%
  bind_rows() %>%
  mutate(lrt.fdr = p.adjust(lrt.pval, method = "BH"), .after = panel) %>%
  mutate(formula = paste0(univ_form, " + Density"))

univ_cox %>%
  arrange(lrt.fdr, lrt.pval)
```

### Multivariate

Start with a limited number of covariates: 

study_group
stage

PAM50 is not suitable, since that will be correlated with immune cell density. Same is likely true of treatment interventions. Grade cannot be used because it is missing for some samples.

Stratify PPBC to allow separate baseline hazards functions for each PPBC subgroup.

```{r}
multiv_form <- paste0(survres," ~ strata(study_group) + stage")
multiv_form
```

Example one cell type:

```{r}
#The LRT pval is the same for all the covariates because 
#it's a comparison between the full and reduced model
#However, coefficient estimates should be different
cox_density(cox_dens, "MPIF27", "CD3+CD8+",
            f = multiv_form,
            tidied=F)

cox_density(cox_dens, "MPIF27", "CD3+CD8+",
            f = multiv_form,
            tidied=T)

```

#### Results 

```{r}
mult_cox <- lapply(relevant_cells, function(x){
  all_cox_density(cox_dens, x,
                  f = multiv_form)
}) %>%
  bind_rows() %>%
  mutate(lrt.fdr = p.adjust(lrt.pval, method = "BH"), .after = panel) %>%
  mutate(formula = paste0(multiv_form, " + density"))

mult_cox %>%
  arrange(lrt.fdr, lrt.pval)
```

## Cox: inv and np subgroups

Check significance in involution samples and nulliparous samples separately.

Involution samples:

```{r}
inv_dens <- filter(cox_dens, study_group == "ppbcpw")
inv_dens <- droplevels(inv_dens)

inv_dens %>%
  select(t_number, stage, study_group) %>%
  distinct() %>%
  group_by(study_group, stage) %>%
  distinct() %>%
  count()
```

Nulliparous samples:

```{r}
np_dens <- filter(cox_dens, study_group == "npbc")
np_dens <- droplevels(np_dens)

np_dens %>%
  select(t_number, stage, study_group) %>%
  distinct() %>%
  group_by(study_group, stage) %>%
  distinct() %>%
  count()
```

### Univariate

As above, but split by group. Formula:

```{r}
univ_form
```

Example single cell type.

```{r}
cox_density(df = inv_dens, panel = "MPIF26", cell_type = "CD20+",
            f = univ_form, tidied=F)
```

#### Involution

```{r}
inv_univ_cox <- lapply(relevant_cells, function(x){
  all_cox_density(inv_dens, x,
                  f = univ_form)
}) %>%
  bind_rows() %>%
  mutate(lrt.fdr = p.adjust(lrt.pval, method = "BH"), .after = panel) %>%
  mutate(formula = paste0(univ_form, " + Density")) %>%
  mutate(subset = "inv", .before = everything())

inv_univ_cox %>%
  arrange(lrt.fdr, lrt.pval)
```

#### Nulliparous

All cell density results by region and panel, arranged by FDR and P value.

```{r}
np_univ_cox <- lapply(relevant_cells, function(x){
  all_cox_density(np_dens, x,
                  f = univ_form)
}) %>%
  bind_rows() %>%
  mutate(lrt.fdr = p.adjust(lrt.pval, method = "BH"), .after = panel) %>%
  mutate(formula = paste0(" + Density"))

np_univ_cox %>%
  arrange(lrt.fdr, lrt.pval)
```

### Multivariate

Given that we are analysing PPBC groups separately, we can omit `strata_group` from this formula.

```{r}
multiv_subform <- paste0(survres, " ~ stage")
multiv_subform
```

#### Involution

All cell density results by region and panel, arranged by FDR and P value.

```{r}
inv_mult_cox <- lapply(relevant_cells, function(x){
  all_cox_density(inv_dens, x,
                  f = multiv_subform)
}) %>%
  bind_rows() %>%
  mutate(lrt.fdr = p.adjust(lrt.pval, method = "BH"), .after = panel) %>%
  mutate(formula = paste0(multiv_subform, " + density"))

inv_mult_cox %>%
  arrange(lrt.fdr, lrt.pval)
```

#### Nulliparous

All cell density results by region and panel, arranged by FDR and P value.

```{r}
np_mult_cox <- lapply(relevant_cells, function(x){
  all_cox_density(np_dens, x,
                  f = multiv_subform)
}) %>%
  bind_rows() %>%
  mutate(lrt.fdr = p.adjust(lrt.pval, method = "BH"), .after = panel) %>%
  mutate(formula = paste0(multiv_subform, " + density"))

np_mult_cox %>%
  arrange(lrt.fdr, lrt.pval)
```

## Kaplan-Meier: Density by study group

KMs require categorical variables. Convert density into quantiles. P values are from logrank tests.

MPIF26

```{r}
cells_26 <- unique(filter(cox_dens, panel == "MPIF26")$cell_type)
cells_26 <- cells_26[!cells_26 %in% c("PanCK+", "Other")]

lapply(cells_26, function(x){
  km_plot(cox_dens, panel = "MPIF26", cell_type = x, outcome,
        f = paste0(survres, " ~ quantile"))
}) %>% set_names(cells_26)
```

MPIF27

```{r}
cells_27 <- unique(filter(cox_dens, panel == "MPIF27")$cell_type)
cells_27 <- cells_27[!cells_27 %in% c("PanCK+", "Other")]

lapply(cells_27, function(x){
  km_plot(cox_dens, panel = "MPIF27", cell_type = x, outcome,
        f =  paste0(survres, " ~ quantile"))
}) %>% set_names(cells_27)
```

## Cox: interaction

Interaction term between group and cell density. Aims to answer the question: which cell densities affect outcome *differently* in involuting vs nulliparous patients. Maybe be statistically underpowered due to lack of more samples.

```{r}
cox_dens %>%
  select(t_number, study_group) %>%
  distinct() %>%
  group_by(study_group) %>%
  count()
```

The analysis below works very similarly to the one implemented with `cox_density()`, except that a formula with an interaction term is automatically generated.

The reduced formula will be:

```{r}
inv_baseform <- paste(multiv_subform, "+ study_group + Total_density")
inv_baseform
```

The interaction term formula is:

```{r}
inv_intform <- paste(inv_baseform, "+ study_group:Total_density")
inv_intform
```

`study_group` is a factor that refers to whether a sample is involuting or nulliparous, and `Total_density` refers to the density of a given cell type. An LRT attempts to determine whether model fit using as formula including the interaction term `study_group:Total_density` is different than the base formula `~stage + study_group + Total_density`. 

Again, this test is likely statistically underpowered. Unfortunately it is not simple to do a formal power calculation between nonbinary predictors in an interaction term.

```{r}
cox_interaction(cox_dens, "MPIF27", "CD20+",
                baseform = inv_baseform, tidied = F)
```

### Results by significance

Also the wrapper function is similar to `all_cox_density()`, except that it invokes `cox_interaction` instead of `cox_density`.

```{r}
interaction_df <- lapply(relevant_cells, function(x){
  all_cox_interaction(cox_dens, x, baseform = inv_baseform)
}) %>%
 bind_rows() %>%
  #Only show the involution-related interaction terms
  filter(str_detect(term, "ppbcpw")) %>%
  mutate(lrt.fdr = p.adjust(lrt.pval, method = "BH"), .after = panel) %>%
  mutate(formula = inv_intform)

```

All cell density results by panel, arranged by FDR and P value.

```{r}
interaction_df %>%
  arrange(lrt.fdr, lrt.pval)
```

## Session info

```{r}
sessionInfo()
```
