---
title: "Tissue Segmentation"
author: "Tycho Bismeijer"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r preliminaries, include=F}
library(conflicted)

ggplot2::theme_set(ggplot2::theme_bw())
```


This report shows details of one sample to explore the segmentation process by density.

```{r libraries, message=F}
library(here)
library(spatstat)
library(tidyverse)
library(yaml)
```


We use the same functions and parameters that are used for segmenting all samples.

```{r source-tissue-segmentation}
source(here('src/spatial/segment_tissue_density.R'))
source(here('src/spatial/plot_segmentation.R'))
```

```{r config}
samples <- c(
    'T21-60303_MPIF26_batch2',
    'T21-60303_MPIF27_batch2')
config <- yaml.load_file(here('src/spatial/segment_tissue_density_config.yaml'))
print(config)
```

```{r load-data}
cells <- map(set_names(samples), function (s) {
    readRDS(paste0(here('data/vectra/processed/objects/'), s, '.Rds'))
})
```

# Points #

A plot of all the cell points (bounding box centers) per cell type and the window.

```{r plot_points}
for (sample in samples) {
  cells_pp <- as_pp_cells(cells[[sample]])
  for (cell_type in levels(cells_pp$marks)) {
    plot(subset(cells_pp, marks == cell_type), chars=1, main=paste0(sample, ' / ', cell_type))
  }
}
```

# Density #

Based on the point locations of the cells, the density of the different tissue compartments are
estimated and plotted. 

```{r estimate_densities}
cell_density <- map(cells, function (cells) {
  cells_pp <- as_pp_cells(cells)
  estimate_densities(cells_pp, config$tissues, config$resolution)
})
```

```{r plot_densities}
for (sample in samples) {
  for (tissue in names(config$tissues)) {
    plot(cell_density[[sample]][[tissue]], main=paste0(sample, ' / ', tissue))
  }
}
```

# Segmentation #

```{r segment}
segmentation <- map(cell_density, function (tissue_density) {
  segment_densities(tissue_density, config$tissues, config$minimum_density)
})

```

```{r plot-segmentation}
for (sample in samples) {
    print(plot_segmentation(segmentation[[sample]], title=sample))
}
```
