---
title: "IG clusters and CD20"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo = F}
library(here)
library(tidyverse)
library(ggpubr)
library(broom)
theme_set(theme_bw())
```

Check to see whether the low, medium and high IG clusters from differential expression correspond to CD20 marker density.

## Load data

Contains marker density info and sample names:

```{r}
dens <- readRDS(here("data/vectra/processed/density_ppbc.Rds"))
```

We have regional density for stroma and tumor. Also calculate total density.

```{r}
cd20 <- dens %>%
  filter(cell_type == "CD20+") %>%
  rename(region=classifier_label) %>%
  mutate(region = factor(region, levels = c("Stroma", "Tumor", "Total"))) %>%
  arrange(t_number, region) #%>%
  #select(-distant_recurrence, time_DRS_months) %>%
  #left_join(., distinct(select(dens, sample_name, death, distant_recurrence)),
            #by = "sample_name")

#DRS is NA in a few samples
head(cd20)
```


Contains IG cluster identities and sample names:

```{r}
igclust <- readxl::read_excel(here("results/clustering/11_inv_clusters.xlsx"),
                              sheet="col_clusters") %>%
  mutate(IG_col_cluster = factor(IG_col_cluster,
                                 levels = c("low", "medium", "high")))
```

Combine

```{r}
cd20 <- inner_join(cd20,select(igclust, -study_group), by = "sample_name")
```

Response variables should be factors for plotting.

```{r}
cd20 <- cd20 %>%
  mutate(death = factor(death, levels = c(0, 1)),
         distant_recurrence = factor(distant_recurrence, levels = c(0, 1)))
```


```{r}
head(cd20)
```

## Tumor and stroma size

Average area tumor vs stroma:

```{r}
dens %>%
  filter(classifier_label != "Total") %>%
  group_by(t_number, panel, classifier_label) %>%
  summarise(av_area = mean(area), .groups="drop") %>%
  ggplot(aes(x = classifier_label, y = av_area)) +
  geom_jitter(aes(color = panel), height = 0 ) +
  geom_boxplot(alpha = 0) +
  ggtitle("Average area of tumor and stroma by panel") +
  ggpubr::stat_compare_means(
    label = "p.signif",
    comparisons = list(
      c("Stroma", "Tumor")
    )) +
  scale_y_sqrt()
```

## Box plot

In all study groups:

```{r}
box_cd20 <- function(df, colorby = "panel"){
  ggplot(df, aes(x = IG_col_cluster, y = density)) +
  geom_jitter(height = 0, aes(color = get(colorby))) +
  geom_boxplot(alpha = 0.1, outlier.shape = NA) +
  facet_wrap(~region) +
  scale_y_sqrt() +
  xlab("Immunoglobulin expression cluster derived from RNAseq") +
  ylab("CD20 Vectra density") +
  ggpubr::stat_compare_means(
    label = "p.signif",
    comparisons = list(
      c("low", "medium"), c("medium", "high"), c("low", "high")
    )) +
    labs(color = colorby)
}

box_cd20(cd20) +
  ggtitle("Relationship between IG cluster status and CD20 density by sample and panel")
```

Study groups separately for OS:

```{r}
lapply(unique(cd20$study_group), function(x){
  box_cd20(filter(cd20, study_group == x), colorby = "death") +
    ggtitle(paste("Relationship between IG cluster status and CD20 density in", x)) +
    scale_color_viridis_d()
})
```

Study groups separately for DRS:

```{r}
lapply(unique(cd20$study_group), function(x){
  box_cd20(filter(cd20, study_group == x), colorby = "distant_recurrence") +
    ggtitle(paste("Relationship between IG cluster status and CD20 density in", x)) +
    scale_color_viridis_d()
})
```

## Linear regression

How well do the regression coefficients line up in the whole dataset?

```{r}
lapply(unique(cd20$region), function(x){
  sub <- filter(cd20, region == x)
  #return(unique(sub$IG_col_cluster))
  fitcd20 <- glm(col_cluster ~ density, family = "poisson", data = sub)
  #summary(fitcd20)
  broom::tidy(fitcd20)
}) %>% set_names(unique(cd20$region)) %>%
  bind_rows(.id = "region") %>%
  filter(term != "(Intercept)") %>%
  #This is the only density we're looking at
  mutate(term = "CD20 density") %>%
  mutate(sig = ifelse(p.value < 0.05, "yes", "no"), .after = region)
  
```

And if we facet by study group?

```{r}
lapply(unique(cd20$study_group), function(y){
  df <- filter(cd20, study_group == y)
  lapply(unique(df$region), function(x){
    sub <- filter(df, region == x)
    fitcd20 <- glm(col_cluster ~ density, family = "poisson", data = sub)
    broom::tidy(fitcd20)
  }) %>% set_names(unique(df$region)) %>%
    bind_rows(.id = "region") %>%
    filter(term != "(Intercept)") %>%
    #This is the only density we're looking at
    mutate(term = "CD20 density") %>%
    mutate(sig = ifelse(p.value < 0.05, "yes", "no"), .after = region)
}) %>% set_names(unique(cd20$study_group)) %>%
  bind_rows(.id = "study_group") %>%
  relocate(p.value, .after=sig)
```

```{r}
sessionInfo()
```


