---
title: "Spatstat"
author: "Tycho Bismeijer & Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    df_print: paged
    highlight: kate
params:
  metadata: "data/external/PPBC_metadata_20220811.xlsx"
  density: "data/vectra/processed/density_ppbc.Rds"
  lib: "src/spatial/outcome.R"
---

```{r preliminaries, include=F}
library(conflicted)

ggplot2::theme_set(ggplot2::theme_bw())
```


```{r libraries, message=F}
library(broom)
library(here)
library(spatstat)
library(survival)
library(survminer)
library(tidyverse)
library(yaml)

source(here(params$lib))
```

```{r load-data-spatstat}
read_spatstat <- function(fn, cross) {
  col_spec <- list( 
    measure = col_character(),
    model = col_character(),
    null_estimate = col_double(),
    observed_estimate = col_double(),
    radius = col_double(),
    delta_estimate = col_double())
  if (cross) {
    col_spec[['cell_type1']] = col_character()
    col_spec[['cell_type2']] = col_character()
  } else {
    col_spec[['cell_type']] = col_character()
  }
  col_spec <- do.call(cols_only, col_spec)
  df <- list.files(fn, '*.tsv', full.names=T) %>%
    set_names() %>%
    map_dfr(read_tsv, col_types = col_spec, .id = 'filename') %>%
    mutate(
      measure = factor(measure),
      model = factor(model)) %>%
    tidyr::extract(filename, c('sample_ID', '_slide', 'experimental_platform', 'batch_HALO'),
                   '(T[0-9]+-[0-9]+(_I+[0-9]+)?)_(MPIF[0-9]+)_(batch[0-9]+)') %>%
      select(-`_slide`)
  if (cross) {
    df <- mutate(df, cell_type1 = factor(cell_type1), cell_type2 = factor(cell_type2))
  } else {
    df <- mutate(df, cell_type = factor(cell_type))
  }
  df
}
sample_id_vars <- c('sample_ID', 'experimental_platform', 'batch_HALO')

spatstat_l <- read_spatstat(here('results/spatial/l'), cross = F)
spatstat_lcross_panck <- read_spatstat(here('results/spatial/lcross_panck'), cross = T) %>%
  mutate(measure = "Lcross_PanCK") %>%
  dplyr::filter(!is.na(cell_type2)) #8 rows have NA for all stats
spatstat_lcross_immune <- read_spatstat(here('results/spatial/lcross_immune'), cross = T) %>%
  mutate(measure = "Lcross_immune")  %>%
  # Current implementation has redundancy with L_cross-PanCK, where PanCK is always cell type 1
  # We are also not interested in Other for this comparison
  dplyr::filter(!cell_type1 %in% c("PanCK+", "Other")) %>%
  # Keep PanCK+ in cell_type2 because different start points may yield different results
  dplyr::filter(!cell_type2 %in% c("Other")) %>%
  dplyr::filter(!is.na(cell_type2)) #62 NA rows

spatstat <- bind_rows(spatstat_l, spatstat_lcross_panck, spatstat_lcross_immune) %>%
  mutate(cell_types = coalesce(cell_type, paste0(cell_type1, "/", cell_type2)))
```

```{r load-data-outcome}
outcome <- read_outcome(here(params$metadata))
```

Samples with outcome but no spatstat:
```{r missing-a}
print(anti_join(outcome, spatstat_l,
                by = c('sample_ID', 'batch_HALO', 'experimental_platform')) %>%
        mutate(stat_missing = "l", .before = everything())) 
print(anti_join(outcome, spatstat_lcross_panck,
                by = c('sample_ID', 'batch_HALO', 'experimental_platform')) %>%
  mutate(stat_missing = "lcross_panck", .before = everything())) 
print(anti_join(outcome, spatstat_lcross_immune,
                by = c('sample_ID', 'batch_HALO', 'experimental_platform')) %>%
  mutate(stat_missing = "lcross_immune", .before = everything())) 
```

Samples with spatstat but no outcome:

```{r missing-b}
print(anti_join(spatstat, outcome,
                by = c('sample_ID', 'batch_HALO', 'experimental_platform')))
```

```{r filter-patients}
outcome <- dplyr::filter(outcome,
  study_group %in% c("npbc", "ppbcpw"),
  stage != "stage IV")
```

## Overview

### L over radius

```{r l-plot}
for (ct in levels(spatstat$cell_type)) {
  plt <- spatstat %>%
    dplyr::filter(measure == 'L', cell_type == ct) %>%
    ggplot(aes(x = radius, y = sample_ID)) +
    geom_tile(aes(fill = observed_estimate)) +
    scale_x_continuous('Radius (mm)') +
    scale_y_discrete('Sample') +
    ggtitle(paste0('L estimate of ', ct))
  print(plt)
}
```

### Lcross panCK+ over radius

Crosses:

```{r}
spatstat %>%
  dplyr::filter(measure == 'Lcross_PanCK') %>%
  select(cell_type1, cell_type2) %>% distinct()
```


```{r lcross-panck-plot}
for (ct in unique(spatstat_lcross_panck$cell_type2)) {
  plt <- spatstat %>%
    dplyr::filter(measure == 'Lcross_PanCK', cell_type1 == 'PanCK+', cell_type2 == ct) %>%
    ggplot(aes(x = radius, y = sample_ID)) +
    geom_tile(aes(fill = observed_estimate)) +
    scale_x_continuous('Radius (mm)') +
    scale_y_discrete('Sample') +
    ggtitle(paste0('Lcross estimate of ', ct, ' to panCK+'))
  print(plt)
}
```

### Lcross immune over radius

Available crosses. x -> y and y -> x comparisons exist, since the starting point differs.

```{r}
immune_crosses <- spatstat %>%
  dplyr::filter(measure == 'Lcross_immune') %>%
  select(cell_type1, cell_type2) %>% 
  dplyr::filter(!cell_type1 %in% c("PanCK+", "Other")) %>%
  dplyr::filter(!cell_type2 %in% c("PanCK+", "Other")) %>%
  distinct()

immune_crosses
```

```{r lcross-immune-plot}
for (i in 1:nrow(immune_crosses)) {
  ct1 <- immune_crosses$cell_type1[i]
  ct2 <- immune_crosses$cell_type2[i]
  plt <- spatstat %>%
    dplyr::filter(measure == 'Lcross_immune', cell_type1 == ct1, cell_type2 == ct2) %>%
    ggplot(aes(x = radius, y = sample_ID)) +
    geom_tile(aes(fill = observed_estimate)) +
    scale_x_continuous('Radius (mm)') +
    scale_y_discrete('Sample') +
    ggtitle(paste0('Lcross estimate of ', ct1, ' to ', ct2))
  print(plt)
}
```

### Histogram at radius 0.03 mm

```{r, de_hist}
inner_join(outcome, spatstat, by = sample_id_vars) %>%
  dplyr::filter(
    radius == 0.03,
    !is.na(death),
    !is.na(time_OS_months),
    !is.na(delta_estimate)) %>%
  group_by(cell_types, measure, model, radius) %>%
  group_walk(function (df_group, key) {
    plt <- ggplot(df_group, aes(x = delta_estimate)) +
      geom_histogram(bins = 30) +
      ggtitle(paste0(names(key), '=', map_chr(key, as.character), collapse = ' '))
    print(plt)
  })
```

## Comparison between groups

```{r compare-groups, message=F}
non_par_test <- function(x, y) {
  valid_obs <- is.finite(x) & is.finite(y)
  x <- x[valid_obs]
  y <- factor(y[valid_obs])
  if (min(table(y) < 2) | length(levels(y)) != 2) {
    str(y)
    tibble()
  } else {
    tidy(wilcox.test(x ~ y))
  }
}


inner_join(outcome, spatstat, by = sample_id_vars) %>%
  dplyr::filter(radius == 0.03) %>%
  group_by(cell_types, measure, model, radius) %>%
  summarize(non_par_test(delta_estimate, study_group)) %>%
  arrange(p.value)
```

## Cox-models

```{r fit-cox}
fit_cox <- function(survival, predictor) {
  model <- coxph(survival ~ predictor)
  model_summary <- summary(model)
  tibble(
    loglikelihood = model$loglik[2],
    n_event = model$nevent,
    statistic_lr = model_summary$logtest[1],
    p_lr = model_summary$logtest[3],
    statistic_score = model_summary$sctest[1],
    statistic_p = model_summary$sctest[3],
    statistic_wald = model_summary$waldtest[1],
    p_wald = model_summary$waldtest[3],
    nominal_p = model_summary$coefficients[1, 5],
    hr = model_summary$conf.int[1, 1],
    hr_min = model_summary$conf.int[1, 3],
    hr_max = model_summary$conf.int[1, 4])
}

cox_results <- inner_join(outcome, spatstat, by = sample_id_vars) %>%
  dplyr::filter(radius < 0.1) %>%
  group_by(cell_types, measure, model, radius) %>%
  summarize(
    fit_cox(Surv(time = time_OS_months, event = death), delta_estimate),
    n = n(),
    n_samples = n_distinct(sample_ID),
    n_data = sum(!is.na(delta_estimate)),
    .groups = 'drop')
```

```{r}
cox_results <- cox_results %>%
  dplyr::filter(!is.na(nominal_p)) %>%
  group_by(cell_types, measure, model) %>%
  mutate(fdr_per_ct = p.adjust(nominal_p, 'fdr')) %>%
  ungroup() %>%
  mutate(fdr = fdr_per_ct / n_distinct(cell_types, measure, model))
```

```{r}
cox_results %>%
  group_by(cell_types, measure, model) %>%
  arrange(-fdr) %>%
  relocate(fdr, .before = everything())
```

Min FDR, average HR, and overall impact (good/bad) across all radii. Note: impact may be different at different distances. See graphs below.

```{r}
cox_results %>%
  group_by(cell_types, measure, model) %>%
  summarize(min_fdr = min(fdr), ave_hr = signif(mean(hr_min, na.rm=T),2)) %>%
  arrange(min_fdr) %>%
  mutate(impact = ifelse(ave_hr > 1, "harmful", "beneficial"))
```

## HR by radius and cell type

Split by metric for legibility.

```{r, fig.width=10}
for(metric in unique(cox_results$measure)){
  plt <- cox_results[cox_results$measure == metric,] %>%
    ggplot(aes(x=radius, y=hr, group='')) +
    geom_ribbon(aes(ymin=hr_min, ymax=hr_max), fill = 'grey70') +
    geom_line(aes(color=!is.na(fdr) & fdr < 0.05)) +
    facet_wrap(~cell_types) +
    scale_y_continuous(trans = scales::log_trans(base=2)) + 
    geom_hline(yintercept=1.0)
  print(plt +
          ggtitle(metric) +
          coord_cartesian(ylim = c(1/10, 10), xlim = c(0, 0.1)))
}


```

CD20-specific graphs.

```{r, fig.height = 4, fig.width=7}
plt <- cox_results %>%
  dplyr::filter(cell_types %in% c("CD20+","PanCK+/CD20+", "CD20+/PanCK+")) %>%
  mutate(cell_types = paste(cell_types, measure)) %>%
  ggplot(aes(x=radius, y=hr, group='')) +
  geom_ribbon(aes(ymin=hr_min, ymax=hr_max), fill = 'grey70') +
  geom_line(aes(color=!is.na(fdr) & fdr < 0.05)) +
  facet_grid(cols = vars(cell_types)) +
  scale_y_continuous(trans = scales::log_trans(base=2)) + 
  geom_hline(yintercept=1.0)
print(plt +
  coord_cartesian(ylim = c(1/10, 10), xlim = c(0, 0.1)))
```

## Kaplan-Meier

Difficult to first generate kaplan-meier estimates before plotting because of limitations in
survfit and ggsurvplot, resultign in ggsurvplot needing the original data.

```{r}
inner_join(outcome, spatstat, by = sample_id_vars) %>%
  dplyr::filter(
    radius == 0.03,
    !is.na(death),
    !is.na(time_OS_months),
    !is.na(delta_estimate)) %>%
  #Cox filter, sig only
  dplyr::filter(cell_types %in% dplyr::filter(cox_results, radius == 0.03, fdr < 0.05)$cell_types) %>%
  group_by(cell_types, measure, model, radius) %>%
  group_walk(function (df_group, key) {
    df_group$de_split <- df_group$delta_estimate > median(df_group$delta_estimate, na.rm=T)
    fit = survfit(Surv(time = time_OS_months, event = death) ~ de_split, data = df_group)
    plt <- ggsurvplot(fit, data = df_group, risk.table = T, conf.int = T, pval = T, pval.method = T) +
      ggtitle(paste0(names(key), '=', map_chr(key, as.character), collapse = ' '))
    print(plt)
  })
```

Look at some CD20 graphs specifically.

```{r}
inner_join(outcome, spatstat, by = sample_id_vars) %>%
  dplyr::filter(
    radius == 0.03,
    !is.na(death),
    !is.na(time_OS_months),
    !is.na(delta_estimate)) %>%
  #CD20-PanCK filter
  dplyr::filter(cell_types %in% c("CD20+","PanCK+/CD20+", "CD20+/PanCK+")) %>%
  group_by(cell_types, measure, model, radius) %>%
  group_walk(function (df_group, key) {
    df_group$de_split <- df_group$delta_estimate > median(df_group$delta_estimate, na.rm=T)
    fit = survfit(Surv(time = time_OS_months, event = death) ~ de_split, data = df_group)
    plt <- ggsurvplot(fit, data = df_group, risk.table = T, conf.int = T, pval = T, pval.method = T) +
      ggtitle(paste0(names(key), '=', map_chr(key, as.character), collapse = ' '))
    print(plt)
  })
```

A different result is seen with a higher radius, as observed by the histogram above.

```{r, fig.width=8}
inner_join(outcome, spatstat, by = sample_id_vars) %>%
  dplyr::filter(
    radius == 0.055,
    !is.na(death),
    !is.na(time_OS_months),
    !is.na(delta_estimate)) %>%
  dplyr::filter(cell_types %in% c("CD20+","PanCK+/CD20+", "CD20+/PanCK+")) %>%
  group_by(cell_types, measure, model, radius) %>%
  group_walk(function (df_group, key) {
    df_group$de_split <- df_group$delta_estimate > median(df_group$delta_estimate, na.rm=T)
    fit = survfit(Surv(time = time_OS_months, event = death) ~ de_split, data = df_group)
    plt <- ggsurvplot(fit, data = df_group, risk.table = T, conf.int = T, pval = T, pval.method = T) +
      ggtitle(paste0(names(key), '=', map_chr(key, as.character), collapse = ' '))
    print(plt)
  })
```

## L vs density

```{r}
dens <- readRDS(here(params$density))
```

Combine second order spatial stats with density stats.

```{r}
spatdens <- inner_join(outcome, spatstat, by = sample_id_vars) %>%
  left_join(., dplyr::select(dplyr::filter(dens, classifier_label == "Total"),
                              sample_ID = t_number, density, cell_types = cell_type),
             by = c("sample_ID", "cell_types"))
```

Plot calculate correlation between cell type density and the L statistic over varius radii.

```{r}
plot_l_dens <- function(df){
  
  df <- df %>%
    select(sample_ID, measure, delta_estimate, density, cell_type, radius, study_group) %>%
    dplyr::filter(radius != 0) %>%
    dplyr::filter(!is.na(delta_estimate) & !is.na(density))
  
  df <- df %>%
    group_by(measure, cell_type, radius, study_group) %>%
    summarise(cor_L_dens = cor(delta_estimate, density,
                       use = "everything", method = "spearman"),
              .groups = "drop")
  
  df %>%
    dplyr::filter(!is.na(cor_L_dens)) %>%
    ggplot(aes(x = radius, y = cor_L_dens, color = study_group)) +
    geom_line() +
    facet_wrap(~cell_type) +
    ggtitle("Correlation between the L statistic and density over distance") +
    ylab("Spearman cor (L, density)")
  
}

plot_l_dens(spatdens)
```

For L cross, we first combine cell type densities.

```{r}
spatLcrossdens <- inner_join(outcome, spatstat, by = sample_id_vars) %>%
  dplyr::filter(!is.na(cell_type1)) %>%
  select(sample_ID, measure, delta_estimate, radius, study_group,
         cell_type1, cell_type2, cell_types) %>%
  # Join densities for celltypes one and two separately
  left_join(., dplyr::select(dplyr::filter(dens, classifier_label == "Total"),
                              sample_ID = t_number, density_1 = density, cell_type1 = cell_type),
             by = c("sample_ID", "cell_type1")) %>% relocate(density_1, .after = cell_type1) %>%
  left_join(., dplyr::select(dplyr::filter(dens, classifier_label == "Total"),
                              sample_ID = t_number, density_2 = density, cell_type2 = cell_type),
             by = c("sample_ID", "cell_type2")) %>% relocate(density_2, .after = cell_type2) %>%
  mutate(density = density_1 + density_2) %>%
  select(-cell_type1:-density_2)
```


```{r, fig.height=8}
plot_lcross_dens <- function(df){
  
  df <- df %>%
    select(sample_ID, measure, delta_estimate, density, cell_types, radius, study_group) %>%
    dplyr::filter(radius != 0) %>%
    dplyr::filter(!is.na(delta_estimate) & !is.na(density))
  
  df <- df %>%
    group_by(measure, cell_types, radius, study_group) %>%
    summarise(cor_Lcross_dens = cor(delta_estimate, density,
                       use = "everything", method = "spearman"),
              .groups = "drop")
  
  df %>%
    dplyr::filter(!is.na(cor_Lcross_dens)) %>%
    ggplot(aes(x = radius, y = cor_Lcross_dens, color = study_group)) +
    geom_line() +
    facet_wrap(~cell_types, ncol = 4) +
    ggtitle("Correlation between the Lcross statistic and density over distance") +
    ylab("Spearman cor (Lcross, density)")
  
}

plot_lcross_dens(spatLcrossdens)
```

## Save data

```{r}
saveRDS(spatstat, here("data/vectra/processed/11_spatstat.Rds"))
saveRDS(spatstat_l, here("data/vectra/processed/11_spatstat_l.Rds"))
saveRDS(spatstat_lcross_immune,
        here("data/vectra/processed/11_spatstat_lcross_immune.Rds"))
saveRDS(spatstat_lcross_panck,
        here("data/vectra/processed/11_spatstat_lcross_panck.Rds"))
```

## Session info

```{r}
sessionInfo()
```
