---
title: "Stainings: CD38, TILs, and TAPC"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: yes
    df_print: tibble
    highlight: kate
params:
    figuredata: "data/figures/00_figuredata.Rds"
    study_km: "src/figures/study_km.R"
    TAPC_PPBC_cor: "figures/Fig2h_TAPC_PPBC_cor.pdf"
    km_TAPC_OS: "figures/Fig2i_TAPC_KM_OS.pdf"
    boxplot_CD38: "figures/Fig2k_CD38_boxplot.pdf"
    km_CD38_OS: "figures/Fig2l_CD38_KM_OS.pdf"
    boxplot_TIL: "figures/Fig4a_TIL_boxplot.pdf"
    km_TAPC_DRS: "figures/supfigs/Supfig13c_TAPC_KM_DRS.pdf"
    km_CD38_DRS: "figures/supfigs/Supfig13d_CD38_KM_DRS.pdf"
---

## Dependencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(ggpubr)
library(survival)
library(survminer)
library(ggrepel)
library(ggcorrplot)
library(conflicted)

conflict_prefer("filter", "dplyr")

theme_set(theme_bw())
```

```{r}
source(here("src/figures/faceted_risktable.R"))
source(here("src/figures/study_km.R"))
```

## Load data

```{r}
figuredata <- readRDS(here(params$figuredata))

head(figuredata)
```

## Colors

```{r}
study_colors <- c("NP-BC"="#521262", "Pr-BC"="#AA96DA", "PP-BCdl"="#112D4E", "PP-BCpw"="#46CDCF")
```

## Figure 2H - Relationship TAPC and study group

### Current version

```{r}
TAPC_PPBC_cor <- figuredata %>%
  filter(!is.na(TAPC_score)) %>%
  ggplot(aes(x = as.integer(study_group), y = TAPC_score)) +
  geom_jitter(aes(color = study_group)) + geom_smooth(method = "lm") +
  #p value in stat_cor comes from `cor.test`
  ggpubr::stat_cor(method = "kendall",
                   aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~"))) +
  scale_color_manual(values = study_colors) +
  scale_x_continuous(breaks = seq(1,4,1), labels= levels(figuredata$study_group)) +
  xlab("Study group") + ylab("TAPC score")

TAPC_PPBC_cor
```

### Save output

```{r}
figureDir = file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$TAPC_PPBC_cor), width = 10, height = 7)
TAPC_PPBC_cor
dev.off()
```

## Fig 2i - KM TAPC OS

Due to power issues, omit lac. High/low is around the median density.

```{r}
km_TAPC_OS <- km(
  figuredata, covariate = "TAPC", event_type = "OS",
  study_groups = c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_OS_months, event=death) ~ study_group + TAPC"),
  returndata = F, returnfit = F
)

km_TAPC_OS
```

### Risk table

```{r}
rt_km_TAPC_OS <-faceted_risktable(km_TAPC_OS, faceted.by = "TAPC") +
  ggtitle("Risk table for TAPC OS")
rt_km_TAPC_OS
```

### Save output

```{r}
pdf(file = here(params$km_TAPC_OS), width = 12, height = 7)
km_TAPC_OS
rt_km_TAPC_OS
dev.off()
```

## Fig 2K - Boxplot CD38

Using intratumoral CD38.

```{r}
boxplot_CD38 <- figuredata %>%
  rename(CD38 = TI_CD38_total) %>%
  filter(!is.na(CD38)) %>%
  ggplot(aes(x = factor(study_group,
                        levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
             y = CD38, color = study_group)) +
  geom_point(position = position_jitter(height = 0), shape = 15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  xlab("Study group") + 
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "NP-BC"))) +
  stat_compare_means(label.y = 65) +
  ggtitle("CD38+ density by study group")

boxplot_CD38
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD38), width = 12, height = 7)
boxplot_CD38
dev.off()
```

## Fig 2L - KM CD38 OS

Due to power issues, omit lac and prbc. High/low is around the median density.

```{r}
km_CD38_OS <- km(
  rename(figuredata, CD38 = TI_CD38), covariate = "CD38", event_type = "OS",
  study_groups = c("PP-BCpw", "NP-BC"),            
  surv_formula=as.formula("Surv(time=time_OS_months, event=death) ~ study_group + CD38"),
  returndata = F, returnfit = F
)

km_CD38_OS
```

### Risk table

```{r}
rt_km_CD38_OS <- faceted_risktable(km_CD38_OS, faceted.by = "CD38")
rt_km_CD38_OS
```

### Save output

```{r}
pdf(file = here(params$km_CD38_OS), width = 12, height = 7)
km_CD38_OS
rt_km_CD38_OS
dev.off()
```

## Fig 4A - TIL% boxplot

Include lac for visualization but exclude it from statistical tests.

```{r}
boxplot_TIL <- figuredata %>%
  filter(!is.na(TILpercent)) %>%
  ggplot(aes(x = study_group, y = TILpercent)) +
  geom_jitter(aes(color = study_group), height=0, width = 0.15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  stat_compare_means(method = "wilcox.test", comparisons = list(
    c("PP-BCpw", "Pr-BC"),
    c("PP-BCpw", "NP-BC")
  )) +
  stat_compare_means(label.y = 100) +
  xlab("Study group") + ylab("TIL %") +
  labs(color = "Study group")

boxplot_TIL
```

### Save output

```{r}
pdf(file = here(params$boxplot_TIL), width = 12, height = 7)
boxplot_TIL
dev.off()
```

## Session info

```{r}
sessionInfo()
```
