---
title: "Stainings: CD38, TILs, and TAPC"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    df_print: tibble
    highlight: kate
params:
    figuredata: "data/figures/00_figuredata.Rds"
    study_km: "src/figures/study_km.R"
    TAPC_PPBC_cor: "figures/Fig2h_TAPC_PPBC_cor.pdf"
    km_TAPC_OS: "figures/Fig2i_TAPC_KM_OS.pdf"
    boxplot_CD38: "figures/Fig2k_CD38_boxplot.pdf"
    km_CD38_OS: "figures/Fig2l_CD38_KM_OS.pdf"
    boxplot_TIL: "figures/Fig4a_TIL_boxplot.pdf"
    km_TAPC_DRS: "figures/supfigs/Supfig13c_TAPC_KM_DRS.pdf"
    km_CD38_DRS: "figures/supfigs/Supfig13d_CD38_KM_DRS.pdf"
    TAPC_CD38_cor: "figures/supfigs/Supfig16_TAPC_CD38_cor.pdf"
    TAPC_iso_cor: "figures/supfigs/Supfig19_TAPC_isotype_cor.pdf"
    reviewer_requests: "figures/reviewer_requests.pdf"
---

Temporary switch to html output to handle LaTeX compling error.

## Dependencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(ggpubr)
library(survival)
library(survminer)
library(ggrepel)
library(ggcorrplot)
library(conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
theme_set(theme_bw())
```

```{r}
source(here("src/figures/faceted_risktable.R"))
source(here("src/figures/study_km.R"))
```

## Load data

```{r}
figuredata <- readRDS(here(params$figuredata))

head(figuredata)
```

## Colors

```{r}
study_colors <- c("NP-BC"="#521262", "Pr-BC"="#AA96DA", "PP-BCdl"="#112D4E", "PP-BCpw"="#46CDCF")
```

## Figure 2H - Relationship TAPC and study group

### Current version

```{r}
TAPC_PPBC_cor <- figuredata %>%
  filter(!is.na(TAPC_score)) %>%
  ggplot(aes(x = as.integer(study_group), y = TAPC_score)) +
  geom_jitter(aes(color = study_group)) + geom_smooth(method = "lm") +
  #p value in stat_cor comes from `cor.test`
  ggpubr::stat_cor(method = "kendall",
                   aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~"))) +
  scale_color_manual(values = study_colors) +
  scale_x_continuous(breaks = seq(1,4,1), labels= levels(figuredata$study_group)) +
  xlab("Study group") + ylab("TAPC score")

TAPC_PPBC_cor
```

By reviewer request

```{r}
figuredata %>% 
  filter(!is.na(TAPC_score)) %>%
  ggplot(aes(x = study_group, y = TAPC_score)) +
  geom_jitter(aes(color = study_group)) +
  #geom_boxplot(alpha = 0) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "NP-BC"),
                                        c("PP-BCpw", "Pr-BC"),
                                        c("PP-BCpw", "PP-BCdl")
                                        )) +
  ggtitle("Wilcoxon test between TAPC scores of PPBC study group") +
  scale_color_manual(values = study_colors) +
  xlab("Study group") + ylab("TAPC score") + labs(color = "Study group")
```


### Save output

```{r}
figureDir = file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$TAPC_PPBC_cor), width = 10, height = 7)
TAPC_PPBC_cor
dev.off()
```

## Fig 2i - KM TAPC OS

Due to power issues, omit lac. High/low is around the median density.

```{r}
km_TAPC_OS <- km(
  figuredata, covariate = "TAPC", event_type = "OS",
  study_groups = c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_OS_months, event=death) ~ study_group + TAPC"),
  returndata = F, returnfit = F
)

km_TAPC_OS
```

### Risk table

```{r}
rt_km_TAPC_OS <-faceted_risktable(km_TAPC_OS, faceted.by = "TAPC") +
  ggtitle("Risk table for TAPC OS")
rt_km_TAPC_OS
```

### Save output

```{r}
pdf(file = here(params$km_TAPC_OS), width = 12, height = 7)
km_TAPC_OS
rt_km_TAPC_OS
dev.off()
```

## Fig 2K - Boxplot CD38

Using intratumoral CD38.

```{r}
boxplot_CD38 <- figuredata %>%
  rename(CD38 = TI_CD38_total) %>%
  filter(!is.na(CD38)) %>%
  ggplot(aes(x = factor(study_group,
                        levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
             y = CD38, color = study_group)) +
  geom_point(position = position_jitter(height = 0), shape = 15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  xlab("Study group") + 
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "NP-BC"))) +
  stat_compare_means(label.y = 65) +
  ggtitle("CD38+ density by study group")

boxplot_CD38
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD38), width = 12, height = 7)
boxplot_CD38
dev.off()
```

## Fig 2L - KM CD38 OS

Due to power issues, omit lac and prbc. High/low is around the median density.

```{r}
km_CD38_OS <- km(
  rename(figuredata, CD38 = TI_CD38), covariate = "CD38", event_type = "OS",
  study_groups = c("PP-BCpw", "NP-BC"),            
  surv_formula=as.formula("Surv(time=time_OS_months, event=death) ~ study_group + CD38"),
  returndata = F, returnfit = F
)

km_CD38_OS
```

### Risk table

```{r}
rt_km_CD38_OS <- faceted_risktable(km_CD38_OS, faceted.by = "CD38")
rt_km_CD38_OS
```

### Save output

```{r}
pdf(file = here(params$km_CD38_OS), width = 12, height = 7)
km_CD38_OS
rt_km_CD38_OS
dev.off()
```

## Fig 4A - TIL% boxplot

Include lac for visualization but exclude it from statistical tests.

```{r}
boxplot_TIL <- figuredata %>%
  filter(!is.na(TILpercent)) %>%
  ggplot(aes(x = study_group, y = TILpercent)) +
  geom_jitter(aes(color = study_group), height=0, width = 0.15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  stat_compare_means(method = "wilcox.test", comparisons = list(
    c("PP-BCpw", "Pr-BC"),
    c("PP-BCpw", "NP-BC")
  )) +
  stat_compare_means(label.y = 100) +
  xlab("Study group") + ylab("TIL %") +
  labs(color = "Study group")

boxplot_TIL
```

### Save output

```{r}
pdf(file = here(params$boxplot_TIL), width = 12, height = 7)
boxplot_TIL
dev.off()
```

## Sup Fig 13C- KM TAPC DRS

```{r}
km_TAPC_DRS <- km(
  figuredata, covariate = "TAPC", event_type = "DRS",
  study_groups = c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + TAPC"),
  returndata = F, returnfit = F
)

km_TAPC_DRS
```

### Risk table

```{r}
rt_km_TAPC_DRS <- faceted_risktable(km_TAPC_DRS, faceted.by = "TAPC") +
  ggtitle("Risk table for TAPC DRS")
rt_km_TAPC_DRS
```

### Save output

```{r}
pdf(file = here(params$km_TAPC_DRS), width = 12, height = 7)
km_TAPC_DRS
rt_km_TAPC_DRS
dev.off()
```    
    
## Sup Fig 13D- KM CD38 DRS

```{r}
km_CD38_DRS <- km(
  rename(figuredata, CD38 = TI_CD38), covariate = "CD38", event_type = "DRS",
  study_groups = c("PP-BCpw", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + CD38"),
  returndata = F, returnfit = F
)

km_CD38_DRS
```

### Risk table

```{r}
rt_km_CD38_DRS <-faceted_risktable(km_CD38_DRS, faceted.by = "CD38") +
  ggtitle("Risk table for CD38 DRS")
rt_km_CD38_DRS
```

### Save output

```{r}
pdf(file = here(params$km_CD38_DRS), width = 12, height = 7)
km_CD38_DRS
rt_km_CD38_DRS
dev.off()
```

## Sup Fig 16- Correlation TAPC and CD38

```{r}
cor_plot <- function(df=figuredata, x, y, colorby, met="kendall", ...){
  
  df <- df[!is.na(df[,x]) & !is.na(df[,y]),]

  df %>%
    ggplot(aes(get(x), get(y))) +
    geom_jitter(aes(color = get(colorby)), height = 1, width = 0.2, shape = 15) +
    geom_smooth(method = "lm", formula = y ~ x) +
    ggpubr::stat_cor(method = met, #label.y = 3.5, label.x = -0.2,
                     aes(label = paste(..rr.label.., ..r.label.., ..p.label..,
                                       sep = "~`,`~"))) +
    xlab(str_replace_all(x, "_", " ")) +
    ylab(str_replace_all(y, "_", " ")) +
    labs(color = str_replace_all(colorby, "_", " "))
}

TAPC_CD38_cor_all <- cor_plot(x="TAPC_score", y="TI_CD38_total", colorby = "study_group") +
  scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
  ggtitle("Kendall correlation between TAPC and CD38 in all samples") +
  ylab("% CD38 positive intratumoral cells")
TAPC_CD38_cor_all
```

```{r}
TAPC_CD38_cor_inv <- cor_plot(
  df = filter(figuredata, study_group == "PP-BCpw"),
  x="TAPC_score", y="TI_CD38_total", colorby = "study_group"
) +
  scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
  ggtitle("Kendall correlation between TAPC and CD38 in PP-BCpw") +
  ylab("% CD38 positive intratumoral cells")

TAPC_CD38_cor_inv
```

### Save output

```{r}
pdf(file = here(params$TAPC_CD38_cor), width = 12, height = 7)
TAPC_CD38_cor_all
TAPC_CD38_cor_inv
dev.off()
```

## Sup fig 19 - Correlation TAPC and antibody isotypes

```{r}
TAPC_Ig_cor_all <- lapply(c("IgA", "IgG"), function(ig){
  cor_plot(figuredata, x="TAPC_score", y=ig, colorby = "study_group") +
    scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
    ggtitle(paste("Kendall correlation between TAPC and", ig, "in all samples"))
}) %>% set_names(c("IgA", "IgG"))

TAPC_Ig_cor_all
```

```{r}
TAPC_Ig_cor_group <- lapply(c("IgA", "IgG"), function(ig){
  cor_plot(figuredata, x="TAPC_score", y=ig, colorby = "study_group") +
    scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
    ggtitle(paste("Kendall correlation between TAPC and", ig, "by study group")) +
    facet_wrap(~study_group)
}) %>% set_names(c("IgA", "IgG"))

TAPC_Ig_cor_group
```

### Save output

```{r}
pdf(file = here(params$TAPC_iso_cor), width = 12, height = 7)
TAPC_Ig_cor_all
TAPC_Ig_cor_group
dev.off()
```

## Reviewer request: CD38 and Ig mRNA

### CD38 high/low categories

```{r}
igs <- c("blue", "red","darkgreen","purple")
names(igs) <- c("IgA", "IgG", "IgM", "IgD")

tmpdf <- figuredata %>%
  rename(CD38 = TI_CD38) %>%
  filter(!is.na(CD38)) %>%
  select(CD38, IgA, IgM, IgG, IgD) %>%
  pivot_longer(cols = starts_with("Ig"), names_to = "isotype", values_to = "expression") %>%
  filter(!is.na(expression)) 

boxplot_ig_CD38 <- lapply(names(igs), function(x){
  tmpdf %>%
    filter(isotype == !!x) %>%
    ggplot(aes(x = CD38, y = expression, color = isotype)) +
    geom_point(position = position_jitter(height = 0), shape = 15) +
    geom_boxplot(alpha = 0) +
    xlab("CD38 category") + 
    stat_compare_means(method = "wilcox.test",
                       comparisons = list(c("low", "high"))) +
    #stat_compare_means() +
    ggtitle(paste("CD38+ and", x, "expression")) +
    scale_color_manual(values = igs)
  })

boxplot_ig_CD38
```

### CD38 numeric correlation

```{r}
CD38_Ig_cor_all <- lapply(c(names(igs)), function(ig){
  cor_plot(figuredata, x= "TI_CD38_total", y=ig, colorby = "study_group",
           met = "spearman") +
    scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
    ggtitle(paste("Spearman correlation between CD38 and", ig, "in all samples"))
}) %>% set_names(igs)

CD38_Ig_cor_all
```

```{r}
CD38_Ig_cor_group <- lapply(names(igs), function(ig){
  cor_plot(figuredata, x="TI_CD38_total", y=ig, colorby = "study_group",
           met = "spearman") +
    scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
    ggtitle(paste("Spearman correlation between CD38 and", ig, "by study group")) +
    facet_wrap(~study_group)
}) %>% set_names(names(igs))

CD38_Ig_cor_group
```

## Reviewer request: TILs and TAPCs

```{r}
TIL_TAPC_cor_all <- 
  figuredata %>% filter(!is.na(TILpercent)) %>%
  cor_plot(figuredata, x= "TAPC_score", y="TILpercent", colorby = "study_group") +
    scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
    ggtitle("Kendall correlation between TAPCs and TILs in all samples")

TIL_TAPC_cor_all
```

```{r}
TIL_TAPC_cor_group <- 
  figuredata %>% filter(!is.na(TILpercent)) %>%
  cor_plot(figuredata, x= "TAPC_score", y="TILpercent", colorby = "study_group") +
  scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
  ggtitle("Kendall correlation between TAPCs and TILs per study group") +
  facet_wrap(~study_group)

TIL_TAPC_cor_group
```

### Save output

```{r}
pdf(file = here(params$reviewer_requests), width = 12, height = 7)
CD38_Ig_cor_all
CD38_Ig_cor_group
TIL_TAPC_cor_all
TIL_TAPC_cor_group
#boxplot_ig_CD38 #is this one useful?
dev.off()
```

## Session info

```{r}
sessionInfo()
```
