---
title: "Cibersort Figures"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: yes
    df_print: tibble
    highlight: kate
params:
    figuredata: "data/figures/00_figuredata.Rds"
    boxplot_cd8: "figures/Fig4b_cibersort_CD8_boxplot.pdf"
    km_cd8: "figures/Fig4c_cibersort_CD8_km.pdf"
    boxplot_b_subtypes: "figures/supfigs/Supfig12b_boxplot_cibersort_B_subtypes.pdf"
    km_plasmaB: "figures/supfigs/Supfig12c_kaplan_cibersort_plasmaB_OS_DRS.pdf"
    km_memoryB: "figures/supfigs/Supfig12d_kaplan_cibersort_memoryB_OS_DRS.pdf"
---

## Dependencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(ggpubr)
library(survival)
library(survminer)
library(ggrepel)
library(ggcorrplot)
library(conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("count", "dplyr")
conflict_prefer("rename", "dplyr")

theme_set(theme_bw())
```

```{r}
source(here("src/figures/study_km.R"))
source(here("src/figures/faceted_risktable.R"))
```

## Load data

```{r}
figuredata <- readRDS(here(params$figuredata))

head(figuredata)
```

## Colors

```{r}
study_colors <- c("NP-BC"="#521262", "Pr-BC"="#AA96DA",
                  "PP-BCdl"="#112D4E", "PP-BCpw"="#46CDCF")
```

## Fig 4B: Cibersort CD8 boxplot

```{r}
boxplot_cd8 <- figuredata %>%
  filter(!is.na(CD8_fraction)) %>%
  ggplot(aes(x = study_group, y = CD8_fraction)) +
  geom_jitter(aes(color = study_group), height = 0, width = 0.15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  stat_compare_means(method = "wilcox.test", comparisons = list(
    c("PP-BCpw", "Pr-BC"),
    c("PP-BCpw", "NP-BC")
  )) +
  stat_compare_means(label.y = 0.2) +
  xlab("Study group") + ylab("Relative CibersortX Fraction") +
  labs(color = "Study group")

boxplot_cd8
```

### Save output

```{r}
figureDir <- file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$boxplot_cd8), width = 10, height = 7)
boxplot_cd8
dev.off()
```

## Fig 4C: KM CD8 Cibersort

Due to power issues, omit lac. High/low is around the median density.

Filtering for this and other KMs is either `time_OS_months > 20 | reason_death == "BC"` or `time_DRS_months > 20 | reason_death == "BC"`, depending on the formula.

```{r}
km_cd8 <- km(
  rename(figuredata, CD8_CD3 = CD8_cat), covariate = "CD8_CD3", event_type = "OS",
  study_groups=c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_OS_months, event=death) ~ study_group + CD8_CD3"),
  returndata = F, returnfit = F
)

km_cd8 <- km_cd8 +
  ggtitle(paste(km_cd8$plot$title, "Cibersort"))

km_cd8
```


### Risk table

```{r}
rt_km_cd8 <- faceted_risktable(km_cd8, faceted.by = "CD8_CD3")
rt_km_cd8
```

### Save output

```{r}
pdf(file = here(params$km_cd8), width = 12, height = 7)
km_cd8
rt_km_cd8
dev.off()
```

## Sup fig 12B: Boxplot B cell subtypes

```{r, fig.height=5, fig.width=8}
boxplot_b_subtypes <- figuredata %>%
  filter(!is.na(Plasma_fraction)) %>%
  select(study_group, Plasma_fraction, naive_B_fraction, memory_B_fraction) %>%
  pivot_longer(cols = c(Plasma_fraction, naive_B_fraction, memory_B_fraction),
               names_to = "B_subtype", values_to = "relative_fraction") %>%
  ggplot(aes(x = study_group, y = relative_fraction)) +
  facet_wrap(~B_subtype) +
  geom_jitter(aes(color = study_group), height = 0, width = 0.15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  stat_compare_means(method = "wilcox.test", comparisons = list(
    c("PP-BCpw", "Pr-BC"),
    c("PP-BCpw", "NP-BC")
  )) +
  stat_compare_means(label.y = 0.3) +
  xlab("Study group") + ylab("Relative CibersortX Fraction") +
  labs(color = "Study group")

boxplot_b_subtypes
```

### Save output

```{r}
supFigDir <- file.path(here("figures/supfigs"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$boxplot_b_subtypes), width = 10, height = 7)
boxplot_b_subtypes
dev.off()
```

## Sup fig 12C: Cibersort Plasma B OS and DRS

### Overall survival

```{r}
km_plasmaB_OS <- km(
  rename(figuredata, plasma_B = Plasma_cat), covariate = "plasma_B", event_type = "OS",
  study_groups=c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_OS_months, event=death) ~ study_group + plasma_B"),
  returndata = F, returnfit = F
) 

km_plasmaB_OS <- km_plasmaB_OS +
  ggtitle(paste(km_plasmaB_OS$plot$title, "Cibersort"))

km_plasmaB_OS
```

### OS Risk table

```{r}
rt_km_plasmaB_OS <- faceted_risktable(km_plasmaB_OS, faceted.by = "plasma_B") +
  ggtitle("Risk table for OS plasma B")
rt_km_plasmaB_OS
```

### Distant recurrence

```{r}
km_plasmaB_DRS <- km(
  rename(figuredata, plasma_B = Plasma_cat), covariate = "plasma_B", event_type = "DRS",
  study_groups=c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + plasma_B"),
  returndata = F, returnfit = F
)

km_plasmaB_DRS <- km_plasmaB_DRS +
  ggtitle(paste(km_plasmaB_DRS$plot$title, "Cibersort"))

km_plasmaB_DRS
```

### DRS Risk table

```{r}
rt_km_plasmaB_DRS <- faceted_risktable(km_plasmaB_DRS, faceted.by = "plasma_B") +
  ggtitle("Risk table for DRS plasma B")
rt_km_plasmaB_DRS
```

### Save output

```{r}
pdf(file = here(params$km_plasmaB), width = 12, height = 7)
km_plasmaB_OS
rt_km_plasmaB_OS
km_plasmaB_DRS
rt_km_plasmaB_DRS
dev.off()
```


## Sup fig 12D: Cibersort Memory B OS and DRS

### Overall survival

```{r}
km_memoryB_OS <- km(
  rename(figuredata, memory_B = memory_B_cat), covariate = "memory_B", event_type = "OS",
  study_groups=c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_OS_months, event=death) ~ study_group + memory_B"),
  returndata = F, returnfit = F
)

km_memoryB_OS <- km_memoryB_OS +
  ggtitle(paste(km_memoryB_OS$plot$title, "Cibersort"))

km_memoryB_OS
```

### OS Risk table

```{r}
rt_km_memoryB_OS <- faceted_risktable(km_memoryB_OS, faceted.by = "memory_B") +
  ggtitle("Risk table for OS memory B")
rt_km_memoryB_OS
```

### Distant recurrence

```{r}
km_memoryB_DRS <- km(
  rename(figuredata, memory_B = Plasma_cat), covariate = "memory_B", event_type = "DRS",
  study_groups=c("PP-BCpw", "Pr-BC", "NP-BC"),
  surv_formula=as.formula("Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + memory_B"),
  returndata = F, returnfit = F
)

km_memoryB_DRS <- km_memoryB_DRS +
  ggtitle(paste(km_memoryB_DRS$plot$title, "Cibersort"))

km_memoryB_DRS
```

### DRS Risk table

```{r}
rt_km_memoryB_DRS <- faceted_risktable(km_memoryB_DRS, faceted.by = "memory_B") +
  ggtitle("Risk table for DRS memory B")
rt_km_memoryB_DRS
```

### Save output

```{r}
pdf(file = here(params$km_memoryB), width = 12, height = 7)
km_memoryB_OS
rt_km_memoryB_OS
km_memoryB_DRS
rt_km_memoryB_DRS
dev.off()
```

## Session info

```{r}
sessionInfo()
```
