---
title: "Boxplot IG signature cluster"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    df_print: tibble
    highlight: kate
params:
    cluster_results: "results/rnaseq/clustering/11_inv_clusters.xlsx"
    dds: "data/rnaseq/processed/08_dds_ovr_inv_vs_rest.Rds"
    bx_annot: "data/rnaseq/processed/bx_annot.Rds"
    figuredata: "data/figures/00_figuredata.Rds"
    fig1b: "figures/Fig1b_ig_boxplot.pdf"
    supfig2: "figures/supfigs/Supfig2_studygroup_km_forest.pdf"
    supfig8: "figures/supfigs/Supfig8_igSig_km_forest.pdf"
---

## Dependencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(here)
library(DESeq2)
library(tidyverse)
library(broom)
library(ggpubr)
library(survival)
library(survminer)

theme_set(theme_bw())
```

## Load data

DESeqDataSet with count matrix.

```{r}
dds <- readRDS(here(params$dds))
```

Gene annotation.

```{r}
bx_annot <- readRDS(here(params$bx_annot))
head(bx_annot)
```

K-means clustering results of DEGs from PPBC-pw vs rest.

```{r}
row_clusters <- readxl::read_excel(here(params$cluster_results),
                                   sheet="row_clusters") %>%
  relocate(row_cluster, ensembl_gene_id, .before = everything()) %>%
  arrange(row_cluster, padj)
head(row_clusters)
```

```{r}
col_clusters <- readxl::read_excel(here(params$cluster_results),
                                   sheet="col_clusters")
head(col_clusters)
```

Aggregated data for figures.

```{r}
figuredata <- readRDS(here(params$figuredata))
```

## Colors

```{r}
#Be sure these are in the same order as the factor level
levels(figuredata$study_group)
study_colors <- c("NP-BC"="#521262", "Pr-BC"="#aa96da",
                  "PP-BCdl"="#282c75", "PP-BCpw"="#46cdcf")
levels(figuredata$IG_col_cluster)
tertile_colors <- c("low"="#ffb600", "medium"="#ff7900","high"="#a10702")
```

## Immune cluster

Cluster 3 is dominated by IG gnes.

```{r}
list(
  cluster_1=row_clusters %>% filter(row_cluster == 1) %>% pull(gene_name),
  cluster_2=row_clusters %>% filter(row_cluster == 2) %>% pull(gene_name),
  cluster_3=row_clusters %>% filter(row_cluster == 3) %>% pull(gene_name)
  )
```

## IG metagene

Create an IG cluster metagene for each sample. For Hanne: see [this discussion on normalization](https://hbctraining.github.io/DGE_workshop/lessons/02_DGE_count_normalization.html) to understand the rationale.

```{r}
countmat <- DESeq2::counts(estimateSizeFactors(dds), normalized=T)
```

Select the IG genes

```{r}
igmat <- countmat[rownames(countmat) %in%
                    filter(row_clusters, row_cluster == 3)$ensembl_gene_id, ]

stopifnot(nrow(igmat) == nrow(filter(row_clusters, row_cluster == 3)))
```

Sum the expression of all genes within the IG cluster and add to figuredata.

```{r}
figuredata <- figuredata %>%
  left_join(., enframe(colSums(igmat), "sample_name", "igClusterScore"),
            by = "sample_name")
```

## Figure 1B - Boxplot IG signature

```{r}
igbp <- figuredata %>%
  filter(!is.na(igClusterScore)) %>%
  ggplot(aes(x = study_group,
             y = igClusterScore)) +
  geom_boxplot(aes(fill = study_group),
               outlier.shape = NA) +
  geom_point(position = position_jitter(), shape = 15) +
  scale_y_log10() +
  scale_fill_manual(values = study_colors) +
  xlab("Study group") + ylab("Log10 sum of normalized counts within IG cluster") +
  ggtitle("Relationship between IG gene signature expression and patient group") +
  stat_compare_means(method = "wilcox.test", comparisons = list(
    c('PP-BCpw', 'NP-BC'),
    c('PP-BCpw', 'Pr-BC')#,
    #c('PP-BCpw', 'PP-BCdl'), # Removed due to lack of samples
    #c('Pr-BC', 'NP-BC') # Removed due to lack of interest
  ), hide.ns = F)

igbp
```

### Save output

```{r}
figureDir <- file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$fig1b), width = 10, height = 7)
igbp
dev.off()
```

## Sup Fig 2 - KM and forest plot study group

In addition to the usual filters for follow up and reason for death, also filter on whether PAM50 and clinical covariates available. OS time and DRS time are applied to both OS and DRS regressions. We do this for both univariate and multivariate analyses so that both analyses have the same number of samples.

### 2A Univariate OS

```{r, fig.width=6.5, fig.height=5}
univariate_km <- function(df, event_type = c("OS", "DRS"), exclude_lac = F, returndata = F){
  
  event_type <- match.arg(event_type)
  
  if(exclude_lac){
    df <- df %>% filter(study_group != "PP-BCdl")
  }
  
  # Filter clinical covariates
  df <- df %>%
    #filter(study_group != "PP-BCdl") %>%
    filter(!is.na(year_diagnosis)) %>%
    filter(!is.na(stage)) %>%
    filter(!is.na(herceptin)) %>%
    filter(!is.na(death), !is.na(time_OS_months)) %>%
    filter(!is.na(distant_recurrence), !is.na(time_DRS_months))
  
  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>%
      filter(time_OS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Overall survival"
    fit <- survfit(Surv(time=time_OS_months, event=death) ~ study_group, data = df)
  } else {
    df <- df %>% filter(time_DRS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Distant recurrence"
    fit <- survfit(Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group, data = df)
  }
  #return(fit)
  
  colors <- study_colors[names(study_colors) %in% unique(df$study_group)]
  
  ggsurvplot(
    fit, data = df,
    xlab = "Months", 
    ylab = paste(y,"probability"),
    palette = colors,
    ggtheme = theme_bw(),
    pval = T,
    legend.labs = names(colors),
    legend.title="Study group",
    risk.table = T,
    risk.table.height = 0.25,
    cumevents = F, cumcensor = F
  )
}

supfig2a <- univariate_km(figuredata, exclude_lac=F, event_type = "OS", returndata = F) +
  ggtitle("Univariate OS")
supfig2a
```

### 2B Univariate DRS

```{r, fig.width=6.5, fig.height=5}
supfig2b <- univariate_km(figuredata, exclude_lac=F, event_type = "DRS", returndata = F) +
  ggtitle("Univariate DRS")
supfig2b
```

### 2C Multivariate OS Forest plot

Using all samples with OS follow up > 20 months and complete covariate data (stage, ciln subtype, treatment).

```{r}
multivariate_km <- function(df, event_type = c("OS", "DRS"),
                            reflevel, exclude_lac = F,
                            returndata = F, returncox=F){
  
  event_type <- match.arg(event_type)
  
  if(exclude_lac){
    df <- df %>% filter(study_group != "PP-BCdl")
  }
  
  # Filter clinical covariates
  df <- df %>%
    filter(!is.na(clin_subtype)) %>%
    filter(!is.na(stage)) %>%
    filter(!is.na(death), !is.na(time_OS_months)) %>%
    filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
    droplevels()

  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>%
      filter(time_OS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    
    y <- "Overall survival"
    
    fit <- coxph(
      Surv(time=time_OS_months, event=death) ~ stage + clin_subtype + study_group 
        + HT + RT + CT + herceptin
      ,
      data = as.data.frame(mutate(df, study_group = relevel(study_group, ref = reflevel)))
      )

  } else {
    
    df <- df %>% filter(time_DRS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Distant recurrence"
    
    fit <- coxph(
      Surv(time=time_DRS_months, event=distant_recurrence) ~ stage + clin_subtype + study_group 
      + HT + RT + CT + herceptin
      ,
      data = as.data.frame(mutate(df, study_group = relevel(study_group, ref = reflevel)))
      )
    
  }
  
  if(returncox){return(fit)}
  
  colors <- study_colors[names(study_colors) %in% unique(df$study_group)]
  
  ggforest(fit,data=df,main=ifelse(event_type == "OS",
                                   "HR: Multivariate overall survival",
                                   "HR: Multivariate distant recurrence"))
}

supfig2c <- multivariate_km(figuredata, reflevel = "PP-BCpw", event_type = "OS", returndata = F) 
supfig2c
```

### 2D Multivariate DRS

```{r}
supfig2d <- multivariate_km(figuredata,  reflevel = "PP-BCpw",
                            event_type = "DRS", returndata = F, returncox = F)
supfig2d
```

### Save output

```{r}
supFigDir <- file.path(here("figures/supfigs"))
dir.create(supFigDir, showWarnings = F)
pdf(file = here(params$supfig2), width = 10, height = 7)
print(supfig2a, newpage = FALSE) #Removes blank first page
supfig2b
supfig2c
supfig2d
dev.off()
```

## Sup Fig 8 - KM and forest IG signatures

### 8A OS KM IG sig

```{r, fig.width=6.5, fig.height=5}
univariate_ig <- function(df, event_type = c("OS", "DRS"),
                          colors = tertile_colors, exclude_lac = F, returndata = F){
  
  event_type <- match.arg(event_type)
  
  if(exclude_lac){
    df <- df %>% filter(study_group != "PP-BCdl")
  }
  
  # Filter clinical covariates
  df <- df %>%
    filter(!is.na(IG_col_cluster)) %>%
    filter(!is.na(year_diagnosis)) %>%
    filter(!is.na(stage)) %>%
    filter(!is.na(herceptin)) %>%
    filter(!is.na(death), !is.na(time_OS_months)) %>%
    filter(!is.na(distant_recurrence), !is.na(time_DRS_months))
  
  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>%
      filter(time_OS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Overall survival"
    fit <- survfit(Surv(time=time_OS_months, event=death) ~ IG_col_cluster,
                   data = df)
  } else {
    df <- df %>% filter(time_DRS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Distant recurrence"
    fit <- survfit(Surv(time=time_DRS_months,
                        event=distant_recurrence) ~ IG_col_cluster,
                   data = df)
  }
  #return(fit)
  colors <- colors[names(colors) %in% unique(df$IG_col_cluster)]
  
  ggsurvplot(
    fit, data = df,
    xlab = "Months", 
    ylab = paste(y,"probability"),
    palette = colors,
    ggtheme = theme_bw(),
    pval = T,
    legend.labs = names(colors),
    legend.title="Ig signatures",
    risk.table = T,
    risk.table.height = 0.25,
    cumevents = F, cumcensor = F
  )
}

supfig8a <- univariate_ig(figuredata, exclude_lac=F, event_type = "OS", returndata = F) +
  ggtitle("Univariate OS: Ig signatures")
supfig8a
```

### 8B DRS KM IG sig

```{r, fig.width=6.5, fig.height=5}
supfig8b <- univariate_ig(figuredata, exclude_lac=F, event_type = "DRS", returndata = F) +
  ggtitle("Univariate DRS: Ig signatures")
supfig8b
```

### 8C Multivariate OS Forest plot

Using all RNA samples with OS follow up > 20 months and complete covariate data (stage, clin subtype, treatment). Because it's RNA, use PAM50 instead of clin subtype.

```{r}
multivariate_ig <- function(df, event_type = c("OS", "DRS"),
                            colors = tertile_colors,
                            reflevel, exclude_lac = F,
                            returndata = F, returncox=F){
  
  event_type <- match.arg(event_type)
  
  if(exclude_lac){
    df <- df %>% filter(study_group != "PP-BCdl")
  }
  
  # Filter clinical covariates
  df <- df %>%
    filter(!is.na(PAM50)) %>%
    filter(!is.na(stage)) %>%
    filter(!is.na(death), !is.na(time_OS_months)) %>%
    filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
    droplevels()

  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>%
      filter(time_OS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    
    y <- "Overall survival"
    
    fit <- coxph(
      Surv(time=time_OS_months, event=death) ~ stage + PAM50 + IG_col_cluster 
        + HT + RT + CT + herceptin
      ,
      data = as.data.frame(mutate(df, IG_col_cluster = relevel(IG_col_cluster, ref = reflevel)))
      )

  } else {
    
    df <- df %>% filter(time_DRS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Distant recurrence"
    
    fit <- coxph(
      Surv(time=time_DRS_months, event=distant_recurrence) ~ stage + PAM50 + IG_col_cluster 
      + HT + RT + CT + herceptin
      ,
      data = as.data.frame(mutate(df, IG_col_cluster = relevel(IG_col_cluster, ref = reflevel)))
      )
    
  }
  
  if(returncox){return(fit)}
  
  colors <- colors[names(colors) %in% unique(df$IG_col_cluster)]
  
  ggforest(fit,data=df,main=ifelse(event_type == "OS",
                                   "HR: Multivariate overall survival",
                                   "HR: Multivariate distant recurrence"))
}

supfig8c <- multivariate_ig(figuredata, reflevel = "high", event_type = "OS", returndata = F) 
supfig8c
```

### 8D Multivariate DRS Forest plot

```{r}
supfig8d <- multivariate_ig(figuredata, reflevel = "high", event_type = "DRS", returndata = F) 
supfig8d
```

### Save output

```{r}
pdf(file = here(params$supfig8), width = 10, height = 7)
print(supfig8a, newpage = FALSE) #Removes blank first page
supfig8b
supfig8c
supfig8d
dev.off()
```

## Session info

```{r}
sessionInfo()
```
