---
title: "Isotype figures"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    df_print: tibble
    highlight: kate
params:
    figuredata: "data/figures/00_figuredata.Rds"
    normat: "data/rnaseq/processed/08_sfnorm_genesymbol_countmatrix.Rds"
    ig_milk_genes: "data/rnaseq/processed/08_ig_and_milk_genes.Rds"
    rt: "src/figures/faceted_risktable.R"
    boxplot_abs: "figures/Fig3a_boxplot_isotypes.pdf"
    km_ig_os: "figures/Fig3bc_km_isotypes.pdf"
    cor_IG_TAPC: "figures/Fig3d_cor_Ig_TAPC.pdf"
    cor_IgA_milk: "figures/supfigs/Supfig17_cor_IgA_milk.pdf"
    km_ig_drs: "figures/supfigs/Supfig18_km_isotypes_drs.pdf"
    cor_IgA_IgG: "figures/supfigs/Supfig20_cor_IgA_IgG.pdf"
---

## Dependencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(ggpubr)
library(survival)
library(survminer)
library(RColorBrewer)
library(ggrepel)
library(DESeq2)
library(conflicted)

conflict_prefer("filter", "dplyr")

theme_set(theme_bw())
```

```{r}
source(here(params$rt))
```

## Load data

```{r}
figuredata <- readRDS(here(params$figuredata))
```

```{r}
normat <- readRDS(here(params$normat))
```

```{r}
ig_milk_genes <- readRDS(here(params$ig_milk_genes))
```

## Colors

```{r}
study_colors <- c("NP-BC"="#521262", "Pr-BC"="#AA96DA",
                  "PP-BCdl"="#112D4E", "PP-BCpw"="#46CDCF")
```

## Figure 3A - Boxplot Ig expression

```{r, fig.height=6, fig.width=8}
boxplot_abs <- figuredata %>%
  select(study_group, IgA:IgM) %>%
  pivot_longer(cols = c(IgA, IgD, IgG, IgM),
               names_to = "isotype", values_to = "normCount") %>%
  filter(!is.na(normCount)) %>%
  ggplot(aes(x = study_group, y = normCount)) +
  geom_jitter(aes(color = isotype), shape = 15, height = 0, width = 0.2) +
  geom_boxplot(alpha = 0) +
  facet_wrap(~isotype, scale = "free_y") +
  scale_colour_discrete() +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "PP-BCdl"),
                                        c("PP-BCpw", "Pr-BC"),
                                        c("PP-BCpw", "NP-BC"))) +
  stat_compare_means() +
  #theme(text = element_text(size=18)) +
  ylab("Normalized Expression Levels") +
  xlab("Study group")

boxplot_abs
```

### Save output

```{r}
figureDir <- file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$boxplot_abs), width = 12, height = 8)
boxplot_abs
dev.off()
```

## Fig 3BC - IG expression KMs

IG expression level is split around the median. Exclude lac from RNAseq analyses due to lack of samples.

```{r}
ig_km <- function(df, event_type = c("OS", "DRS"), returndata = F){
  
  event_type <- match.arg(event_type)
  
  df <- df %>%
    filter(
      study_group %in% c("PP-BCpw", "Pr-BC", "NP-BC")
    )
  
  df <- df %>%
    select(study_group, IgA_rank:IgM_rank,
           death, time_OS_months,
           distant_recurrence, time_DRS_months,
           reason_death) %>%
    pivot_longer(cols = c(IgA_rank:IgM_rank), names_to = "isotype",
                 values_to = "rank") %>%
    mutate(isotype = str_remove(isotype, "_rank")) %>%
    mutate(isotype_group = paste(isotype, rank)) %>%
    filter(!is.na(rank)) %>%
    droplevels()
  
  if(returndata){return(df)}
  
  # There are a lot of facets, we can either make the chart very large, or
  # Use a for loop to plot the isotypes separately

  plt <- list()

  for(ig in c("IgA", "IgD", "IgM", "IgG")){
    
    dfi <- df[df$isotype == ig,]

    if(event_type == "OS"){
      dfi <- dfi %>% filter(!is.na(death), !is.na(time_OS_months)) %>%
        filter(time_OS_months > 20 | reason_death == "BC") %>%
        as.data.frame()
      y <- "Overall survival"
      fit <- survfit(Surv(time=time_OS_months, event=death) ~ study_group + isotype_group, data = dfi)
    } else {
      dfi <- dfi %>% filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
        filter(time_DRS_months > 20 | reason_death == "BC") %>%
        as.data.frame()
      y <- "Distant recurrence"
      fit <- survfit(Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + isotype_group, data = dfi)
    }
    #return(fit)
    plt[[length(plt)+1]] <- ggsurvplot(
      fit, data = dfi,
      xlab = "Months", 
      ylab = paste(y,"probability"),
      palette = study_colors[names(study_colors) %in% unique(dfi$study_group)],
      ggtheme = theme_bw(),
      pval = T,
      facet.by = "isotype_group",
      legend.title="Study group:"
    )
  }
   
  names(plt) <-  c("IgA", "IgD", "IgM", "IgG")
  plt
}

km_ig_os <- ig_km(figuredata, event_type = "OS", returndata = F)

km_ig_os
```

### Risk table

```{r}
rt_km_ig_os <- lapply(names(km_ig_os), function(x){
  faceted_risktable(km_ig_os[[x]], faceted.by = "isotype_group") +
    ggtitle(paste("Risk table for", x))
}) %>% set_names(names(km_ig_os))
rt_km_ig_os
```

### Save output

```{r}
pdf(file = here(params$km_ig_os), width = 9, height = 6)
km_ig_os
rt_km_ig_os
dev.off()
```

## Fig 3D - Correlation IgA/IgG and TAPC in PP-BCpw

```{r}
cor_IG_TAPC <- 
  list(
    IgA = figuredata %>%
      filter(study_group == "PP-BCpw") %>%
      filter(!is.na(TAPC_score)) %>%
      ggplot(aes(x=TAPC_score, y = IgA)) +
      geom_jitter(aes(color = as.factor(TAPC_score)), height = 0, width = 0.15) +
      geom_boxplot(alpha = 0, aes(group = TAPC_score), color = "darkgray"#, linetype="dashed"
      ) +
      geom_smooth(method = "lm", formula = y ~ x, linetype="dashed") +
      ggpubr::stat_cor(method = "kendall", label.y = 3.5, label.x = -0.2,
                       aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~"))) +
      ggtitle("Kendall's correlation IgA and TAPC in PP-BCpw") +
      xlab("TAPC score") + ylab("Normalized IgA expression") +
      labs(color = "TAPC score") +
      scale_color_brewer(palette = "Spectral"),
    
    IgD = figuredata %>%
      filter(study_group == "PP-BCpw") %>%
      filter(!is.na(TAPC_score)) %>%
      ggplot(aes(x=TAPC_score, y = IgG)) +
      geom_jitter(aes(color = as.factor(TAPC_score)), height = 0, width = 0.15) +
      geom_boxplot(alpha = 0, aes(group = TAPC_score), color = "darkgray"#, linetype="dashed"
      ) +
      geom_smooth(method = "lm", formula = y ~ x, linetype="dashed") +
      ggpubr::stat_cor(method = "kendall", #label.y = 3.5, label.x = -0.2,
                       aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~"))) +
      ggtitle("Kendall's correlation IgG and TAPC in PP-BCpw") +
      xlab("TAPC score") + ylab("Normalized IgG expression") +
      labs(color = "TAPC score") +
      scale_color_brewer(palette = "Spectral")
  )

cor_IG_TAPC
```

### Save output

```{r}
pdf(file = here(params$cor_IG_TAPC), width = 12, height = 8)
cor_IG_TAPC
dev.off()
```

## Sup Fig 17: Correlation IgA and milk

Similar to previous figures, manual selection of some milk-associated genes, based on volcano plots of diffex genes and Cook's outliers.

```{r}
milk_genes <- ig_milk_genes$milk_genes
milk_genes
```

As previous figures, compute metagene for DE milk proteins and IgA on the basis of size factor. Use log transformation with a pseudocount for visualization.
Note: there is also IgA expression in `figuredata`, but it's computed based on TMM. Size factor normalization is preferred here for consistency with previous figures related to IG expression.

### All samples

```{r}
# Create milk and IG metagenes
milk_metagene <- colSums(normat[rownames(normat) %in% milk_genes,])
IgA_norm <- colSums(normat[rownames(normat) %in% c("IGHA1", "IGHA2"),])

igA_milk_cor_all <- tibble(sample_name = colnames(normat), milk_metagene, IgA_norm) %>%
  left_join(., select(figuredata, sample_name, study_group), by = "sample_name") %>%
  ggplot(aes(x = IgA_norm, y = milk_metagene)) +
  geom_point(aes(color = study_group), shape = 15) +
  scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
  geom_smooth(method = "lm", formula = y ~ x) +
      ggpubr::stat_cor(method = "spearman", #label.y = 3.5, label.x = -0.2,
                       aes(label = paste(..rr.label.., ..r.label.., ..p.label..,
                                         sep = "~`,`~"))) +
      ggtitle("Spearman correlation IgA and milk metagene: All groups") +
  ylab("Metagene milk proteins\n(log10 size factor normalized + 1)") +
  xlab("IgA isotype counts\n(log10 size factor normalized + 1)") +
  labs(color = "Study group")

igA_milk_cor_all
```

### By study group

```{r}
igA_milk_cor_group <- tibble(sample_name = colnames(normat), milk_metagene, IgA_norm) %>%
  left_join(., select(figuredata, sample_name, study_group), by = "sample_name") %>%
  ggplot(aes(x = IgA_norm, y = milk_metagene)) +
  geom_point(aes(color = study_group), shape = 15) +
  scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
  geom_smooth(method = "lm", formula = y ~ x) +
      ggpubr::stat_cor(method = "spearman", #label.y = 3.5, label.x = -0.2,
                       aes(label = paste(..rr.label.., ..r.label.., ..p.label..,
                                         sep = "~`,`~"))) +
      ggtitle("Spearman correlation IgA and milk metagene: All groups") +
  ylab("Metagene milk proteins\n(log10 size factor normalized + 1)") +
  xlab("IgA isotype counts\n(log10 size factor normalized + 1)") +
  labs(color = "Study group") +
  facet_wrap(~study_group, scales = "free_x")

igA_milk_cor_group
```

### Save output

```{r}
pdf(file = here(params$cor_IgA_milk), width = 12, height = 8)
igA_milk_cor_all
igA_milk_cor_group
dev.off()
```

## Sup fig 18: Ab isotype DRS

```{r}
km_ig_drs <- ig_km(figuredata, event_type = "DRS", returndata = F)

km_ig_drs
```

### Risk table

```{r}
rt_km_ig_drs <- lapply(names(km_ig_drs), function(x){
  faceted_risktable(km_ig_drs[[x]], faceted.by = "isotype_group") +
    ggtitle(paste("Risk table for", x))
}) %>% set_names(names(km_ig_drs))
rt_km_ig_drs
```

### Save output

```{r}
pdf(file = here(params$km_ig_drs), width = 12, height = 8)
km_ig_drs
rt_km_ig_drs
dev.off()
```

## Sup fig 20: Correlation IgA and IgG

### All samples

```{r}
cor_plot <- function(df=figuredata, x, y, colorby, cormet="spearman", ...){
  
  df <- df[!is.na(df[,x]) & !is.na(df[,y]),]

  df %>%
    ggplot(aes(get(x), get(y))) +
    geom_jitter(aes(color = get(colorby)), height = 1, width = 0.2, shape = 15) +
    geom_smooth(method = "lm", formula = y ~ x) +
    ggpubr::stat_cor(method = cormet, #label.y = 3.5, label.x = -0.2,
                     aes(label = paste(..rr.label.., ..r.label.., ..p.label..,
                                       sep = "~`,`~"))) +
    xlab(str_replace_all(x, "_", " ")) +
    ylab(str_replace_all(y, "_", " ")) +
    labs(color = str_replace_all(colorby, "_", " "))
}

cor_IgA_IgG_all <- cor_plot(figuredata, x="IgA", y="IgG", colorby = "study_group") +
  scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
  ggtitle("Spearman correlation between IgA and IgG in all samples") +
  xlab("TMM log2 IgA isotype expression")+
  ylab("TMM log2 IgG isotype expression")
cor_IgA_IgG_all
```

### PPBCpw

```{r}
cor_IgA_IgG_inv <- cor_plot(
  filter(figuredata, study_group == "PP-BCpw"),
  x="IgA", y="IgG", colorby = "study_group"
) +
  scale_color_manual(values = study_colors[names(study_colors) != "rest"]) +
  ggtitle("Spearman correlation between IgA and IgG in all samples") +
  xlab("TMM log2 IgA isotype expression")+
  ylab("TMM log2 IgG isotype expression")

cor_IgA_IgG_inv
```

### Save output

```{r}
pdf(file = here(params$cor_IgA_IgG), width = 12, height = 8)
cor_IgA_IgG_all
cor_IgA_IgG_inv
dev.off()
```

## Session info

```{r}
sessionInfo()
```

