---
title: "Spatial figures"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: yes
    df_print: tibble
    highlight: kate
params:
    figuredata: "data/figures/00_figuredata.Rds"
    sampledata: "data/external/sample_data.tsv"
    density: "data/vectra/processed/density_ppbc.Rds"
    radius: 0.03
    spatstat_l: "data/vectra/processed/11_spatstat_l.Rds"
    spatstat_lcross_immune: "data/vectra/processed/11_spatstat_lcross_immune.Rds"
    spatstat_lcross_panck: "data/vectra/processed/11_spatstat_lcross_panck.Rds"
    fig2b: "figures/Fig2b_CD20_density_boxplot.pdf"
    fig2c: "figures/Fig2c_CD20_density_km.pdf"
    rt_fig2c: "figures/Fig2c_risktable.csv"
    fig2e: "figures/Fig2e_CD20_l_boxplot.pdf"
    fig2f: "figures/Fig2f_Cd20_l_km.pdf"
    rt_fig2f: "figures/Fig2f_risktable.csv"
    fig4d: "figures/Fig4d_CD8_density_boxplot_KM.pdf"
---

## Dependencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(here)
library(tidyverse)
library(ggpubr)
library(survival)
library(survminer)
library(ggrepel)
library(conflicted)

conflict_prefer("filter", "dplyr")

theme_set(theme_bw())
```

## Load data

Linkage of t-numbers to patient_IDs.

```{r}
sampledata <- read_tsv(here(params$sampledata))
```

Processed outcome data.

```{r}
figuredata <- readRDS(here(params$figuredata))
```

Cell density from Vectra data.

```{r}
density <- readRDS(here(params$density)) %>%
  mutate(original_group = study_group, .after = study_group) %>%
  mutate(study_group = recode(study_group, "npbc"="NP-BC", "prbc"="Pr-BC",
                              "ppbcdl"="PP-BCdl", "ppbcpw"="PP-BCpw")) %>%
  rename(sample_ID = t_number)
```

L stat (same cell type)

```{r}
l <- readRDS(here(params$spatstat_l))
```

L cross immune <-> immune

```{r}
lximmune <- readRDS(here(params$spatstat_lcross_immune))
```

L cross immune <-> panck

```{r}
lxpanck <- readRDS(here(params$spatstat_lcross_panck))
```

## Colors

```{r}
figuredata$study_group %>% levels()
study_colors <- c("NP-BC"="#521262", "Pr-BC"="#AA96DA", "PP-BCdl"="#112D4E", "PP-BCpw"="#46CDCF")
```

## Link outcomes

Add the t-numbers to the figure data.

```{r}
outcomedata <- sampledata %>%
  filter(sample_type == "slide" & !is.na(sample_ID) & !is.na(batch_HALO) & Included == 1) %>%
  select(patient_ID, sample_ID) %>%
  distinct() %>%
  inner_join(., figuredata, by = "patient_ID")

stopifnot(nrow(filter(outcomedata, duplicated(sample_ID))) == 0)
```

## Figure 2B - CD20 density boxplot

```{r}
fig2b <- density %>%
  filter(cell_type == "CD20+" & classifier_label == "Total") %>%
  ggplot(aes(x = factor(study_group,
                        levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
             y = log10(density + 1), color = study_group)) +
  geom_point(position = position_jitter(height = 0), shape = 15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  xlab("Study group") + #ylab("CD20+ multiplex IF density") +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "NP-BC"))) +
  stat_compare_means(label.y = 3.7) +
  ggtitle("CD20+ density by study group")

fig2b
```

### Save output

```{r}
figureDir = file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$fig2b), width = 10, height = 7)
fig2b
dev.off()
```

## Figure 2C - CD20 density KM

Due to power issues, only use PP-BCpw and NP-BC. High/low is around the median density.

```{r}
density_km <- function(df, cell, event_type = c("OS", "DRS"), returndata = F){
  
  stopifnot(cell %in% df$cell_type)
  
  event_type <- match.arg(event_type)
  
  df <- df %>% 
    filter(
      classifier_label == "Total",
      study_group %in% c("PP-BCpw", "NP-BC"),
      cell_type %in% cell,
      !is.na(density)
    ) %>%
    rename(density_int = density)
  
  med <- median(df$density_int)
  #print(med)
  
  df$density <- factor(ifelse(df$density_int > med, "high", "low"),
                       levels = c("low", "high"))
  df <- df %>% relocate(density, .after = density_int)
  
  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>% filter(!is.na(death), !is.na(time_OS_months)) %>%
      as.data.frame()
    y <- "Overall survival"
    fit <- survfit(Surv(time=time_OS_months, event=death) ~ study_group + density, data = df)
  } else {
    df <- df %>% filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
      as.data.frame()
    y <- "Distant recurrence"
    fit <- survfit(Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + density, data = df)
  }
  
  ggsurvplot(
    fit, data = df,
    xlab = "Months", 
    ylab = paste(y,"probability"),
    title = paste(y, cell, "stratified by median density:", signif(med, 2)),
    palette = study_colors[names(study_colors) %in% unique(df$study_group)],
    ggtheme = theme_bw(),
    pval = T,
    facet.by = "density",
    legend.title="Study group:"
  )
}

fig2c <- density_km(density, cell = "CD20+", event_type = "OS", returndata = F)
fig2c
```

### Risk table

```{r, fig.height=7, fig.width=12}
risktable <- function(ggkm, response, text.size = 8){
  ggplot(ggkm$data, aes(time, study_group)) + 
    ggrepel::geom_text_repel(aes(label = n.risk)) +
    facet_wrap(as.formula(paste("~", response))) + theme_bw() + 
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.background = element_rect(fill="white", size = 1),
          axis.text.x = element_text(color="black", size = text.size+2),
          axis.text.y = element_text(color="black", size = text.size),
          axis.title = element_text(color="black", size = text.size+4),
          strip.text = element_text(color="black", size = text.size))+
    ggtitle("Risk Tables") +
    xlab("Time") + ylab("")
}

rt_fig2c <- risktable(fig2c, response = "density")
rt_fig2c
```

### Save output

```{r}
pdf(file = here(params$fig2c), width = 12, height = 7)
fig2c
rt_fig2c
dev.off()
```

Also save risk table as csv.

```{r}
write_csv(fig2c$data, here(params$rt_fig2c))
```

## Figure 2E - CD20 delta L boxplot

```{r}
lstat_boxplot <-  function(df, cell, outcome=outcomedata, rad=params$radius,
                           pseudocount = 0.01, returndata = F) {
  
  stopifnot(class(rad)=="numeric")  
  stopifnot(cell %in% df$cell_type)
  
  df <- df %>%
    left_join(., outcome, by="sample_ID") %>%
    filter(
      cell_type %in% cell,
      radius == rad,
      !is.na(delta_estimate)
    )
  
  if(returndata==T){return(df)}
  
  df %>%
    ggplot(aes(x = factor(study_group,
                          levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
               y = log10(delta_estimate + pseudocount), color = study_group)) +
    geom_point(position = position_jitter(height = 0), shape = 15) +
    geom_boxplot(alpha = 0) +
    scale_color_manual(values = study_colors) +
    xlab("Study group") + #ylab("CD20+ multiplex IF density") +
    stat_compare_means(method = "wilcox.test",
                       comparisons = list(c("PP-BCpw", "NP-BC"))) +
    stat_compare_means() +
    ggtitle("CD20+ density by study group") +
    ylab(paste("Log10 delta estimate +", pseudocount))
  
}

fig2e <- lstat_boxplot(l, cell = "CD20+")
fig2e
```
### Save output

```{r}
figureDir = file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$fig2e), width = 10, height = 7)
fig2e
dev.off()
```

## Figure 2F - CD20 delta L KM

Due to power issues, only use PP-BCpw and NP-BC. Stratification is around the median delta L estimate.

```{r}
lstat_km <- function(df, cell, outcome=outcomedata, rad=params$radius,
                     event_type = c("OS", "DRS"), returndata = F){
  
  stopifnot(class(rad)=="numeric")  
  stopifnot(cell %in% df$cell_type)
  
  event_type <- match.arg(event_type)
  
  df <- df %>%
    left_join(., outcome, by="sample_ID") %>%
    filter(
      study_group %in% c("PP-BCpw", "NP-BC"),
      cell_type %in% cell,
      radius == rad,
      !is.na(delta_estimate)
    )
  
  med <- median(df$delta_estimate)
  #print(med)
  
  df$delta_l <- factor(ifelse(df$delta_estimate > med, "high", "low"),
                       levels = c("low", "high"))
  df <- df %>% relocate(delta_l, .after = delta_estimate)
  
  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>% filter(!is.na(death), !is.na(time_OS_months)) %>%
      as.data.frame()
    y <- "Overall survival"
    fit <- survfit(Surv(time=time_OS_months, event=death) ~ study_group + delta_l, data = df)
  } else {
    df <- df %>% filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
      as.data.frame()
    y <- "Distant recurrence"
    fit <- survfit(Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + delta_l, data = df)
  }
  
  ggsurvplot(
    fit, data = df,
    xlab = "Months", 
    ylab = paste(y,"probability"),
    title = paste(y, cell, "stratified by median delta L:", signif(med, 2)),
    palette = study_colors[names(study_colors) %in% unique(df$study_group)],
    ggtheme = theme_bw(),
    pval = T,
    facet.by = "delta_l",
    legend.title="Study group:"
  )
}

fig2f <- lstat_km(l, cell = "CD20+", event_type = "OS", returndata = F)
fig2f
```

### Risk table

```{r, fig.height=7, fig.width=12}
rt_fig2f <- risktable(fig2f, response = "delta_l")
rt_fig2f
```

### Save output

```{r}
pdf(file = here(params$fig2f), width = 12, height = 7)
fig2f
rt_fig2f
dev.off()
```

Also save risk table as csv.

```{r}
write_csv(fig2f$data, here(params$rt_fig2f))
```

## Figure 4D - CD8 density boxplot and KM

### Boxplot

```{r}
fig4d <- density %>%
  filter(cell_type == "CD3+CD8+" & classifier_label == "Total") %>%
  ggplot(aes(x = factor(study_group,
                        levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
             y = log10(density + 1), color = study_group)) +
  geom_point(position = position_jitter(height = 0), shape = 15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  xlab("Study group") + #ylab("CD20+ multiplex IF density") +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "NP-BC"))) +
  stat_compare_means(label.y = 3.7) +
  ggtitle("CD3+CD8+ density by study group")

fig4d
```

### Kaplan-Meier

```{r}
fig4d2 <- density_km(density, cell = "CD3+CD8+", event_type = "OS", returndata = F)
fig4d2
```

### Risk table 

```{r, fig.height=7, fig.width=12}
rt_fig4d2 <- risktable(fig4d2, response = "density")
rt_fig4d2
```

### Save output

```{r}
pdf(file = here(params$fig4d), width = 12, height = 7)
fig4d
fig4d2
rt_fig4d2
dev.off()
```

Also save risk table as csv.

```{r}
write_csv(fig4d2$data, here(params$rt_fig4d))
```

## Session info

```{r}
sessionInfo()
```
