---
title: "Spatial figures"
author: "Kat Moore"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: yes
    df_print: tibble
    highlight: kate
params:
    figuredata: "data/figures/00_figuredata.Rds"
    sampledata: "data/external/sample_data.tsv"
    density: "data/vectra/processed/density_ppbc.Rds"
    radius: 0.03
    spatstat_l: "data/vectra/processed/11_spatstat_l.Rds"
    spatstat_lcross_immune: "data/vectra/processed/11_spatstat_lcross_immune.Rds"
    spatstat_lcross_panck: "data/vectra/processed/11_spatstat_lcross_panck.Rds"
    boxplot_CD20_density: "figures/Fig2b_CD20_density_boxplot.pdf"
    km_CD20_density_OS: "figures/Fig2c_CD20_density_km_OS.pdf"
    boxplot_CD20_lstat: "figures/Fig2e_CD20_l_boxplot.pdf"
    km_CD20_lstat_OS: "figures/Fig2f_Cd20_l_km_OS.pdf"
    boxplot_CD8_density_km_OS: "figures/Fig4d_CD8_density_boxplot_km_OS.pdf"
    boxplot_CD8_l_km_OS: "figures/Fig4e_CD8_l_boxplot_km_OS.pdf"
    boxplot_CD8_PancK_lcross_km_OS: "figures/Fig4f_CD8_PanCk_lcross_boxplot_km_OS.pdf"
    boxplot_CD8_CD4_lcross_km_OS: "figures/Fig4g_CD8_CD4_lcross_boxplot_km_OS.pdf"
    boxplot_CD20_CD8_lcross_km_OS: "figures/Fig4h_CD20_CD8_lcross_boxplot_km_OS.pdf"
    boxplot_CD20_CD4_lcross_km_OS: "figures/Fig4i_CD20_CD4_lcross_boxplot_km_OS.pdf"
    km_CD20_density_DRS: "figures/supfigs/Supfig13a_CD20_density_km_DRS.pdf"
    km_CD20_l_DRS: "figures/supfigs/Supfig13b_CD20_l_km_DRS.pdf"
    boxplot_CD20_Panck_lcross: "figures/supfigs/Supfig14a_CD20_Panck_lcross_boxplot.pdf"
    km_CD20_Panck_lcross_km_OS: "figures/supfigs/Supfig14b_CD20_Panck_lcross_km_OS.pdf"
    km_CD20_Panck_lcross_km_DRS: "figures/supfigs/Supfig14c_CD20_Panck_lcross_km_DRS.pdf"
    misc_spatial_drs: "figures/supfigs/Supfig23b-g_misc_spatial_km_drs.pdf"
---

## Dependencies

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(ggpubr)
library(survival)
library(survminer)
library(ggrepel)
library(conflicted)

conflict_prefer("filter", "dplyr")

theme_set(theme_bw())
```

```{r}
source(here("src/figures/faceted_risktable.R"))
```

## Load data

Linkage of t-numbers to patient_IDs.

```{r}
sampledata <- read_tsv(here(params$sampledata))
```

Processed outcome data.

```{r}
figuredata <- readRDS(here(params$figuredata))
```

Cell density from Vectra data.

```{r}
density <- readRDS(here(params$density)) %>%
  mutate(original_group = study_group, .after = study_group) %>%
  mutate(study_group = recode(study_group, "npbc"="NP-BC", "prbc"="Pr-BC",
                              "ppbcdl"="PP-BCdl", "ppbcpw"="PP-BCpw")) %>%
  rename(sample_ID = t_number)
```

L stat (same cell type)

```{r}
l <- readRDS(here(params$spatstat_l))
```

L cross immune <-> immune

```{r}
lximmune <- readRDS(here(params$spatstat_lcross_immune))
```

L cross immune <-> panck

```{r}
lxpanck <- readRDS(here(params$spatstat_lcross_panck))
```

## Colors

```{r}
figuredata$study_group %>% levels()
study_colors <- c("NP-BC"="#521262", "Pr-BC"="#AA96DA", "PP-BCdl"="#112D4E", "PP-BCpw"="#46CDCF")
```

## Link outcomes

Add the t-numbers to the figure data.

```{r}
outcomedata <- sampledata %>%
  filter(sample_type == "slide" & !is.na(sample_ID) & !is.na(batch_HALO) & Included == 1) %>%
  select(patient_ID, sample_ID) %>%
  distinct() %>%
  inner_join(., figuredata, by = "patient_ID")

stopifnot(nrow(filter(outcomedata, duplicated(sample_ID))) == 0)
```

## Figure 2B - CD20 density boxplot

```{r}
boxplot_CD20_density <- density %>%
  filter(cell_type == "CD20+" & classifier_label == "Total") %>%
  ggplot(aes(x = factor(study_group,
                        levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
             y = log10(density + 1), color = study_group)) +
  geom_point(position = position_jitter(height = 0), shape = 15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  xlab("Study group") + #ylab("CD20+ multiplex IF density") +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "NP-BC"))) +
  stat_compare_means(label.y = 3.7) +
  ggtitle("CD20+ density by study group")

boxplot_CD20_density
```

### Save output

```{r}
figureDir = file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$boxplot_CD20_density), width = 10, height = 7)
boxplot_CD20_density
dev.off()
```

## Figure 2C - CD20 density KM OS

Due to power issues, only use PP-BCpw and NP-BC. High/low is around the median density.

```{r}
density_km <- function(df, cell, event_type = c("OS", "DRS"), returndata = F){
  
  stopifnot(cell %in% df$cell_type)
  
  event_type <- match.arg(event_type)
  
  df <- df %>% 
    filter(
      classifier_label == "Total",
      study_group %in% c("PP-BCpw", "NP-BC"),
      cell_type %in% cell,
      !is.na(density)
    ) %>%
    rename(density_int = density)
  
  med <- median(df$density_int)
  #print(med)
  
  df$density <- factor(ifelse(df$density_int > med, "high", "low"),
                       levels = c("low", "high"))
  df <- df %>% relocate(density, .after = density_int)
  
  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>% filter(!is.na(death), !is.na(time_OS_months)) %>%
      filter(time_OS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Overall survival"
    fit <- survfit(Surv(time=time_OS_months, event=death) ~ study_group + density, data = df)
  } else {
    df <- df %>% filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
      filter(time_DRS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Distant recurrence"
    fit <- survfit(Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + density, data = df)
  }
  
  ggsurvplot(
    fit, data = df,
    xlab = "Months", 
    ylab = paste(y,"probability"),
    title = paste(y, cell, "stratified by median density:", signif(med, 2)),
    palette = study_colors[names(study_colors) %in% unique(df$study_group)],
    ggtheme = theme_bw(),
    pval = T,
    facet.by = "density",
    legend.title="Study group"
  )
}

km_CD20_density_OS <- density_km(density, cell = "CD20+", event_type = "OS", returndata = F)
km_CD20_density_OS
```

### Risk table

```{r}
rt_km_CD20_density_OS <- faceted_risktable(km_CD20_density_OS, faceted.by = "density") +
  ggtitle("Risk table for CD20+ density OS")
rt_km_CD20_density_OS
```

### Save output

```{r}
pdf(file = here(params$km_CD20_density_OS), width = 12, height = 7)
km_CD20_density_OS
rt_km_CD20_density_OS
dev.off()
```

## Figure 2E - CD20 delta L boxplot

```{r}
lstat_boxplot <-  function(df, cell, outcome=outcomedata, rad=params$radius,
                           returndata = F, ...) {
  
  stopifnot(class(rad)=="numeric")  
  stopifnot(cell %in% df$cell_type)
  
  df <- df %>%
    left_join(., outcome, by="sample_ID") %>%
    filter(
      cell_type %in% cell,
      radius == rad,
      !is.na(delta_estimate)
    )
  
  if(returndata==T){return(df)}
  
  df %>%
    ggplot(aes(x = factor(study_group,
                          levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
               y = delta_estimate, color = study_group)) +
    geom_point(position = position_jitter(height = 0), shape = 15) +
    geom_boxplot(alpha = 0) +
    scale_color_manual(values = study_colors) +
    xlab("Study group") + #ylab("CD20+ multiplex IF density") +
    stat_compare_means(method = "wilcox.test",
                       comparisons = list(c("PP-BCpw", "NP-BC"))) +
    stat_compare_means(...) +
    ggtitle(paste(cell, "clustering by study group")) +
    ylab(paste0("Delta estimate for ", cell)) +
    labs(color = "study group")
  
}

boxplot_CD20_lstat <- lstat_boxplot(l, cell = "CD20+", label.y = 1)
boxplot_CD20_lstat
```
### Save output

```{r}
figureDir = file.path(here("figures"))
dir.create(figureDir, showWarnings = F)
pdf(file = here(params$boxplot_CD20_lstat), width = 10, height = 7)
boxplot_CD20_lstat
dev.off()
```

## Figure 2F - CD20 delta L KM

Due to power issues, only use PP-BCpw and NP-BC. Stratification is around the median delta L estimate.

```{r}
lstat_km <- function(df, cell, outcome=outcomedata, rad=params$radius,
                     event_type = c("OS", "DRS"), returndata = F){
  
  stopifnot(class(rad)=="numeric")  
  stopifnot(cell %in% df$cell_type)
  
  event_type <- match.arg(event_type)
  
  df <- df %>%
    left_join(., outcome, by="sample_ID") %>%
    filter(
      study_group %in% c("PP-BCpw", "NP-BC"),
      cell_type %in% cell,
      radius == rad,
      !is.na(delta_estimate)
    )
  
  med <- median(df$delta_estimate)
  #print(med)
  
  df$delta_l <- factor(ifelse(df$delta_estimate > med, "high", "low"),
                       levels = c("low", "high"))
  df <- df %>% relocate(delta_l, .after = delta_estimate)
  
  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>% filter(!is.na(death), !is.na(time_OS_months)) %>%
      filter(time_OS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Overall survival"
    fit <- survfit(Surv(time=time_OS_months, event=death) ~ study_group + delta_l, data = df)
  } else {
    df <- df %>% filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
      filter(time_DRS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Distant recurrence"
    fit <- survfit(Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + delta_l, data = df)
  }
  
  ggsurvplot(
    fit, data = df,
    xlab = "Months", 
    ylab = paste(y,"probability"),
    title = paste(y, cell, "stratified by median delta L:", signif(med, 2)),
    palette = study_colors[names(study_colors) %in% unique(df$study_group)],
    ggtheme = theme_bw(),
    pval = T,
    facet.by = "delta_l",
    legend.title="Study group"
  )
}

km_CD20_lstat_OS <- lstat_km(l, cell = "CD20+", event_type = "OS", returndata = F)
km_CD20_lstat_OS
```

### Risk table

```{r}
rt_km_CD20_lstat_OS <- faceted_risktable(km_CD20_lstat_OS, faceted.by = "delta_l") +
  ggtitle("Risk table for delta L: CD20+ OS")
rt_km_CD20_lstat_OS
```

### Save output

```{r}
pdf(file = here(params$km_CD20_lstat_OS), width = 12, height = 7)
km_CD20_lstat_OS
rt_km_CD20_lstat_OS
dev.off()
```

## Figure 4D - CD8 density boxplot and KM

### Boxplot

```{r}
boxplot_CD8_density <- density %>%
  filter(cell_type == "CD3+CD8+" & classifier_label == "Total") %>%
  ggplot(aes(x = factor(study_group,
                        levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
             y = log10(density + 1), color = study_group)) +
  geom_point(position = position_jitter(height = 0), shape = 15) +
  geom_boxplot(alpha = 0) +
  scale_color_manual(values = study_colors) +
  xlab("Study group") + #ylab("CD20+ multiplex IF density") +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c("PP-BCpw", "NP-BC"))) +
  stat_compare_means(label.y = 3.7) +
  ggtitle("CD3+CD8+ density by study group")

boxplot_CD8_density
```

### Kaplan-Meier

```{r}
km_CD8_density_OS <- density_km(density, cell = "CD3+CD8+", event_type = "OS", returndata = F)
km_CD8_density_OS
```

### Risk table 

```{r}
rt_km_CD8_density_OS <- faceted_risktable(km_CD8_density_OS, faceted.by = "density") +
  ggtitle("Risk table for CD3+CD8+ density OS")
rt_km_CD8_density_OS
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD8_density_km_OS), width = 12, height = 7)
boxplot_CD8_density
km_CD8_density_OS
rt_km_CD8_density_OS
dev.off()
```

## Figure 4E - CD8+ delta L boxplot and KM

### Boxplot

```{r}
boxplot_CD8_l <- lstat_boxplot(l, cell = "CD3+CD8+", label.y = 0.25)
boxplot_CD8_l
```

### Kaplan-Meier

```{r}
km_CD8_l_OS <- lstat_km(l, cell = "CD3+CD8+", event_type = "OS", returndata = F)
km_CD8_l_OS
```

### Risk table 

```{r}
rt_km_CD8_l_OS <- faceted_risktable(km_CD8_l_OS, faceted.by = "delta_l") +
  ggtitle("Risk table for CD3+CD8+ delta L OS")
rt_km_CD8_l_OS
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD8_l_km_OS), width = 12, height = 7)
boxplot_CD8_l
km_CD8_l_OS
rt_km_CD8_l_OS
dev.off()
```

## Figure 4F - CD3+CD8+ <-> PanCK L cross boxplot and KM

### Boxplot

```{r}
lcross_boxplot <-  function(df, cell.1, cell.2, outcome=outcomedata,
                            rad=params$radius, returndata = F, ...) {
  
  stopifnot(class(rad)=="numeric")  
  stopifnot(cell.1 %in% df$cell_type1)
  stopifnot(cell.2 %in% df$cell_type2)
  
  df <- df %>%
    left_join(., outcome, by="sample_ID") %>%
    filter(
      cell_type1 %in% cell.1,
      cell_type2 %in% cell.2,
      radius == rad,
      !is.na(delta_estimate)
    ) %>%
    mutate(cell = paste(cell_type1, cell_type2, sep=" <-> "), .before = cell_type1)
  
  cell <- unique(df$cell)
  stopifnot(length(cell)==1)
  
  if(returndata==T){return(df)}
  
  df %>%
    filter(!is.na(delta_estimate)) %>%
    ggplot(aes(x = factor(study_group,
                          levels = c("NP-BC", "Pr-BC", "PP-BCdl","PP-BCpw")),
               y = delta_estimate, color = study_group)) +
    geom_point(position = position_jitter(height = 0), shape = 15) +
    geom_boxplot(alpha = 0) +
    scale_color_manual(values = study_colors) +
    xlab("Study group") + #ylab("CD20+ multiplex IF density") +
    stat_compare_means(method = "wilcox.test",
                       comparisons = list(c("PP-BCpw", "NP-BC"))) +
    stat_compare_means(...) +
    ggtitle(paste(cell, "clustering by study group")) +
    ylab(paste0(paste0("Delta estimate for ", cell))) +
    labs(color = "study group")
  
}

boxplot_CD8_Panck_lcross <- lcross_boxplot(lximmune, cell.1 = "CD3+CD8+", cell.2 = "PanCK+",
                                           label.y = 0.01, returndata=F)
boxplot_CD8_Panck_lcross
```

### Kaplan-meier

```{r}
lcross_km <- function(df, cell.1, cell.2, outcome=outcomedata, rad=params$radius,
                      event_type = c("OS", "DRS"), returndata = F){
  
  stopifnot(class(rad)=="numeric")  
  stopifnot(cell.1 %in% df$cell_type1)
  stopifnot(cell.2 %in% df$cell_type2)
  
  event_type <- match.arg(event_type)
  
  df <- df %>%
    left_join(., outcome, by="sample_ID") %>%
    filter(
      study_group %in% c("PP-BCpw", "NP-BC"),
      cell_type1 %in% cell.1,
      cell_type2 %in% cell.2,
      radius == rad,
      !is.na(delta_estimate)
    ) %>%
    mutate(cell = paste(cell_type1, cell_type2, sep=" <-> "), .before = cell_type1)
  
  cell <- unique(df$cell)
  stopifnot(length(cell)==1)
  
  med <- median(df$delta_estimate)
  #print(med)
  
  df$delta_l <- factor(ifelse(df$delta_estimate > med, "high", "low"),
                       levels = c("low", "high"))
  df <- df %>% relocate(delta_l, .after = delta_estimate)
  
  if(returndata){return(df)}
  
  if(event_type == "OS"){
    df <- df %>% filter(!is.na(death), !is.na(time_OS_months)) %>%
      filter(time_OS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Overall survival"
    fit <- survfit(Surv(time=time_OS_months, event=death) ~ study_group + delta_l, data = df)
  } else {
    df <- df %>% filter(!is.na(distant_recurrence), !is.na(time_DRS_months)) %>%
      filter(time_DRS_months > 20 | reason_death == "BC") %>%
      as.data.frame()
    y <- "Distant recurrence"
    fit <- survfit(Surv(time=time_DRS_months, event=distant_recurrence) ~ study_group + delta_l, data = df)
  }
  
  ggsurvplot(
    fit, data = df,
    xlab = "Months", 
    ylab = paste(y,"probability"),
    title = paste(y, cell, "stratified by median delta L:", signif(med, 2)),
    palette = study_colors[names(study_colors) %in% unique(df$study_group)],
    ggtheme = theme_bw(),
    pval = T,
    facet.by = "delta_l",
    legend.title="Study group"
  )
}

km_CD8_Panck_lcross_OS <- lcross_km(
  lximmune, cell.1 = "CD3+CD8+", cell.2 = "PanCK+",
  event_type = "OS", returndata = F
)

km_CD8_Panck_lcross_OS
```

### Risk table 

```{r}
rt_km_CD8_Panck_lcross_OS <- faceted_risktable(km_CD8_Panck_lcross_OS, faceted.by = "delta_l") +
  ggtitle("Risk table for delta L: CD3+CD8+ <-> PanCK+ OS")
rt_km_CD8_Panck_lcross_OS
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD8_PancK_lcross_km_OS), width = 12, height = 7)
boxplot_CD8_Panck_lcross
km_CD8_Panck_lcross_OS
rt_km_CD8_Panck_lcross_OS
dev.off()
```

## Figure 4G - CD3+CD8+ <-> CD3+CD8- L cross boxplot and KM

### Boxplot

```{r}
boxplot_CD8_CD4_lcross <- lcross_boxplot(lximmune, cell.1 = "CD3+CD8+", cell.2 = "CD3+CD8-",
                                         label.y = 0.15, returndata=F)
boxplot_CD8_CD4_lcross
```

### Kaplan-meier

```{r}
km_CD8_CD4_lcross_OS <- lcross_km(lximmune, cell.1 = "CD3+CD8+", cell.2 = "CD3+CD8-",
                                  event_type = "OS", returndata = F)
km_CD8_CD4_lcross_OS
```

### Risk table 

```{r}
rt_km_CD8_CD4_lcross_OS <- faceted_risktable(km_CD8_CD4_lcross_OS, faceted.by = "delta_l") +
  ggtitle("Risk table for delta L: CD3+CD8+ <-> CD3+CD8- OS")
rt_km_CD8_CD4_lcross_OS
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD8_CD4_lcross_km_OS), width = 12, height = 7)
boxplot_CD8_CD4_lcross
km_CD8_CD4_lcross_OS
rt_km_CD8_CD4_lcross_OS
dev.off()
```

## Figure 4H - CD20+ <-> CD3+CD8+ L cross boxplot and KM

### Boxplot

```{r}
boxplot_CD20_CD8_lcross <- lcross_boxplot(lximmune, cell.1 = "CD20+", cell.2 = "CD3+CD8+",
                                          label.y = 0.126, returndata=F)
boxplot_CD20_CD8_lcross
```

### Kaplan-meier

```{r}
km_CD20_CD8_lcross_OS <- lcross_km(lximmune, cell.1 = "CD20+", cell.2 = "CD3+CD8+",
                                   event_type = "OS", returndata = F)
km_CD20_CD8_lcross_OS
```

### Risk table 

```{r}
rt_km_CD20_CD8_lcross_OS <- faceted_risktable(km_CD20_CD8_lcross_OS, faceted.by =  "delta_l") +
  ggtitle("Risk table for delta L: CD20+ <-> CD3+CD8+ OS")
rt_km_CD20_CD8_lcross_OS
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD20_CD8_lcross_km_OS), width = 12, height = 7)
boxplot_CD20_CD8_lcross
km_CD20_CD8_lcross_OS
rt_km_CD20_CD8_lcross_OS
dev.off()
```

## Figure 4I - CD20+ <-> CD3+CD8- L cross boxplot and KM

### Boxplot

```{r}
boxplot_CD20_CD4_lcross <- lcross_boxplot(lximmune, cell.1 = "CD20+", cell.2 = "CD3+CD8-",
                                          label.y = 0.22, returndata=F)
boxplot_CD20_CD4_lcross
```

### Kaplan-meier

```{r}
km_CD20_CD4_lcross_OS <- lcross_km(lximmune, cell.1 = "CD20+", cell.2 = "CD3+CD8-",
                                   event_type = "OS", returndata = F)
km_CD20_CD4_lcross_OS
```

### Risk table 

```{r}
rt_km_CD20_CD4_lcross_OS <- faceted_risktable(km_CD20_CD4_lcross_OS, faceted.by =  "delta_l") +
  ggtitle("Risk table for delta L CD20+ <-> CD3+CD8- OS")
rt_km_CD20_CD4_lcross_OS
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD20_CD4_lcross_km_OS), width = 12, height = 7)
boxplot_CD20_CD4_lcross
km_CD20_CD4_lcross_OS
rt_km_CD20_CD4_lcross_OS
dev.off()
```

## Sup fig 13A - KM CD20 density DRS

```{r}
km_CD20_density_DRS <- density_km(density, cell = "CD20+", event_type = "DRS", returndata = F)
km_CD20_density_DRS
```

### Risk table

```{r}
rt_km_CD20_density_DRS <- faceted_risktable(km_CD20_density_DRS, faceted.by = "density") +
  ggtitle("Risk table for CD20+ density DRS")
rt_km_CD20_density_DRS
```

### Save output

```{r}
pdf(file = here(params$km_CD20_density_DRS), width = 12, height = 7)
km_CD20_density_DRS
rt_km_CD20_density_DRS
dev.off()
```

## Sup fig 13B - KM CD20 L DRS

```{r}
km_CD20_l_DRS <- lstat_km(l, cell = "CD20+", event_type = "DRS", returndata = F)
km_CD20_l_DRS
```

### Risk table

```{r}
rt_km_CD20_l_DRS <- faceted_risktable(km_CD20_l_DRS, faceted.by = "delta_l") +
  ggtitle("Risk table for CD20+ L stat DRS")
rt_km_CD20_l_DRS
```

### Save output

```{r}
pdf(file = here(params$km_CD20_l_DRS), width = 12, height = 7)
km_CD20_l_DRS
rt_km_CD20_l_DRS
dev.off()
```

## Sup fig 14A- Boxplot CD20 <-> Panck L cross

```{r}
boxplot_CD20_Panck_lcross <- lcross_boxplot(lximmune, cell.1 = "CD20+", cell.2 = "PanCK+",
                                            label.y = 0.06, returndata=F)
boxplot_CD20_Panck_lcross
```

### Save output

```{r}
pdf(file = here(params$boxplot_CD20_Panck_lcross), width = 12, height = 7)
boxplot_CD20_Panck_lcross
dev.off()
```

## Sup fig 14B - KM OS CD20+ <-> Panck L cross

```{r}
km_CD20_Panck_lcross_km_OS <- lcross_km(lximmune, cell.1 = "CD20+", cell.2 = "PanCK+",
                                  event_type = "OS", returndata = F)
km_CD20_Panck_lcross_km_OS
```

### Risk table 

```{r}
rt_km_CD20_Panck_lcross_km_OS <- faceted_risktable(km_CD20_Panck_lcross_km_OS, faceted.by = "delta_l") +
  ggtitle("Risk table for delta L: OS CD20+ <-> Panck L cross")
rt_km_CD20_Panck_lcross_km_OS
```

### Save output

```{r}
pdf(file = here(params$km_CD20_Panck_lcross_km_OS), width = 12, height = 7)
km_CD20_Panck_lcross_km_OS
rt_km_CD20_Panck_lcross_km_OS
dev.off()
```

## Sup fig 14C - KM DRS CD20+ <-> Panck L cross

```{r}
km_CD20_Panck_lcross_km_DRS <- lcross_km(lximmune, cell.1 = "CD20+", cell.2 = "PanCK+",
                                         event_type = "DRS", returndata = F)
km_CD20_Panck_lcross_km_DRS
```

### Risk table 

```{r}
rt_km_CD20_Panck_lcross_km_DRS <- faceted_risktable(km_CD20_Panck_lcross_km_DRS, faceted.by = "delta_l") +
  ggtitle("Risk table for delta L: DRS CD20+ <-> Panck L cross")
rt_km_CD20_Panck_lcross_km_DRS
```

### Save output

```{r}
pdf(file = here(params$km_CD20_Panck_lcross_km_DRS), width = 12, height = 7)
km_CD20_Panck_lcross_km_DRS
rt_km_CD20_Panck_lcross_km_DRS
dev.off()
```

## Sup fig 23 - Misc spatial drs

### KMs

L stat:

```{r}
misc_km_drs <- list()

misc_km_drs[[length(misc_km_drs)+1]] <- lstat_km(
  l, cell = "CD3+CD8+", event_type = "DRS", returndata = F
  )

names(misc_km_drs) <- "CD3+CD8+"
```

L cross:

```{r}
drs_cross <- tibble(
  cell1=c("CD3+CD8+","CD3+CD8+", "CD20+", "CD20+"),
  cell2=c("CD3+CD8-", "PanCK+","CD3+CD8-","CD3+CD8+")
)

drs_cross
```

```{r}
for(i in 1:nrow(drs_cross)){
 misc_km_drs[[length(misc_km_drs)+1]] <- lcross_km(
   lximmune, cell.1 = drs_cross$cell1[i], cell.2 = drs_cross$cell2[i],
   event_type = "DRS", returndata = F
 )
 names(misc_km_drs)[length(misc_km_drs)] <- paste(drs_cross$cell1[i],
                                                  drs_cross$cell2[i],
                                                  sep = " <-> ")
}
```

```{r}
misc_km_drs
```

### Risk table 

```{r}
rt_misc_km_drs <- lapply(names(misc_km_drs), function(x){
  faceted_risktable(misc_km_drs[[x]], faceted.by = "delta_l") +
    ggtitle(paste("Risk table for delta L: DRS", x))
}) %>%
  set_names(paste("Risk table",names(misc_km_drs)))

rt_misc_km_drs
```

### Save output

```{r}
pdf(file = here(params$misc_spatial_drs), width = 12, height = 7)
misc_km_drs
rt_misc_km_drs
dev.off()
```

## Session info

```{r}
sessionInfo()
```
